---
name: geom_abline
permalink: ggplot2/geom_abline/
description: How to use the abline geom in ggplot2 to add a line with specified slope and intercept to the plot.
layout: base
thumbnail: thumbnail/ipython_graph_email.jpg
language: ggplot2
page_type: example_index
display_as: basic
order: 1
redirect_from: ggplot2/line-shapes/
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Line
add line for mean using <code>geom_vline</code>

```{r}
library(plotly)

set.seed(1234)
dat <- data.frame(cond = factor(rep(c("A","B"), each=200)),
                   rating = c(rnorm(200),rnorm(200, mean=.8)))

fig <-  ggplot(dat, aes(x=rating)) +
    geom_histogram(binwidth=.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(rating, na.rm=T)),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1)

fig <-  ggplotly(p)


fig
```

### Histogram
overlaid histograms with <code>geom_vline</code>

```{r}
library(plotly)
library(plyr)
cdat <- ddply(dat, "cond", summarise, rating.mean=mean(rating))

# Overlaid histograms with means
fig <-  ggplot(dat, aes(x=rating, fill=cond)) +
    geom_histogram(binwidth=.5, alpha=.5, position="identity") +
    geom_vline(data=cdat, aes(xintercept=rating.mean),
               linetype="dashed", size=1)

fig <-  ggplotly(p)


fig
```

### Histogram Means
histograms with <code>geom_vline</code> means

```{r}
library(plotly)
library(plyr)
cdat <- ddply(dat, "cond", summarise, rating.mean=mean(rating))

# With mean lines
fig <-  ggplot(dat, aes(x=rating)) + geom_histogram(binwidth=.5, colour="black", fill="white") +
    facet_grid(cond ~ .) +
    geom_vline(data=cdat, aes(xintercept=rating.mean),
               linetype="dashed", size=1, colour="red")

fig <-  ggplotly(p)


fig
```

### Density Plots
density plots with <code>geom_vline</code> means

```{r}
library(plotly)
library(plyr)
cdat <- ddply(dat, "cond", summarise, rating.mean=mean(rating))

# Density plots with means
fig <-  ggplot(dat, aes(x=rating, colour=cond)) +
    geom_density() +
    geom_vline(data=cdat, aes(xintercept=rating.mean),
               linetype="dashed", size=1)


fig <-  ggplotly(p)


fig
```

### Horizontal Line
add horizontal line with <code>geom_hline</code>

```{r}
library(plotly)

dat <- read.table(header=TRUE, text='
      cond xval yval
   control 11.5 10.8
   control  9.3 12.9
   control  8.0  9.9
   control 11.5 10.1
   control  8.6  8.3
   control  9.9  9.5
   control  8.8  8.7
   control 11.7 10.1
   control  9.7  9.3
   control  9.8 12.0
 treatment 10.4 10.6
 treatment 12.1  8.6
 treatment 11.2 11.0
 treatment 10.0  8.8
 treatment 12.9  9.5
 treatment  9.1 10.0
 treatment 13.4  9.6
 treatment 11.6  9.8
 treatment 11.5  9.8
 treatment 12.0 10.6
')

# The basic scatterplot
fig <-  ggplot(dat, aes(x=xval, y=yval, colour=cond)) + 
  geom_point()

# Add a horizontal line
fig <-  p + geom_hline(aes(yintercept=10))

fig <-  ggplotly(p)


fig
```

### Mean Line
add mean line with <code>geom_hline</code>

```{r}
library(plotly)

dat <- read.table(header=TRUE, text='
      cond xval yval
   control 11.5 10.8
   control  9.3 12.9
   control  8.0  9.9
   control 11.5 10.1
   control  8.6  8.3
   control  9.9  9.5
   control  8.8  8.7
   control 11.7 10.1
   control  9.7  9.3
   control  9.8 12.0
 treatment 10.4 10.6
 treatment 12.1  8.6
 treatment 11.2 11.0
 treatment 10.0  8.8
 treatment 12.9  9.5
 treatment  9.1 10.0
 treatment 13.4  9.6
 treatment 11.6  9.8
 treatment 11.5  9.8
 treatment 12.0 10.6
')

# The basic scatterplot
fig <-  ggplot(dat, aes(x=xval, y=yval, colour=cond)) + 
  geom_point()

mean1 <- mean(dat[dat$cond == "control", "xval"])
mean2 <- mean(dat[dat$cond == "treatment", "xval"])

# Add colored lines for the mean xval of each group
fig <-  p + 
  geom_vline(aes(xintercept=mean1), colour="green") + 
  geom_vline(aes(xintercept=mean2), colour="lightblue")

fig <-  ggplotly(p)


fig
```

### Geom_vline & Geom_hline
use <code>geom_vline</code> with <code>geom_hline</code>

```{r}
library(plotly)

dat <- read.table(header=TRUE, text='
      cond xval yval
   control 11.5 10.8
   control  9.3 12.9
   control  8.0  9.9
   control 11.5 10.1
   control  8.6  8.3
   control  9.9  9.5
   control  8.8  8.7
   control 11.7 10.1
   control  9.7  9.3
   control  9.8 12.0
 treatment 10.4 10.6
 treatment 12.1  8.6
 treatment 11.2 11.0
 treatment 10.0  8.8
 treatment 12.9  9.5
 treatment  9.1 10.0
 treatment 13.4  9.6
 treatment 11.6  9.8
 treatment 11.5  9.8
 treatment 12.0 10.6
')

# The basic scatterplot
fig <-  ggplot(dat, aes(x=xval, y=yval, colour=cond)) + geom_point()

# Add a red dashed vertical line
fig <-  p + geom_hline(aes(yintercept=10)) +
    geom_vline(aes(xintercept=11.5), colour="#BB0000", linetype="dashed")

fig <-  ggplotly(p)

fig
```

These ggplot2 examples were inspired by <a href="http://www.cookbook-r.com/Graphs/Lines_(ggplot2)/">the Cookbook for R</a>.
