---
name: Intro to Animations
permalink: ggplot2/animations/
description: How to create animations in ggplot2 with Plotly.
layout: base
thumbnail: thumbnail/animations.gif
language: ggplot2
page_type: example_index
display_as: animations
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Frames

Now, along with `data` and `layout`, `frames` is added to the keys that `figure` allows. Your `frames` key points to a list of figures, each of which will be cycled through upon instantiation of the plot.

### Basic Example

```{r}
library(plotly)

df <- data.frame(
  x = c(1,2,3,4), 
  y = c(1,2,3,4), 
  f = c(1,2,3,4)
)

fig <-  ggplot(df, aes(x, y)) +
    geom_point(aes(frame = f))

fig <-  ggplotly(p)

fig
```

### Mulitple Trace Animations

```{r}
library(plotly)
library(gapminder)


fig <-  ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10()

fig <-  ggplotly(p)

fig
```

### Add Animation Options

```{r}
library(plotly)

fig <-  p %>% 
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  )

fig
```

### Add Button Options

```{r}
library(plotly)

fig <-  p %>% 
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  )

fig
```

### Add Slider Options

```{r}
library(plotly)

fig <-  p %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )

fig
```

### Advanced Example

```{r}
library(plotly)
library(gapminder)

fig <-  ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10()

fig <-  ggplotly(p) %>% 
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  ) %>% 
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )

fig
```

### Reference 

To read more on animations see [The Plotly Book](https://cpsievert.github.io/plotly_book/key-frame-animations.html).
