---
name: geom_contour
permalink: ggplot2/geom_contour/
description: How to make a contour in ggplot2 using geom_contour.
layout: base
thumbnail: thumbnail/geom_contour.jpg
language: ggplot2
page_type: example_index
display_as: basic
order: 3
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic geom\_contour plot
geom\_contour produces a similar output to geom\_density\_2d, except it uses a third variable for the values rather than frequency. The volcano dataset comes pre-loaded on R.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

fig <-  ggplot(df, aes(Var1, Var2, z= value)) +
  geom_contour() +
  scale_fill_distiller(palette = "Spectral", direction = -1)
fig <-  ggplotly(p)

fig
```

### Coloured Plot
[See here](https://ggplot2.tidyverse.org/reference/scale_brewer.html) for a list of colour palettes that come with the brewer (discrete) and distiller (continuous) packages.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

fig <-  ggplot(df, aes(Var1, Var2, z= value, colour=stat(level))) +
  geom_contour() +
  scale_colour_distiller(palette = "YlGn", direction = 1)

ggplotly(p)
```

### Filled Plot
It's possible to colour in each of the layers, by changing geom\_contour to stat\_contour as below. As the edges of the graph indicate, filled contour plots only work when each layer is an enclosed shape rather than an open line; a geom more suited to this functionality would be geom\_tile or geom\_raster.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

fig <-  ggplot(df, aes(Var1, Var2, z= value)) +
  stat_contour(geom="polygon",aes(fill=stat(level))) +
  scale_fill_distiller(palette = "Spectral", direction = -1)

ggplotly(p)
```

