---
name: geom_hex
permalink: ggplot2/geom_hex/
description: How to make a hexagonal two-dimensional heatmap in ggplot2 using geom_hex. Examples of coloured and facetted graphs.
layout: base
thumbnail: thumbnail/geom_hex.jpg
language: ggplot2
page_type: example_index
display_as: statistical
order: 3
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic 2d Heatmap
See also [geom_bin2d](https://plot.ly/ggplot2/geom_bin2d/) for a similar geom with rectangular bins. Note: facetting is supported in geom\_bin2d but not geom\_hex.

Source: [Department of Canadian Heritage](https://open.canada.ca/data/en/dataset/a0bff264-1c80-41ee-aef9-e7da347c5158)

```{r}
library(plotly)

english_french <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/english_french.csv",stringsAsFactors = FALSE)

fig <-  ggplot(english_french, aes(x=engperc,y=frenperc)) + 
  geom_hex() +
  labs(title = "Distribution of Canadian areas by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "# of census \nsubdivisions")
fig <-  ggplotly(p)

fig
```

### Customized Colours
Let's flip the colour scheme so that lighter colours denote larger numbers than darker colours. We should also move to a logarithmic scale, since as it is, the very large value in the bottom right overshadows all other values.

```{r}
library(plotly)

english_french <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/english_french.csv",stringsAsFactors = FALSE)

fig <-  ggplot(english_french, aes(x=engperc,y=frenperc)) + 
  geom_hex() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of Canadian towns by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "# of census \nsubdivisions")
fig <-  ggplotly(p)

fig
```

### Weighted Data
In the previous graphs, each observation represented a single census subdivision - this counted small towns of 500 people equally with cities like Montreal and Toronto. We can weight the data by the "total" column (i.e. total population) to make this a graph of population.

```{r}
library(plotly)

english_french <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/english_french.csv",stringsAsFactors = FALSE)

fig <-  ggplot(english_french, aes(x=engperc, y=frenperc, weight=total)) + 
  geom_hex() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of the Canadian population by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "population")
fig <-  ggplotly(p)

fig
```

### Customized Appearance
We can modify the graph's appearance - for example, if the grey background makes it difficult to make out the paler shades of blue, we can change the theme to one with a white background. Included also is a way to change the font. You can find a list [here](http://ggplot2.tidyverse.org/reference/theme.html) of all the theme elements that you can modify.

```{r}
library(plotly)

english_french <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/english_french.csv",stringsAsFactors = FALSE)

fig <-  ggplot(english_french, aes(x=engperc,y=frenperc, weight=total)) + 
  geom_hex(bins = 20) +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of Canadian towns by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "population") +
  theme_bw() +
  theme(text = element_text(family = 'Fira Sans'))
fig <-  ggplotly(p)

fig
```

