---
name: geom_raster
permalink: ggplot2/geom_raster/
description: How to make a 2-dimensional heatmap in ggplot2 using geom_raster.
layout: base
thumbnail: thumbnail/geom_raster.jpg
language: ggplot2
page_type: example_index
display_as: basic
order: 7
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic 2d Heatmap
geom\_raster creates a coloured heatmap, with two variables acting as the x- and y-coordinates and a third variable mapping onto a colour. (It is coded similarly to geom\_tile and is generated more quickly.) This uses the volcano dataset that comes pre-loaded with R.

```{r}
library(reshape2)
library(plotly)

df <- melt(volcano)

fig <-  ggplot(df, aes(Var1, Var2)) +
  geom_raster(aes(fill=value)) +
  labs(x="West to East",
       y="North to South",
       title = "Elevation map of Maunga Whau")

ggplotly(p)
```

### Customized 2d Heatmap
This uses the Spectral palette from [ColorBrewer](https://ggplot2.tidyverse.org/reference/scale_brewer.html); a full list of palettes is here.

```{r}
library(reshape2)
library(plotly)

df <- melt(volcano)

fig <-  ggplot(df, aes(Var1, Var2)) +
  geom_raster(aes(fill=value)) +
  scale_fill_distiller(palette = "Spectral", direction = -1) +
  labs(x="West to East",
       y="North to South",
       title = "Elevation map of Maunga Whau",
       fill = "Elevation") +
  theme(text = element_text(family = 'Fira Sans'),
        plot.title = element_text(hjust = 0.5))

ggplotly(p)
```

