---
description: How to Style Markers in ggplot2 with Plotly.
name: Styling Markers
permalink: ggplot2/marker-style/
thumbnail: thumbnail/marker-style.gif
layout: base
language: ggplot2
display_as: file_settings
page_type: u-guide
order: 19
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

# Default point plot

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(wt, mpg))
p <-  p + geom_point()

ggplotly(p)
```
# Add colour

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(wt, mpg))
p <-  p + geom_point(aes(colour = factor(cyl)))

ggplotly(p)
```

# Changing shapes of data points


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(wt, mpg))
p <-  p + geom_point(aes(shape = factor(cyl)))

ggplotly(p)
```

# Changing size of data points


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(wt, mpg))
p <-  p + geom_point(aes(size = qsec))

ggplotly(p)
```


# Manually setting aesthetics


```{r}
library(plotly)
library(ggplot2)

p <-  ggplot(mtcars, aes(wt, mpg)) + geom_point(colour = "red", size = 3)

ggplotly(p)
```

# Optional shape arguments

For shapes that have a border (like shape 21), you can colour the inside and outside separately. Use the stroke aesthetic to modify the width of the border.


```{r}
library(plotly)
library(ggplot2)

p <-    
 ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "whiet", size = 5, stroke = 5)
  
ggplotly(p)
```


# Mix multiples shapes

You can create interesting shapes by layering multiple points of different sizes.

Default plot:


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(mpg, wt, shape = factor(cyl)))
p <-    
 p +
  geom_point(aes(colour = factor(cyl)), size = 4) +
  geom_point(colour = "grey90", size = 1.5)
  
ggplotly(p)
```


Mixed shapes:


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(mpg, wt, shape = factor(cyl)))
p <-    
 p +
  geom_point(colour = "black", size = 4.5) +
  geom_point(colour = "pink", size = 4) +
  geom_point(aes(shape = factor(cyl)))
  
ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->
