---
name: Heatmaps
permalink: ggplot2/heatmaps/
description: How to make Heatmaps plots in ggplot2 with Plotly.
layout: base
thumbnail: thumbnail/heatmap.jpg
language: ggplot2
page_type: u-guide
display_as: statistical
order: 2
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic 2d Heatmap
See also geom\_hex for a similar geom with hexagonal bins. Note: facetting is supported in geom\_bin2d but not geom\_hex.

Source: [Department of Canadian Heritage](https://open.canada.ca/data/en/dataset/a0bff264-1c80-41ee-aef9-e7da347c5158)

```{r}
library(plotly)

english_french <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/english_french.csv",stringsAsFactors = FALSE)

p <- ggplot(english_french, aes(x=engperc,y=frenperc)) + 
  geom_bin2d() +
  labs(title = "Distribution of Canadian areas by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "# of census \nsubdivisions")
ggplotly(p)
```

`geom_raster` creates a coloured heatmap, with two variables acting as the x- and y-coordinates and a third variable mapping onto a colour. (It is coded similarly to geom\_tile and is generated more quickly.) This uses the volcano dataset that comes pre-loaded with R.

```{r}
library(reshape2)
library(plotly)

df <- melt(volcano)

p <- ggplot(df, aes(Var1, Var2)) +
  geom_raster(aes(fill=value)) +
  labs(x="West to East",
       y="North to South",
       title = "Elevation map of Maunga Whau")

ggplotly(p)
```

### Customized Colours
This uses the Spectral palette from [ColorBrewer](https://ggplot2.tidyverse.org/reference/scale_brewer.html); a full list of palettes is here.

```{r}
library(reshape2)
library(plotly)

df <- melt(volcano)

p <- ggplot(df, aes(Var1, Var2)) +
  geom_raster(aes(fill=value)) +
  scale_fill_distiller(palette = "Spectral", direction = -1) +
  labs(x="West to East",
       y="North to South",
       title = "Elevation map of Maunga Whau",
       fill = "Elevation") +
  theme(text = element_text(family = 'Fira Sans'),
        plot.title = element_text(hjust = 0.5))

ggplotly(p)
```


Let's flip the colour scheme so that lighter colours denote larger numbers than darker colours. We should also move to a logarithmic scale, since as it is, the very large value in the bottom right overshadows all other values.

```{r}
library(plotly)

p <- ggplot(english_french, aes(x=engperc,y=frenperc)) + 
  geom_bin2d() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of Canadian towns by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "# of census \nsubdivisions")
ggplotly(p)
```

### Weighted Data
In the previous graphs, each observation represented a single census subdivision - this counted small towns of 500 people equally with cities like Montreal and Toronto. We can weight the data by the "total" column (i.e. total population) to make this a graph of population.

```{r}
library(plotly)

p <- ggplot(english_french, aes(x=engperc, y=frenperc, weight=total)) + 
  geom_bin2d() +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of the Canadian population by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "population")
fig <- ggplotly(p)


fig
```

### With Facets
We can facet the graphic with the "region" column, and set "bins" to 20, so that the graph is 20 x 20 sides. 

```{r}
library(plotly)

p <- ggplot(english_french, aes(x=engperc,y=frenperc, weight=total)) + 
  geom_bin2d(bins = 20) +
  facet_wrap(~factor(region, levels = c("Atlantic","Québec","Ontario","Prairies","British Columbia"))) +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of Canadian towns by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "population")
fig <- ggplotly(p)


fig
```

### Customized Appearance
We can modify the graph's appearance - for example, if the grey background makes it difficult to make out the paler shades of blue, we can change the theme to one with a white background. Included also is a way to change the font.

```{r}
library(plotly)

p <- ggplot(english_french, aes(x=engperc,y=frenperc, weight=total)) + 
  geom_bin2d(bins = 20) +
  facet_wrap(~factor(region, levels = c("Atlantic","Québec","Ontario","Prairies","British Columbia"))) +
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") +
  labs(title = "Distribution of Canadian towns by English and French fluency",
       x = "% fluent in English",
       y = "% fluent in French",
       fill = "population") +
  theme_bw() +
  theme(text = element_text(family = 'Fira Sans'))
fig <- ggplotly(p)


fig
```

### Basic geom\_tile graph
This graph, compiled by [Jeff Zimmerman](https://docs.google.com/spreadsheets/d/1HI-dikWN64clxSGJu2QV8C64R9Bkzt8K-jFaeHj4X7k/edit#gid=0), shows how often hitters swing and miss at fastballs, based on their velocity and spin rate. Colour schemes are from ColorBrewer; a complete list of palettes is available [here](https://ggplot2.tidyverse.org/reference/scale_brewer.html).

```{r}
library(plotly)
spinrates <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/spinrates.csv",
                      stringsAsFactors = FALSE)

p <- ggplot(spinrates, aes(x=velocity, y=spinrate)) +
  geom_tile(aes(fill = swing_miss)) +
  scale_fill_distiller(palette = "YlGnBu") +
  labs(title = "Likelihood of swinging and missing on a fastball",
       y = "spin rate (rpm)")

ggplotly(p)
```

### Adjusting appearance
The *direction* option sets which side of the colour scheme maps onto the low values and which side maps onto the high; it defaults to -1 but could be adjusted to 1.

Also [adjusted the theme](https://ggplot2.tidyverse.org/reference/ggtheme.html).

```{r}
library(plotly)
spinrates <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/spinrates.csv",
                      stringsAsFactors = FALSE)

p <- ggplot(spinrates, aes(x=velocity, y=spinrate)) +
  geom_tile(aes(fill = swing_miss)) +
  scale_fill_distiller(palette = "YlGnBu", direction = 1) +
  theme_light() +
  labs(title = "Likelihood of swinging and missing on a fastball",
       y = "spin rate (rpm)")

ggplotly(p)
```

### geom\_tile with viridis colour scheme
[Viridis colour schemes](https://ggplot2.tidyverse.org/reference/scale_brewer.html) are uniform in both colour and black-and-white, as well as for those with colour-blindness. There are five colour schemes: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").

```{r}
library(plotly)
spinrates <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/spinrates.csv",
                      stringsAsFactors = FALSE)

p <- ggplot(spinrates, aes(x=velocity, y=spinrate)) +
  geom_tile(aes(fill = swing_miss)) +
  scale_fill_viridis_c(option = "B", direction = -1) +
  labs(title = "Likelihood of swinging and missing on a fastball",
       y = "spin rate (rpm)") +
  theme_light()

ggplotly(p)
```

