---
description: How to work with Legends in ggplot2 with Plotly.
name: Legends
permalink: ggplot2/legend/
thumnail_github: legends.gif
layout: base
language: ggplot2
display_as: file_settings
page_type: u-guide
order: 14
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Default box plot


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Change the legend position

The position of the legend can be changed using the function `theme()` as follow:


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p + theme(legend.position="bottom")

ggplotly(p)
```
Note that, the argument `legend.position` can be also a numeric vector `c(x,y)`. In this case it is possible to position the legend inside the plotting area. x and y are the coordinates of the legend box. Their values should be between 0 and 1. `c(0,0)` corresponds to the `bottom left` and `c(1,1)` corresponds to the `top right` position.


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p + theme(legend.position = c(0.8, 0.2))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Change the legend title and text font styles



```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p <- p + theme(legend.title = element_text(colour="red", size=10, 
                                      face="italic"))
p <- p + theme(legend.text = element_text(colour="blue", size=10, 
                                     face="bold"))

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Change the background color of the legend box



```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p <- p + theme(legend.background = element_rect(fill="lightblue",
                                  size=0.5, linetype="solid", 
                                  colour ="red"))

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Change the order of legend items



```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p + scale_x_discrete(limits=c("2", "0.5", "1"))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Remove the plot legend



```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()

p <- p + theme(legend.title = element_blank())
p <- p + theme(legend.position='none')

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->
