---
description: How to make Network Graphs in ggplot2 with Plotly.
name: Network Graphs
permalink: ggplot2/network-graphs/
thumbnail: thumbnail/net.jpg
layout: base
language: ggplot2
display_as: scientific
page_type: u-guide
order: 12
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Random graph

Let’s start with an undirected Bernoulli random graph, with 10 nodes named “a, b, …, i, j”, and a rather high likelihood of an edge to exist between them:

```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

# vertex names
network.vertex.names(net) = letters[1:10]

p <- ggnet2(net)


ggplotly(p)
```
The `net` argument is the only compulsory argument of `ggnet2`. It can be a `network` object or any object that can be coerced to that class through its `edgeset.constructors` functions, such as adjacency matrixes, incidence matrixes and edge lists.

If the `intergraph` package is installed, `net` can also be an `igraph` one-mode network object, which is the only type of network that the package can convert from the `igraph` to the `network` class.

<!--------------------- EXAMPLE BREAK ------------------------->

## Node color and size

The most basic properties that one might want to change at that stage are the size and color of the nodes, or the size and color of the edges. Let’s modify each of these properties:

```{r}
library(plotly)
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

# vertex names
network.vertex.names(net) = letters[1:10]

p <- ggnet2(net, node.size = 6, node.color = "black", edge.size = 1, edge.color = "grey")

ggplotly(p)
```
The vertex-related arguments of `ggnet2` start with `node`, and its edge-related arguments start with edge. The `node.color` and node.size arguments can be abbreviated:
`ggnet2(net, size = 6, color = "black", edge.size = 1, edge.color = "grey")`

It also possible to pass a vector of node colors directly to `ggnet2`, as long as it has the same number of elements as the network has nodes:


```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

# vertex names
network.vertex.names(net) = letters[1:10]

p <- ggnet2(net, size = 6, color = rep(c("tomato", "steelblue"), 5))

ggplotly(p)
```
The color, shape, size and transparency of nodes can all be set through these basic methods, or by passing a vertex attribute to them. Let’s first see how to position the nodes.

<!--------------------- EXAMPLE BREAK ------------------------->

## Node colors

Let’s now assign a vertex attribute called phono, which indicates whether the name of the vertex is a vowel or a consonant.
This attribute can be passed to `ggnet2` to indicate that the nodes belong to a group. All the user has to do is to pass the name of the vertex attribute to the color argument, which will find it in the list of vertex attributes and use it to map the colors of the nodes:

```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")

p <- ggnet2(net, color = "phono")

ggplotly(p)
```
By default, `ggnet2` assigns a grayscale color to each group. To modify this behavior, let’s review three different options. The first one consists in “hard-coding” the colors into the graph by assigning them to a vertex attribute, and then in passing this attribute to `ggnet2`:

```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")

p <- ggnet2(net, color = "color")

ggplotly(p)
```

`ggnet2` returns a `ggplot` object, so the underlying data can be accessed by requesting the `data` component of the plot. The structure of that component always contains the following columns, which match the names of `ggplot2` arguments.
```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")

ggnet2(net, color = "phono", size = 1:10)$data
```

This means that you can append any `ggplot2` component to the graph by passing additional aesthetics to it, which allows for a fair amount of “plot hacking”. In this example, we use `ggnet2` to get the basic data structure in place, while sizing the nodes to 0. The nodes are then plotted manually, by overlaying several `geom` objects:
```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")

p <- ggnet2(net, color = "phono", palette = "Set1", size = 0) +
  geom_point(aes(color = color), size = 12, color = "white") +
  geom_point(aes(color = color), size = 12, alpha = 0.5) +
  geom_point(aes(color = color), size = 9) +
  geom_text(aes(label = toupper(substr(color, 1, 1))), color = "white", fontface = "bold") +
  guides(color = FALSE)

ggplotly(p)
```
<!--------------------- EXAMPLE BREAK ------------------------->

## Node size

It is common to size the nodes of a network by their centrality or by some other indicator of interest. Just like its `color` argument, the `size` argument of `ggnet2` can take a single numeric value, a vector of values, or a vertex attribute:

```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")

p <- ggnet2(net, size = "phono", size.palette = c("vowel" = 10, "consonant" = 1))

ggplotly(p)
```

When the `size` attribute is not a single numeric value, the maximum size of the nodes is determined by the `max_size` argument, just like in the `scale_size_area` controller of `ggplot2`, which `ggnet2` emulates to compute the relative size of the nodes:

`ggnet2(net, size = sample(0:2, 10, replace = TRUE), max_size = 9)`

`ggnet2` can also size nodes by calculating their in-degree, out-degree, or total (Freeman) degree, using the `degree` function of the `sna` package. All the user has to do is to pass the `indegree`, `outdegree`, or `freeman` option to the `weight` argument (`degree` is also understood, and is equivalent to `freeman`).

`ggnet2` gives the user further control over the node size by providing a quick way to cut the node sizes into quantiles, using the `size.cut` argument. If set to `TRUE`, it defaults to quartiles, but any numeric value above 1 is acceptable:


```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")

p <- ggnet2(net, size = "degree", size.cut = 3)

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Adding labels

Through the `label` argument, `ggnet2` can label the nodes of a network by using their vertex names, another vertex attribute, or any other vector of labels.

The size of the labels, which is automatically set to half of the node size, is controlled by the `label.size` argument, their color by the `label.color` argument, and their level of transparency by the `label.alpha` argument:

```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

p <-ggnet2(net, size = 12, label = TRUE, label.alpha = 0.75, label.size = 5, color = "black", label.color = "white")

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Changing shapes

The shapes and transparency of the nodes can be set exactly like the color and size of the nodes, either through a single value, a vector of (numeric) values, or a vertex attribute. This allows to create nodes that can be distinguished even in the plot loses its colors:
```{r}
library(plotly)
library(ggnet)
library(network)
library(sna)
library(ggplot2)

# random graph
net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)

net %v% "phono" = ifelse(letters[1:10] %in% c("a", "e", "i"), "vowel", "consonant")
net %v% "color" = ifelse(net %v% "phono" == "vowel", "steelblue", "tomato")


p <-ggnet2(net, color = "phono", shape = "phono")

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->