---
description: How to make Random Walk in ggplot2 with Plotly.
name: Random Walk
permalink: ggplot2/random-walk/
thumnail_github: static-image
layout: base
language: ggplot2
display_as: advanced
page_type: u-guide
order: 2
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

A random walk can be thought of as a random process in which a token or a marker is randomly moved around some space, that is, a space with a metric used to compute distance. In the visualizations below, we will be using scatter plots as well as a colorscale to denote the time sequence of the walk.

## Random Walk in 1D

The jitter in the data points along the x and y axes are meant to illuminate where the points are being drawn and what the tendency of the random walk is.

```{r}
library(plotly)
library(ggplot2)

P1 <- cumsum(rnorm(1000))
dat <- data.frame(x = seq_along(P1), y = P1)
p <- ggplot(dat, aes(x = x, y = y)) + geom_line()

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Random Walk in 2D


```{r}
library(plotly)
require(ggplot2)

random_walk <- function(n.org, steps, left.p = .5, up.p = .5, plot = TRUE){
 
 
whereto <- matrix(ncol = 2)
 
for(x in 1:n.org){
walker <- matrix(c(0,0), nrow = steps+1, ncol = 2, byrow = T)
 
for(i in 1:steps){
# left/right = 1/0
horizontal <- rbinom(1, 1, left.p)
 
# distance 2
h.dist <- abs(rnorm(1, 0, 1))
 
# Horizontal Movement
if(horizontal == 0){
walker[i+1,1] <- walker[i,1] + h.dist
}
if(horizontal == 1){
walker[i+1,1] <- walker[i,1] - h.dist
}
 
# up/down = 1/0
vertical <- rbinom(1, 1, up.p)
 
#distance 2
v.dist <- abs(rnorm(1, 0, 1))
 
# Vertical Movement
if(vertical == 1){
walker[i+1,2] <- walker[i,2] + v.dist
}
if(vertical == 0){
walker[i+1,2] <- walker[i,2] - v.dist
}
}
 
whereto <- rbind(whereto, walker)
}
 
id <- rep(1:n.org, each = 1001)
colnames(whereto) <- c("x" , "y")
whereto <- as.data.frame(whereto)
whereto <- cbind(whereto[2:nrow(whereto),], org = factor(id))
 
p <- ggplot(whereto, aes(x = x, y = y, colour = org))
p <- p + geom_path()
 
return(p)
}
 
p <- random_walk(1, 1000, .5, .5)

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->