---
description: How to make Bar Plots plots ggplot2 with Plotly.
name: Bar Plots
display_as: basic
order: 3
permalink: ggplot2/bar-charts/
thumbnail: thumbnail/bar.jpg
layout: base
language: ggplot2
page_type: u-guide
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

`geom_bar` is designed to make it easy to create bar charts that show counts (or sums of weights).

# Default bar plot

```{r}
library(plotly)
g <- ggplot(mpg, aes(class))  
p <-  g + geom_bar()

ggplotly(p)
```



```{r}
library(plotly)
g <- ggplot(mpg, aes(class))  
p <-  g + geom_bar(aes(weight = displ))

ggplotly(p)
```

# Add colour 

```{r}
library(plotly)

dat <- data.frame(
    time = factor(c("Lunch","Dinner"), levels=c("Lunch","Dinner")),
    total_bill = c(14.89, 17.23)
)

p <- ggplot(data=dat, aes(x=time, y=total_bill, fill=time)) +
    geom_bar(stat="identity")

fig <- ggplotly(p)

fig
```

Setting custom colours:

```{r}
library(plotly)

dat1 <- data.frame(
    sex = factor(c("Female","Female","Male","Male")),
    time = factor(c("Lunch","Dinner","Lunch","Dinner"), levels=c("Lunch","Dinner")),
    total_bill = c(13.53, 16.81, 16.24, 17.42)
)

p <- ggplot(data=dat1, aes(x=time, y=total_bill, fill=sex)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") +
    scale_fill_manual(values=c("#999999", "#E69F00"))

fig <- ggplotly(p)

fig
```


# Stacking bar plots

Bar plots are automatically stacked when multiple bars are at the same location. The order of the fill is designed to match the legend.

```{r}
library(plotly)
g <- ggplot(mpg, aes(class))  
p <-  g + geom_bar(aes(fill = drv))

ggplotly(p)
```



# Showing mean

```{r}
library(plotly)
df <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))  
p <-    
 ggplot(df, aes(trt, outcome)) +
  geom_col()
  
ggplotly(p)
```


`geom_point()` displays exactly the same information and doesn't require the y-axis to touch zero.


```{r}
library(plotly)
df <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))  
p <-    
 ggplot(df, aes(trt, outcome)) +
  geom_point()
  
ggplotly(p)
```


You can also use `geom_bar()` with continuous data, in which case it will show counts at unique locations.

```{r}
library(plotly)
df <- data.frame(x = rep(c(2.9, 3.1, 4.5), c(5, 10, 4)))  
p <-  ggplot(df, aes(x)) + geom_bar()

ggplotly(p)
```



# Using binwidth 


```{r}
library(plotly)
df <- data.frame(x = rep(c(2.9, 3.1, 4.5), c(5, 10, 4)))  
p <-  ggplot(df, aes(x)) + geom_histogram(binwidth = 0.5)

ggplotly(p)
```


# Error Bars
barplot with error bars

```{r}
library(plotly)
library(dplyr)
set.seed(123)

df <- diamonds[sample(1:nrow(diamonds), size = 1000),]

df.summ <- df %>% group_by(cut) %>% summarize(Mean = mean(table), Min = min(table), Max = max(table))

p <- ggplot(df.summ, aes(x = cut, y = Mean, ymin = Min, ymax = Max, fill = cut)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar() + 
  ggtitle("Bar chart with Error Bars")

fig <- ggplotly(p)

fig
```

# Stacked Bar Chart
<code>geom_bar</code> with stacked traces

```{r}
library(plotly)
library(dplyr)
library(reshape)

df <- structure(c(106487, 495681, 1597442, 2452577, 2065141, 2271925, 4735484, 3555352, 8056040, 4321887, 2463194, 347566, 621147, 1325727, 1123492, 800368, 761550, 1359737, 1073726, 36, 53, 141, 41538, 64759, 124160, 69942, 74862, 323543, 247236, 112059, 16595, 37028, 153249, 427642, 1588178, 2738157, 2795672, 2265696, 11951, 33424, 62469, 74720, 166607, 404044, 426967, 38972, 361888, 1143671, 1516716, 160037, 354804, 996944, 1716374, 1982735, 3615225, 4486806, 3037122, 17, 54, 55, 210, 312, 358, 857, 350, 7368, 8443, 6286, 1750, 7367, 14092, 28954, 80779, 176893, 354939, 446792, 33333, 69911, 53144, 29169, 18005, 11704, 13363, 18028, 46547, 14574, 8954, 2483, 14693, 25467, 25215, 41254, 46237, 98263, 185986), .Dim = c(19, 5), .Dimnames = list(c("1820-30", "1831-40", "1841-50", "1851-60", "1861-70", "1871-80", "1881-90", "1891-00", "1901-10", "1911-20", "1921-30", "1931-40", "1941-50", "1951-60", "1961-70", "1971-80", "1981-90", "1991-00", "2001-06"), c("Europe", "Asia", "Americas", "Africa", "Oceania")))
df.m <- melt(df)
df.m <- rename(df.m, Period = Var1, Region = Var2)

p <- ggplot(df.m, aes(x = Period, y = value/1e+06,fill = Region)) + ggtitle("Migration to the United States by Source Region (1820-2006), In Millions")
p <- p + geom_bar(stat = "identity", position = "stack")

fig <- ggplotly(p)

fig
```

# Ordered Bar Chart
ordering variable in <code>geom_bar</code>

```{r}
library(plotly)
library(plyr)

dane<-data.frame(x=1:10,y=seq(-5,4),g=rep(c('A','B'),each=5))
dane$x<-as.factor(dane$x)

p <- ggplot(data=dane,aes(x=x,y=y,fill=g)) +
    geom_bar(stat="identity")

fig <- ggplotly(p)

fig
```

# Precentages
using <code>geom_bar</code> to show percentages

```{r}
library(plotly)
set.seed(123)

df <- diamonds[sample(1:nrow(diamonds), size = 1000),]

p <- ggplot(df, aes(x = color)) + 
  geom_bar(aes(y = ..count../sum(..count..), fill = cut)) + 
  scale_fill_brewer(palette = "Set3") + 
  ylab("Percent") + 
  ggtitle("Show precentages in bar chart")
  
fig <- ggplotly(p)

fig
```

Set manual colors using <code>geom_bar</code> to manually specify colors.

```{r}
library(plotly)
library(RColorBrewer)
set.seed(123)

df <- diamonds[sample(1:nrow(diamonds), size = 1000),]

# Simply use fill = a vector of colors
p <- ggplot(df, aes(x = color)) +
  geom_bar(fill = brewer.pal(length(unique(df$color)), "Set3")) + 
  ylab("Count") + 
  ggtitle("Specify manual colors in a bar chart")
  
fig <- ggplotly(p)

fig
```

# Reordered Bar Chart
Re-ordering bars shown using <code>geom_bar</code>.

```{r}
library(plotly)

df <- data.frame(x = as.factor(LETTERS[1:5]), 
                 y = sample(10:20, size = 5))

# First change factor levels
df$x <- factor(df$x, levels = c("C", "B", "A", "D", "E"))

# Plot
p <- ggplot(df, aes(x, y, fill = x)) + 
  geom_bar(stat = "identity") +
  ggtitle("Bar Chart with changed factor levels")

fig <- ggplotly(p)

fig
```
