---
description: How to make Violin Plots in ggplot2 with Plotly.
name: Violin Plots
permalink: ggplot2/violin/
thumbnail: thumbnail/violin.jpg
layout: base
language: ggplot2
display_as: statistical
page_type: u-guide
order: 10
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

# Default violin plot

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin()

plotly::ggplotly(p)
```

# Flip plot orientation

```{r}
library(plotly)
library(ggplot2)

p <-   
 ggplot(mtcars, aes(mpg, factor(cyl))) +
  geom_violin()
  
plotly::ggplotly(p)
```

With `geom_violin()`, the y-axis must always be the continuous variable, and the x-axis the categorical variable. To create horizontal violin graphs, keep the x- and y-variables as is and add `coord_flip()`.

```{r}
library(plotly)
library(ggplot2)

district_density <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/district_density.csv", stringsAsFactors = FALSE)
district_density$cluster <- factor(district_density$cluster, levels=c("Pure urban", "Urban-suburban mix", "Dense suburban", "Sparse suburban", "Rural-suburban mix", "Pure rural"))
district_density$region <- factor(district_density$region, levels=c("West", "South", "Midwest", "Northeast"))

p <- ggplot(district_density,aes(x=cluster, y=dem_margin, fill=cluster)) +
  geom_violin(colour=NA) +
  geom_hline(yintercept=0, alpha=0.5) +
  labs(title = "Democratic performance in the 2018 House elections, by region and density",
       x = "Density Index\nfrom CityLab",
       y = "Margin of Victory/Defeat") +
  coord_flip()

ggplotly(p)
```

# Add data points with jitter

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin() + geom_jitter(height = 0, width = 0.1)

plotly::ggplotly(p)
```

# Scaling maximum width

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(scale = "count")

plotly::ggplotly(p)
```


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(scale = "width")

plotly::ggplotly(p)
```

# Disabling default trim

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(trim = FALSE)

plotly::ggplotly(p)
```



# Closer density fit


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(adjust = .5)

plotly::ggplotly(p)
```


# Adding fill

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(aes(fill = cyl))

plotly::ggplotly(p)
```

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(aes(fill = factor(cyl)))

plotly::ggplotly(p)
```




```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(aes(fill = factor(vs)))

plotly::ggplotly(p)
```



```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(aes(fill = factor(am)))

plotly::ggplotly(p)
```


# Changing border colour

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(fill = "grey80", colour = "#3366FF")

plotly::ggplotly(p)
```


# Enabling quartiles


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg))
p <- p + geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))

plotly::ggplotly(p)
```


# Example


```{r}
library(plotly)
library(ggplot2)

p <-   
 if (require("ggplot2movies")) {
  # Scale transformations occur before the density statistics are computed.
  # Coordinate transformations occur afterwards.  Observe the effect on the
  # number of outliers.
  m <- ggplot(movies, aes(y = votes, x = rating, group = cut_width(rating, 0.5)))
  m + geom_violin()
  m +
    geom_violin() +
    scale_y_log10()
  m +
    geom_violin() +
    coord_trans(y = "log10")
  m +
    geom_violin() +
    scale_y_log10() + coord_trans(y = "log10")

  # Violin plots with continuous x:
  # Use the group aesthetic to group observations in violins
  ggplot(movies, aes(year, budget)) +
    geom_violin()
  ggplot(movies, aes(year, budget)) +
    geom_violin(aes(group = cut_width(year, 10)), scale = "width")
}

plotly::ggplotly(p)
```

### Basic violin plot
A basic violin plot showing how Democratic vote share in the 2018 elections to the US House of Representatives varied by level of density. A horizontal bar is added, to divide candidates who lost from those who won.

Source: [Dave Wassermann and Ally Flinn](https://docs.google.com/spreadsheets/d/1WxDaxD5az6kdOjJncmGph37z0BPNhV1fNAH_g7IkpC0/htmlview?sle=true#gid=0) for the election results and CityLab for its [Congressional Density Index](https://github.com/theatlantic/citylab-data/tree/master/citylab-congress). Regional classifications are according to the Census Bureau.

```{r}
library(plotly)
library(ggplot2)

district_density <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/district_density.csv", stringsAsFactors = FALSE)
district_density$cluster <- factor(district_density$cluster, levels=c("Pure urban", "Urban-suburban mix", "Dense suburban", "Sparse suburban", "Rural-suburban mix", "Pure rural"))
district_density$region <- factor(district_density$region, levels=c("West", "South", "Midwest", "Northeast"))

p <- ggplot(district_density,aes(x=cluster, y=dem_margin, fill=cluster)) +
  geom_violin(colour=NA) +
  geom_hline(yintercept=0, alpha=0.5) +
  labs(title = "Democratic performance in the 2018 House elections, by region and density",
       x = "Density Index\nfrom CityLab",
       y = "Margin of Victory/Defeat")

ggplotly(p)
```


### Add facetting
Including facetting by region.

```{r}
library(plotly)
library(ggplot2)

district_density <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/district_density.csv", stringsAsFactors = FALSE)
district_density$cluster <- factor(district_density$cluster, levels=c("Pure urban", "Urban-suburban mix", "Dense suburban", "Sparse suburban", "Rural-suburban mix", "Pure rural"))
district_density$region <- factor(district_density$region, levels=c("West", "South", "Midwest", "Northeast"))

p <- ggplot(district_density,aes(x=cluster, y=dem_margin, fill=cluster)) +
  geom_violin(colour=NA) +
  geom_hline(yintercept=0, alpha=0.5) +
  facet_wrap(~region) +
  labs(title = "Democratic performance in the 2018 House elections, by region and density",
       x = "Density Index\nfrom CityLab",
       y = "Margin of Victory/Defeat") +
  coord_flip()

ggplotly(p)
```

### Customized Appearance
Add colour to the facet titles, centre-align the title, rotate the y-axis title, change the font, and get rid of the unnecessary legend. Note that `coord_flip()` flips the axes for the variables and the titles, but does not flip `theme()` elements.

```{r}
library(plotly)
library(ggplot2)

district_density <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/district_density.csv", stringsAsFactors = FALSE)
district_density$cluster <- factor(district_density$cluster, levels=c("Pure urban", "Urban-suburban mix", "Dense suburban", "Sparse suburban", "Rural-suburban mix", "Pure rural"))
district_density$region <- factor(district_density$region, levels=c("West", "South", "Midwest", "Northeast"))

p <- ggplot(district_density,aes(x=cluster, y=dem_margin, fill=cluster)) +
  geom_violin(colour=NA) +
  geom_hline(yintercept=0, alpha=0.5) +
  facet_wrap(~region) +
  labs(title = "Democratic performance in the 2018 House elections, by region and density",
       x = "Density Index\nfrom CityLab",
       y = "Margin of Victory/Defeat") +
  coord_flip() +
  theme(axis.title.y = element_text(angle = 0, vjust=0.5),
        plot.title = element_text(hjust = 0.5),
        strip.background = element_rect(fill="lightblue"),
        text = element_text(family = 'Fira Sans'),
        legend.position = "none")

ggplotly(p)
```

### Rotated Axis Text
Rotated the x-axis text 45 degrees, and used `facet_grid` to create a 4x1 facet (compared to `facet_wrap`, which defaults to 2x2).

```{r}
library(plotly)
library(ggplot2)

district_density <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/district_density.csv", stringsAsFactors = FALSE)
district_density$cluster <- factor(district_density$cluster, levels=c("Pure urban", "Urban-suburban mix", "Dense suburban", "Sparse suburban", "Rural-suburban mix", "Pure rural"))
district_density$region <- factor(district_density$region, levels=c("West", "South", "Midwest", "Northeast"))

p <- ggplot(district_density,aes(x=cluster, y=dem_margin, fill=cluster)) +
  geom_violin(colour=NA) +
  geom_hline(yintercept=0, alpha=0.5) +
  facet_grid(.~region) +
  labs(title = "Democratic performance in the 2018 House elections, by region and density",
       x = "Density Index\nfrom CityLab",
       y = "Margin of Victory/Defeat") +
  theme(axis.text.x = element_text(angle = -45),
        plot.title = element_text(hjust = 0.5),
        strip.background = element_rect(fill="lightblue"),
        text = element_text(family = 'Fira Sans'),
        legend.position = "none")

ggplotly(p)
```
