---
name: Contour Plots
permalink: ggplot2/contour-plots/
description: How to make Contour Plots in ggplot2 with Plotly.
layout: base
thumnail_github: contour-plots.png
language: ggplot2
page_type: u-guide
display_as: scientific
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic `geom_contour` plot
`geom_contour` produces a similar output to `geom_density_2d`, except it uses a third variable for the values rather than frequency. The volcano dataset comes pre-loaded on R.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

p <- ggplot(df, aes(Var1, Var2, z= value)) +
  geom_contour() +
  scale_fill_distiller(palette = "Spectral", direction = -1)

ggplotly(p)
```

### Coloured Plot
[See here](https://ggplot2.tidyverse.org/reference/scale_brewer.html) for a list of colour palettes that come with the brewer (discrete) and distiller (continuous) packages.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

p <- ggplot(df, aes(Var1, Var2, z= value, colour=stat(level))) +
  geom_contour() +
  scale_colour_distiller(palette = "YlGn", direction = 1)

ggplotly(p)
```

### Filled Plot
It's possible to colour in each of the layers, by changing `geom_contour` to `stat_contour` as below. As the edges of the graph indicate, filled contour plots only work when each layer is an enclosed shape rather than an open line; a geom more suited to this functionality would be `geom_tile` or `geom_raster`.

```{r}
library(plotly)
library(reshape2)
df <- melt(volcano)

p <- ggplot(df, aes(Var1, Var2, z= value)) +
  stat_contour(geom="polygon",aes(fill=stat(level))) +
  scale_fill_distiller(palette = "Spectral", direction = -1)

ggplotly(p)
```

