---
name: Simple Features in Maps
permalink: ggplot2/maps-sf/
description: How to use Simple Features in Maps with Plotly.
layout: base
thumnail_github: maps-sf.png
language: ggplot2
page_type: example_index
display_as: maps
order: 9
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Introduction

In order to complete the examples below, you'll require installing additional packages (`install.packages("packageName")`):
- [sf](https://github.com/r-spatial/sf)

The examples below use the library [simple features](https://r-spatial.github.io/sf/) to read in the shape files before plotting the features with Plotly.

### Basic sf 

```{r}
library(plotly)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

fig <- ggplotly(
  ggplot(nc) +
  geom_sf(aes(fill = AREA))
) 

fig
```


### Using Native Plotly

Alternatively, you can use [`plot_ly`, `plot_geo`, or `plot_mapbox`](https://plotly.com/r/maps-sf/).

### Reference

See [https://plotly.com/r/reference/](https://plotly.com/r/reference/) for more information and chart attribute options! If you would like to read more on visualizing geo-spatial data with sf and ggplotly click [here](https://blog.cpsievert.me/2018/01/30/learning-improving-ggplotly-geom-sf/).
