---
description: How to make Scatterplot Matrix using `ggpairs()` in ggplot2 with Plotly.
name: Scatterplot Matrix
permalink: ggplot2/splom/
thumbnail: thumbnail/splom_image.jpg
layout: base
language: ggplot2
display_as: statistical
page_type: u-guide
order: 6
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Default splot plot

The `ggpairs()` function of the `GGally` package allows to build a great scatterplot matrix.

Scatterplots of each pair of numeric variable are drawn on the left part of the figure. Pearson correlation is displayed on the right. Variable distribution is available on the diagonal.

```{r}
library(plotly)
library(GGally)
 
data <- data.frame( var1 = 1:100 + rnorm(100,sd=20), v2 = 1:100 + rnorm(100,sd=27), v3 = rep(1, 100) + rnorm(100, sd = 1)) 
data$v4 = data$var1 ** 2 
data$v5 = -(data$var1 ** 2) 
 
p <- ggpairs(data, title="correlogram with ggpairs()") 

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Visualize correlation

he `ggcorr()` function allows to visualize the correlation of each pair of variable as a square. Note that the `method` argument allows to pick the correlation type you desire.

```{r}
library(plotly)
library(GGally)
 
data <- data.frame( var1 = 1:100 + rnorm(100,sd=20), v2 = 1:100 + rnorm(100,sd=27), v3 = rep(1, 100) + rnorm(100, sd = 1)) 
data$v4 = data$var1 ** 2 
data$v5 = -(data$var1 ** 2) 
 
p <- ggcorr(data, method = c("everything", "pearson")) 

p
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Split by group
It is possible to use `ggplot2` aesthetics on the chart, for instance to color each category.

```{r}
library(plotly)
library(GGally)
 
data(flea)

p <- ggpairs(flea, columns = 2:4, ggplot2::aes(colour=species)) 

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Change plot types

Change the type of plot used on each part of the `correlogram`. This is done with the `upper` and `lower` argument.

```{r}
library(plotly)
library(GGally)
 
data(tips, package = "reshape")

p <- ggpairs(
      tips[, c(1, 3, 4, 2)],
      upper = list(continuous = "density", combo = "box_no_facet"),
      lower = list(continuous = "points", combo = "dot_no_facet")
)

p
```

<!--------------------- EXAMPLE BREAK ------------------------->