---
description: How to make Strip Charts in ggplot2 with Plotly.
name: Strip Charts
permalink: ggplot2/strip-charts/
thumbnail: thumbnail/figure-labels.png
layout: base
language: ggplot2
display_as: statistical
page_type: u-guide
order: 14
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Basic stripchart


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

ggplot(ToothGrowth, aes(x=dose, y=len)) + 
  geom_jitter()

p<-ggplot(ToothGrowth, aes(x=dose, y=len)) + 
  geom_jitter(position=position_jitter(0.2))
p <- p + coord_flip()

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Change point size


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
        geom_jitter(position=position_jitter(0.2), cex=1.2)

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Change shape


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
        geom_jitter(position=position_jitter(0.2), shape=17)

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Add summary statistics


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
  geom_jitter(position=position_jitter(0.2), cex=1.2)
p <- p + stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

ggplotly(p)
```

To add standard deviation use `mean_sdl` function which computes the mean plus or minus a constant times the standard deviation.

```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
    geom_jitter(position=position_jitter(0.2))
p + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red")

ggplotly(p)
```

You can change the representation of the statistics by changing `geom`, for example setting it to `crossbar`.


<!--------------------- EXAMPLE BREAK ------------------------->

## Add box plot

To add a box plot you can use `geom_boxplot()`. Likewise, you can add a notched boxplot with `geom_boxplot(notch = TRUE)` and a violin plot with `geom_violin(trim = FALSE)`.  

```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + 
        geom_boxplot()+
        geom_jitter(position=position_jitter(0.2))

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Colour data by groups


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, color=dose)) +
  geom_jitter(position=position_jitter(0.2))

ggplotly(p)
```


```{r}
library(plotly)
library(ggplot2)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p <- ggplot(ToothGrowth, aes(x=dose, y=len, color=dose, shape=dose)) + 
  geom_jitter(position=position_jitter(0.2))+
  labs(title="Plot of length  by dose",x="Dose (mg)", y = "Length")
p + theme_classic()

fig <- ggplotly(p)


ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->
