---
description: How to do kNN Classification in ggplot2 with Plotly.
name: kNN Classification
permalink: ggplot2/knn-classification/
thumbnail: thumbnail/knn-classification.png
layout: base
language: ggplot2
display_as: ai_ml
page_type: u-guide
order: 2
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

`ggfortify` supports `stats::kmeans` class. You must explicitly pass original data to `autoplot` function via `data` keyword. Because kmeans object doesn't store original data. The result will be automatically colorized by categorized cluster.

## XXXXXXXXXXXXX


```{r}
library(plotly)
library(ggfortify)

set.seed(1)
p <- autoplot(kmeans(USArrests, 3), data = USArrests)

ggplotly(p)
```


```{r}
library(plotly)
library(ggfortify)

p <- autoplot(kmeans(USArrests, 3), data = USArrests, label = TRUE, label.size = 3)

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Using `cluster` package

{`ggfortify`} supports `cluster::clara`, `cluster::fanny`, `cluster::pam` as well as `cluster::silhouette` classes. Because these instances should contains original data in its property, there is no need to pass original data explicitly.

```{r}
library(plotly)
library(ggfortify)
library(cluster)

p <- autoplot(clara(iris[-5], 3))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Plotting cluster converx

Specifying `frame = TRUE` in autoplot for `stats::kmeans` and `cluster::*` draws convex for each cluster.


```{r}
library(plotly)
library(ggfortify)
library(cluster)

p <-  autoplot(fanny(iris[-5], 3), frame = TRUE)

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Plotting cluster probability ellipse

If you want probability ellipse, specify whatever supported in `ggplot2::stat_ellipse`'s type keyword via `frame.type` option.

```{r}
library(plotly)
library(ggfortify)
library(cluster)

p <- autoplot(pam(iris[-5], 3), frame = TRUE, frame.type = 'norm')

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Silhouette plot


```{r}
library(plotly)
library(ggfortify)
library(cluster)

p <- autoplot(silhouette(pam(iris[-5], 3L)))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->