---
description: How to work with Axes in ggplot2 with Plotly.
name: Axes
permalink: ggplot2/axes/
thumbnail: thumbnail/axes.png
layout: base
language: ggplot2
display_as: file_settings
page_type: u-guide
order: 13
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Default plot


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(name = "Displacement")

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Changing axis limits


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(limits = c(0, 600))

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Changing axis ticks

 The `breaks` argument is used to specify where the ticks appear. It takes a numeric vector equal to the length of the number of ticks.

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(breaks = c(150, 300, 450))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Changing tick labels

You can change the tick labels using the `labels` argument. In the below example, we use words instead of numbers. When adding labels, we need to ensure that the length of the `breaks` and `labels` are same.

```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(breaks = c(150, 300, 450),
  labels = c('One Hundred Fifty', 'Three Hundred', 'Four Hundred Fifity'))

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Changing position of labels


```{r}
library(ggplot2)

p <- ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(position = 'top')

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Y-axis example

Everything above using `scale_x_continuous()` can be used in the same way for y axis with `scale_y_continuous()`.


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) + geom_point(aes(disp, mpg)) +
  scale_y_continuous(name = "Miles Per Gallon", limits = c(0, 45),
                     breaks = c(0, 15, 30, 45), position = 'right',
                     labels = c('zero', 'fifteen', 'thirty', 'fourtyfive'))

ggplotly(p)

```


<!--------------------- EXAMPLE BREAK ------------------------->

## Discrete axis labels


```{r}
library(plotly)
library(ggplot2)

p <- ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(labels = c("4" = "Four", "6" = "Six", "8" = "Eight"))

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Remove Axis Labels

```{r}
library(plotly)
library(ggplot2)

df <- diamonds[sample(1:nrow(diamonds), size = 1000),]

p <- ggplot(df, aes(carat, price, color = cut)) +
  geom_point() +
  theme(axis.text = element_blank())

 ggplotly(p)
```

## Vertical Text Orientation

```{r}
library(plotly)
library(ggplot2)

lab <- paste("Vertical Label", c(1, 2, 3, 4, 5))

ds <- data.frame(x = sample(lab, size = 1000, replace = T),
                 y = sample(LETTERS[1:5], size = 1000, replace = T))

p <- ggplot(ds, aes(x = x, fill = y)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Vertical Axis Labels")

ggplotly(p)
```

## Angled Text Orientation

```{r}
library(plotly)
library(ggplot2)

lab <- paste("Angle Label", c(1, 2, 3, 4, 5))

ds <- data.frame(x = sample(lab, size = 1000, replace = T),
                 y = sample(LETTERS[1:5], size = 1000, replace = T))

p <- ggplot(ds, aes(x = x, fill = y)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle("Angle Axis Labels")

ggplotly(p)
```