---
description: How to make Parallel Coordinates Plot in ggplot2 with Plotly.
name: Parallel Coordinates Plot
permalink: ggplot2/parallel-coordinates-plot/
thumnail_github: parallel-coordinates-plot.png
layout: base
language: ggplot2
display_as: scientific
page_type: u-guide
order: 9
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Default parallel coordinates plot

The input dataset must be a data frame with several numeric variables, each being used as a vertical axis on the chart. Columns number of these variables are specified in the `columns` argument of the function.

For this example, a categoric variable is used to color lines, as specified in the `groupColumn` variable.
```{r}
library(plotly)
library(GGally)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5
    )

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Custom colour, theme and style

This is the same chart as the previous one, except for the following customizations:

* color palette is improved thanks to the `viridis` package
* title is added with `title`, and customized in `theme`
* dots are added with `showPoints`
* a bit of transparency is applied to lines with `alphaLines`
* `theme_ipsum()` is used for the general appearance

```{r}
library(plotly)
library(GGally)
library(viridis)
library(hrbrthemes)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for the Iris Data",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Scaling

Scaling transforms the raw data to a new scale that is common with other variables. It is a crucial step to compare variables that do not have the same unit, but can also help otherwise as shown in the example below.

The `ggally` package offers a `scale` argument. Four possible options are applied on the same dataset below:

* `globalminmax` → No scaling
* `uniminmax` → Standardize to Min = 0 and Max = 1
* `std` → Normalize univariately (substract mean & divide by sd)
* `center` → Standardize and center variables
```{r}
library(plotly)
library(GGally)
library(viridis)
library(hrbrthemes)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="globalminmax",
    showPoints = TRUE, 
    title = "No scaling",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")

ggplotly(p)
```

Standardize to Min = 0 and Max = 1:
```{r}
library(plotly)
library(GGally)
library(viridis)
library(hrbrthemes)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="uniminmax",
    showPoints = TRUE, 
    title = "Standardize to Min = 0 and Max = 1",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")

ggplotly(p)
```

Normalize univariately (substract mean & divide by sd):
```{r}
library(plotly)
library(GGally)
library(viridis)
library(hrbrthemes)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="std",
    showPoints = TRUE, 
    title = "Normalize univariately (substract mean & divide by sd)",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")

ggplotly(p)
```

Standardize and center variables:

```{r}
library(plotly)
library(GGally)
library(viridis)
library(hrbrthemes)

data <- iris

p <- ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="center",
    showPoints = TRUE, 
    title = "Standardize and center variables",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Highlight a group


```{r}
library(plotly)
library(GGally)
data <- iris

p <- data %>%
  arrange(desc(Species)) %>%
  ggparcoord(
    columns = 1:4, groupColumn = 5, order = "anyClass",
    showPoints = TRUE, 
    title = "Original",
    alphaLines = 1
    ) + 
  scale_color_manual(values=c( "#69b3a2", "#E8E8E8", "#E8E8E8") ) +
  theme_ipsum()+
  theme(
    legend.position="Default",
    plot.title = element_text(size=10)
  ) +
  xlab("")

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->