---
description: How to make Filled Area Plots in ggplot2 with Plotly.
name: Filled Area
permalink: ggplot2/filled-area-plots/
thumbnail: thumbnail/area.jpg
layout: base
language: ggplot2
display_as: basic
page_type: u-guide
order: 7
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

## Basic stacked area plot

The data frame used as input to build a stacked area chart requires 3 columns:
* `x`: numeric variable used for the X axis, often it is a time.
* `y`: numeric variable used for the Y axis. What are we looking at?
* `group`: one shape will be done per group.

The chart is built using the `geom_area()` function.

```{r}
library(plotly)
library(ggplot2)
library(dplyr)

time <- as.numeric(rep(seq(1,7),each=7))
value <- runif(49, 10, 100)
group <- rep(LETTERS[1:7],times=7)
data <- data.frame(time, value, group)

p <- ggplot(data, aes(x=time, y=value, fill=group)) + 
    geom_area()

ggplotly(p)
```


<!--------------------- EXAMPLE BREAK ------------------------->

## Control stacking order

The gallery offers a post dedicated to reordering with ggplot2. This step can be tricky but the code below shows how to:

* give a specific order with the `factor()` function.
* order alphabetically using `sort()`
* order following values at a specific data

```{r}
library(plotly)
library(ggplot2)
library(dplyr)

time <- as.numeric(rep(seq(1,7),each=7))
value <- runif(49, 10, 100)
group <- rep(LETTERS[1:7],times=7)
data <- data.frame(time, value, group)

data$group <- factor(data$group , levels=c("B", "A", "D", "E", "G", "F", "C") )

p <- ggplot(data, aes(x=time, y=value, fill=group)) + 
    geom_area()

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->

## Proportional stack
In a proportional stacked area graph, the sum of each year is always equal to hundred and value of each group is represented through percentages.

```{r}
library(plotly)
library(ggplot2)
library(dplyr)
data <- data  %>%
  group_by(time, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))

p <- ggplot(data, aes(x=time, y=percentage, fill=group)) + 
    geom_area(alpha=0.6 , size=1, colour="black")

ggplotly(p)
```

<!--------------------- EXAMPLE BREAK ------------------------->
