---
description: How to use shapes in R. Two examples on highlighting regions by adding
  shapes to your R charts.
display_as: file_settings
language: r
layout: base
name: Shapes
order: 13
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/shapes/
thumbnail: thumbnail/shape.jpg
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
### Lines

```{r}
library(plotly)

s <- seq.int(0, 15)
fig <- plot_ly(x = ~s, y = ~sin(s), mode = "lines")

# initiate a line shape object
line <- list(
  type = "line",
  line = list(color = "pink"),
  xref = "x",
  yref = "y"
)

lines <- list()
for (i in c(0, 3, 5, 7, 9, 13)) {
  line[["x0"]] <- i
  line[["x1"]] <- i + 2
  line[c("y0", "y1")] <- sin(i + 1)
  lines <- c(lines, list(line))
}

fig <- layout(fig, title = 'Highlighting with Lines', shapes = lines)

fig
```

### Rectangles
```{r}
library(plotly)
fig <- plot_ly(economics, x = ~date, y = ~uempmed, name = "unemployment")

# add shapes to the layout
fig <- layout(fig, title = 'Highlighting with Rectangles',
             shapes = list(
               list(type = "rect",
                    fillcolor = "blue", line = list(color = "blue"), opacity = 0.3,
                    x0 = "1980-01-01", x1 = "1985-01-01", xref = "x",
                    y0 = 4, y1 = 12.5, yref = "y"),
               list(type = "rect",
                 fillcolor = "blue", line = list(color = "blue"), opacity = 0.2,
                 x0 = "2000-01-01", x1 = "2005-01-01", xref = "x",
                 y0 = 4, y1 = 12.5, yref = "y")))

fig
```

### Circles

Circles are centered around  ((`x0`+`x1`)/2, (`y0`+`y1`)/2))

```{r}
library(plotly)
df <- diamonds[sample(nrow(diamonds), 1000), ]
fig <- plot_ly(df, x = ~carat, y = ~price, text = ~paste("Clarity: ", clarity),
             mode = "markers", color = ~carat, size = ~carat)
fig <- layout(fig, title = 'Highlighting Regions with Circles',
            shapes = list(
              list(type = 'circle',
                   xref = 'x', x0 = .2, x1 = .7,
                   yref = 'y', y0 = 20, y1 = 3000,
                   fillcolor = 'rgb(50, 20, 90)', line = list(color = 'rgb(50, 20, 90)'),
                   opacity = 0.2),
              list(type = 'circle',
                   xref = 'x', x0 = .75, x1 = 1.5,
                   yref = 'y', y0 = 2500, y1 = 7500,
                   fillcolor = 'rgb(30, 100, 120)', line = list(color = 'rgb(30, 100, 120)'),
                   opacity = 0.2),
               list(type = 'circle',
                   xref = 'x', x0 = 1.6, x1 = 2.5,
                   yref = 'y', y0 = 12500, y1 = 18500,
                   fillcolor = 'rgb(90, 200, 75)', line = list(color = 'rgb(90, 200, 75)'),
                   opacity = 0.2)))

fig
```

### Drawing Shapes on Cartesian Plots

You can create layout shapes programatically, but you can also draw shapes manually by setting the `dragmode` to one of the shape-drawing modes: `drawline`,`drawopenpath`, `drawclosedpath`, `drawcircle`, or `drawrect`. If you need to switch between different shape-drawing or other dragmodes (panning, selecting, etc.), modebar buttons can be added in the `config` of your figure to select the dragmode. 

If you switch to a different dragmode such as pan or zoom, you will need to select the drawing tool in the modebar to go back to shape drawing.

This shape-drawing feature is particularly interesting for annotating graphs, in particular [image traces](https://plotly.com/r/displaying-images/#drawing-shapes-on-images). 

Once you have drawn shapes, you can select and modify an existing shape by clicking on its boundary (note the arrow pointer). Its fillcolor turns to pink to highlight the activated shape and then you can

- drag and resize it for lines, rectangles and circles/ellipses
- drag and move individual vertices for closed paths
- move individual vertices for open paths.

An activated shape is deleted by cliking on the `eraseshape` button.

Drawing or modifying a shape triggers a `relayout` event, which [can be captured by a callback inside a Dash For R application](https://dash.plotly.com/interactive-graphing).

```{r}
library(plotly)

df <- diamonds[sample(nrow(diamonds), 1000), ]

fig <- plot_ly(df, x = ~carat, y = ~price, text = ~paste("Clarity: ", clarity),
             mode = "markers", color = ~carat, size = ~carat)

fig <- layout(fig, dragmode="drawrect", xaxis = list(title = 'Click and drag inside the figure to draw a rectangle or select another shape in the modebar'))

fig <- fig %>%
  config(modeBarButtonsToAdd = list("drawine", "drawopenpath", "drawclosedpath", "drawcircle", "drawrect", "eraseshape" ) )

fig

```

### Style of user-drawn shapes

The layout `newshape` attribute controls the visual appearance of new shapes drawn by the user. `newshape` attributes have the same names as layout shapes.

Note on shape opacity: having a new shape's opacity > 0.5 makes it possible to activate a shape by clicking inside the shape (for opacity <= 0.5 you have to click on the border of the shape), but you cannot start a new shape within an existing shape (which is possible for an opacity <= 0.5).

```{r}
library(plotly)

df <- diamonds[sample(nrow(diamonds), 1000), ]

fig <- plot_ly(df, x = ~carat, y = ~price, text = ~paste("Clarity: ", clarity),
             mode = "markers", color = ~carat, size = ~carat)

fig <- layout(fig, dragmode="drawrect", newshape=list(fillcolor="yellow", opacity=0.5), xaxis = list(title = 'Click and drag inside the figure to draw a rectangle or select another shape in the modebar'))

fig <- fig %>%
  config(modeBarButtonsToAdd = list("drawine", "drawopenpath", "drawclosedpath", "drawcircle", "drawrect", "eraseshape" ) )

fig

```
### Reference
Check out our <b>[reference page](https://plotly.com/r/reference/#layout-shapes)</b> for more information on using shapes!