---
description: How to modify the legend in R graphs. Nine examples of how to move, color,
  and hide the legend.
display_as: file_settings
language: r
layout: base
name: Legends
order: 4
output:
  html_document:
    keep_md: true
page_type: example_index
permalink: r/legend/
thumbnail: thumbnail/legends.gif
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Legend Names

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3')
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')

fig
```

### Hiding the Legend

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines')
fig <- fig %>% add_trace(y = ~Tree2)
fig <- fig %>% add_trace(y = ~Tree3)
fig <- fig %>% add_trace(y = ~Tree4)
fig <- fig %>% add_trace(y = ~Tree5)
fig <- fig %>% layout(showlegend = FALSE)

fig
```

### Hiding Legend Entries

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3', showlegend = FALSE)
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')

fig
```

### Positioning the Legend Inside the Plot

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3')
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')
fig <- fig %>% layout(legend = list(x = 0.1, y = 0.9))

fig
```

### Positioning the Legend Outside the Plot

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3')
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')
fig <- fig %>% layout(legend = list(x = 100, y = 0.5))

fig
```

### Changing the Legend Orientation

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3')
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')
fig <- fig %>% layout(legend = list(orientation = 'h'))

fig
```

### Styling the Legend

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

l <- list(
  font = list(
    family = "sans-serif",
    size = 12,
    color = "#000"),
  bgcolor = "#E2E2E2",
  bordercolor = "#FFFFFF",
  borderwidth = 2)


fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines', name = 'Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, name = 'Tree 2')
fig <- fig %>% add_trace(y = ~Tree3, name = 'Tree 3')
fig <- fig %>% add_trace(y = ~Tree4, name = 'Tree 4')
fig <- fig %>% add_trace(y = ~Tree5, name = 'Tree 5')
fig <- fig %>% layout(legend = l)

fig
```

### Grouped Legend

Plotly legends are interactive. Click on the legend entries to hide and show traces. The *legendgroup* key groups legend entries so that clicking on one legend entry will hide or show all of the traces in that group.

```{r}
library(plotly)
library(tidyr)
library(plyr)

data <- spread(Orange, Tree, circumference)
data <- rename(data, c("1" = "Tree1", "2" = "Tree2", "3" = "Tree3", "4" = "Tree4", "5" = "Tree5"))

fig <- plot_ly(data, x = ~age, y = ~Tree1, type = 'scatter', mode = 'lines',
        legendgroup = 'group1', name = 'Zone 1 - Tree 1')
fig <- fig %>% add_trace(y = ~Tree2, legendgroup = 'group2', name = 'Zone 2 - Tree 1')
fig <- fig %>% add_trace(y = ~Tree3, legendgroup = 'group1', name = 'Zone 1 - Tree 2')
fig <- fig %>% add_trace(y = ~Tree4, legendgroup = 'group2', name = 'Zone 2 - Tree 2')
fig <- fig %>% add_trace(y = ~Tree5, legendgroup = 'group1', name = 'Zone 1 - Tree 3')

fig
```

### Subplot Grouped Legend

```{r}
library(plotly)

df <- data.frame(x = c("a","b","c"), y = c(2,3,2), y2 = c(4,2,4))

fig1 <- df
fig1 <- fig1 %>% plot_ly(
    type = 'bar', 
    x = ~x, 
    y = ~y, 
    color = ~x, 
    legendgroup = ~x
  )
fig1 <- fig1 %>% layout(
    xaxis = list(
      showgrid = F
    ),
    yaxis = list(
      showgrid = F
    )
  )

fig2 <- df
fig2 <- fig2 %>% plot_ly(
    type = 'bar', 
    x = ~x, 
    y = ~y2, 
    color =  ~x, 
    legendgroup = ~x, 
    showlegend = F
  )
fig2 <- fig2 %>% layout(
    xaxis = list(
      showgrid = F
    ),
    yaxis = list(
      showgrid = F
    )
  )

fig <- subplot(fig1, fig2, nrows = 2, shareX = T)

fig
```

Reference

See [https://plot.ly/r/reference/#layout-legend](https://plot.ly/r/reference/#layout-legend) for more information and chart attribute options!