---
description: How to make a histogram in R.
display_as: statistical
language: r
layout: base
name: Histograms
order: 3
output:
  html_document:
    keep_md: true
page_type: example_index
permalink: r/histograms/
thumbnail: thumbnail/histogram.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Basic Histogram

```{r}
library(plotly)
fig <- plot_ly(x = ~rnorm(50), type = "histogram")

fig
```

#### Normalized Histogram

```{r}
library(plotly)
fig <- plot_ly(x = ~rnorm(50),
             type = "histogram",
             histnorm = "probability")

fig
```

#### Specify Binning Function

```{r}
library(plotly)

x = c("Apples","Apples","Apples","Oranges", "Bananas")
y = c("5","10","3","10","5")

fig <- plot_ly(y=y, x=x, histfunc='sum', type = "histogram")
fig <- fig %>% layout(yaxis=list(type='linear'))

fig
```

#### Horizontal Histogram

```{r}
library(plotly)
fig <- plot_ly(y = ~rnorm(50), type = "histogram")

fig
```

#### Overlaid Histograms

```{r}
fig <- plot_ly(alpha = 0.6)
fig <- fig %>% add_histogram(x = ~rnorm(500))
fig <- fig %>% add_histogram(x = ~rnorm(500) + 1)
fig <- fig %>% layout(barmode = "overlay")

fig
```

#### Stacked Histograms

```{r}
fig <- plot_ly(alpha = 0.6)
fig <- fig %>% add_histogram(x = ~rnorm(500))
fig <- fig %>% add_histogram(x = ~rnorm(500) + 1)
fig <- fig %>% layout(barmode = "stack")

fig
```

#### Cumulative Histogram

```{r}
library(plotly)
fig <- plot_ly(x = ~rnorm(50),
             type = "histogram",
             cumulative = list(enabled=TRUE))

fig
```
### Share bins between histograms

In this example both histograms have a compatible bin settings using [bingroup](https://plotly.com/r/reference/#histogram-bingroup) attribute. 
```{r}
library(plotly)

fig <- plot_ly(
  type='histogram',
  x=~rnorm(100, 5),
  bingroup=1)

fig <- fig %>% add_trace(
  type='histogram',
  x=~rnorm(20, 5),
  bingroup=1)

fig <- fig %>% layout(
  barmode="overlay",
  bargap=0.1)

fig
```
Note that traces on the same subplot, and with the same `barmode` ("stack", "relative", "group") are forced into the same `bingroup`, however traces with `barmode = "overlay"` and on different axes (of the same axis type) can have compatible bin settings. Histogram and [histogram2d](https://plotly.com/r/2D-Histogram/) trace can share the same `bingroup`.

```{r}
library(plotly)

fig <- plot_ly(
  type='histogram',
  x=~rnorm(100, 5))

fig <- fig %>% add_trace(
  type='histogram',
  x=~rnorm(20, 5))

fig <- fig %>% layout(
  barmode="stack",
  bargap=0.1)

fig
```
### Reference

See [https://plotly.com/r/reference/#histogram](https://plotly.com/r/reference/#histogram) for more information and chart attribute options!
