---
name: Time Series
permalink: r/time-series/
description: How to plot date and time in R. An example of a time series plot with the POSIXct and Sys.Date classes.
layout: base
thumbnail: thumbnail/time-series.jpg
language: r
page_type: example_index
display_as: financial
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Dates

```{r}
library(plotly)
today <- Sys.Date()
tm <- seq(0, 600, by = 10)
x <- today - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "days from today"))

fig
```

### Hiding Weekends and Holidays
The `rangebreaks` attribute available on x- and y-axes of type date can be used to hide certain time-periods. In the example below, we show two plots: one in default mode to show gaps in the data, and one where we hide weekends and holidays to show an uninterrupted trading history. Note the smaller gaps between the grid lines for December 21 and January 4, where holidays were removed. Check out the reference for more options: https://plotly.com/r/reference/#layout-xaxis-rangebreaks


```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')
fig <- plot_ly(
  type = "scatter",
  x = as.Date(df$Date, format= "%Y-%m-%d"), 
  y = df$AAPL.High,
  name = 'AAPL High',
  mode = "markers",
)
fig <- fig %>%
  layout(
    title = "Time Series with Custom Date-Time Format",
    xaxis = list(
      type = "date",
      range=c('2015-12-01', '2016-01-15')
    )
  )
fig
```

```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')
fig <- plot_ly(
  type = "scatter",
  x = as.Date(df$Date, format= "%Y-%m-%d"), 
  y = df$AAPL.High,
  name = 'AAPL High',
  mode = "markers",
)
fig <- fig %>%
  layout(
    title = "Time Series with Custom Date-Time Format",
    xaxis = list(
      type = "date",
      range=c('2015-12-01', '2016-01-15'),
      rangebreaks = list( 
        list(bounds=c("sat", "mon")),
        list(values=c("2015-12-25", "2016-01-01"))
      )
    )
  )
fig
```
### POSIXlt date time class with timezone

```{r}
library(plotly)
now_lt <- as.POSIXlt(Sys.time(), tz = "GMT")
tm <- seq(0, 600, by = 10)
x <- now_lt - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "seconds from now in GMT"))

fig
```

### POSIXct date time class without timezone

```{r}
library(plotly)
now_ct <- as.POSIXct(Sys.time())
tm <- seq(0, 600, by = 10)
x <- now_ct - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "seconds from now in", Sys.timezone()))

fig
```
