---
name: Getting Started with Plotly
permalink: r/getting-started/
description: How to get started making charts with Plotly's R graphing library.
page_type: example_index
layout: base
language: r
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

# Getting Started with Plotly for R

[`plotly`](https://github.com/ropensci/plotly) is an R package for creating interactive web-based graphs via the open source JavaScript graphing library [`plotly.js`](http://plot.ly/javascript).

As of version 2.0 (November 17, 2015), graphs created with the `plotly` R package are rendered *locally* through the [`htmlwidgets`](http://www.htmlwidgets.org/) framework.

## Installation

### Download from CRAN

Use the `install.package()` function to install the [`plotly R package`](https://cran.r-project.org/web/packages/plotly/index.html) from [CRAN](https://cran.r-project.org/!

```r
install.packages("plotly")
```

### Download from GitHub

Alternatively, you can install the latest development version of `plotly` from GitHub via the [`devtools`](https://cran.r-project.org/web/packages/devtools/index.html) R package:

```r
devtools::install_github("ropensci/plotly")
```

### Note For `RStudio` Users

`RStudio` users should ensure that they are using the latest [RStudio release](https://www.rstudio.com/products/rstudio/download/) in order to ensure compatibility with the `htmlwidgets` R package.

## Rendering Charts

By default, the `plotly` R package runs locally in your web browser or in the `RStudio` viewer.

```{r}
library(plotly)
fig <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
fig
```

Simply printing the plot object will render the chart locally in your web browser or in the `RStudio` viewer.

Graphs created with the `plotly` R package are interactive! 

Click on legend entries to hide/show traces, click-and-drag on the chart to zoom, double-click to autoscale, shift-and-drag to pan.

## Next Steps

Once you have installed the `plotly` R package, learn more about the [fundamentals of making charts](https://plot.ly/r/plotly-fundamentals/) and start making [basic charts](https://plot.ly/r/basic-charts/).