---
name: Subplots
permalink: r/subplots/
description: How to graph subplots and small multiples in R with Plotly.
layout: base
thumbnail: thumbnail/subplots.jpg
language: r
page_type: example_index
display_as: multiple_axes
order: 4
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic Subplots

The subplot() function provides a flexible interface for merging plotly objects into a single object (i.e., view).

```{r}
library(plotly)
fig1 <- plot_ly(economics, x = ~date, y = ~unemploy)
fig1 <- fig1 %>% add_lines(name = ~"unemploy")
fig2 <- plot_ly(economics, x = ~date, y = ~uempmed)
fig2 <- fig2 %>% add_lines(name = ~"uempmed")
fig <- subplot(fig1, fig2)

fig
```

### Scaled Subplots

Although subplot() accepts an arbitrary number of plot objects, passing a list of plots can save typing and redundant code when dealing with a large number of plots. For example, you can share the x-axis by utlising shareX, set axis ID, and and specify the number of of rows with nrows.

```{r}

fig <- economics
fig <- fig %>% tidyr::gather(variable, value, -date)
fig <- fig %>% transform(id = as.integer(factor(variable)))
fig <- fig %>% plot_ly(x = ~date, y = ~value, color = ~variable, colors = "Dark2",
          yaxis = ~paste0("y", id))
fig <- fig %>% add_lines()
fig <- fig %>% subplot(nrows = 5, shareX = TRUE)

fig
```

### Recursive Subplots

The subplot() function returns a plotly object so it can be modified like any other plotly object. This effectively means that subplots work recursively (i.e., you can have subplots within subplots).

```{r}
plotList <- function(nplots) {
  lapply(seq_len(nplots), function(x) plot_ly())
}
s1 <- subplot(plotList(6), nrows = 2, shareX = TRUE, shareY = TRUE)
s2 <- subplot(plotList(2), shareY = TRUE)
fig <- subplot(s1, s2, plot_ly(), nrows = 3, margin = 0.04, heights = c(0.6, 0.3, 0.1))

fig
```

For more information on subplots check [the plotly book](https://cpsievert.github.io/plotly_book/merging-plotly-objects.html)
