---
name: Scatter and Line Plots
permalink: r/line-and-scatter/
description: How to create line and scatter plots in R. Examples of basic and advanced scatter plots, time series line plots, colored charts, and density plots.
layout: base
thumbnail: thumbnail/line-and-scatter.jpg
language: r
page_type: example_index
display_as: basic
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
### Basic Scatter Plot

```{r}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length)

fig
```

### Styled Scatter Plot

```{r}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length,
               marker = list(size = 10,
                             color = 'rgba(255, 182, 193, .9)',
                             line = list(color = 'rgba(152, 0, 0, .8)',
                                         width = 2)))
fig <- fig %>% layout(title = 'Styled Scatter',
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))

fig
```

### Plotting Markers and Lines

```{r}
library(plotly)

trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x)
fig <- fig %>% add_trace(y = ~trace_0, name = 'trace 0',mode = 'lines')
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers')
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')

fig
```

It is also possible to pass the first trace in the plot_ly function. In such cases, the type of graph has to be specified, as shown below:

```{r}
library(plotly)

trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x, y = ~trace_0, name = 'trace 0', type = 'scatter', mode = 'lines')
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers')
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')

fig
```

See more examples of line charts [here](https://plot.ly/r/line-charts/).

### Qualitative Colorscales

```{r}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species)

fig
```

### ColorBrewer Palette Names

```{r}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, colors = "Set1")

fig
```

### Custom Color Scales

The `colors` argument also accepts a character vector of any valid R color code(s).

```{r}
library(plotly)

pal <- c("red", "blue", "green")

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, colors = pal)

fig
```

To ensure a particular data value gets mapped to particular color, provide a character vector of color codes, and match the names attribute accordingly.

```{r}
library(plotly)

pal <- c("red", "blue", "green")
pal <- setNames(pal, c("virginica", "setosa", "versicolor"))

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, colors = pal)

fig
```

### Mapping Data to Symbols

```{r}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, type = 'scatter',
  mode = 'markers', symbol = ~Species, symbols = c('circle','x','o'),
  color = I('black'), marker = list(size = 10))

fig
```

### Adding Color and Size Mapping

```{r}
library(plotly)

d <- diamonds[sample(nrow(diamonds), 1000), ]

fig <- plot_ly(
  d, x = ~carat, y = ~price,
  color = ~carat, size = ~carat
)

fig
```

### Data Labels on Hover

```{r}
library(plotly)

d <- diamonds[sample(nrow(diamonds), 1000), ]

fig <- plot_ly(
  d, x = ~carat, y = ~price,
  # Hover text:
  text = ~paste("Price: ", price, '$<br>Cut:', cut),
  color = ~carat, size = ~carat
)

fig
```

#Reference

See [https://plot.ly/r/reference/#scatter](https://plot.ly/r/reference/#scatter) for more information and chart attribute options!
