---
description: How to make an interactive box plot in R. Examples of box plots in R
  that are grouped, colored, and display the underlying data distribution.
display_as: statistical
language: r
layout: base
name: Box Plots
order: 2
output:
  html_document:
    keep_md: true
page_type: example_index
permalink: r/box-plots/
thumbnail: thumbnail/box.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic Boxplot

```{r}
library(plotly)
fig <- plot_ly(y = ~rnorm(50), type = "box")
fig <- fig %>% add_trace(y = ~rnorm(50, 1))

fig
```

### Choosing The Algorithm For Computing Quartiles

By default, quartiles for box plots are computed using the **linear** method (for more about linear interpolation, see #10 listed on http://www.amstat.org/publications/jse/v14n3/langford.html and https://en.wikipedia.org/wiki/Quartile for more details).

However, you can also choose to use an exclusive or an inclusive algorithm to compute quartiles.

The **exclusive** algorithm uses the median to divide the ordered dataset into two halves. If the sample is odd, it does not include the median in either half. Q1 is then the median of the lower half and Q3 is the median of the upper half.

The **inclusive** algorithm also uses the median to divide the ordered dataset into two halves, but if the sample is odd, it includes the median in both halves. Q1 is then the median of the lower half and Q3 the median of the upper half.

```{r}
library(plotly)
fig <- plot_ly(y = list(1,2,3,4,5), type = "box", quartilemethod="exclusive") # or "inclusive", or "linear" by default

fig
```

### Modifying The Algorithm For Computing Quartiles

For an explanation of how each algorithm works, see [Choosing The Algorithm For Computing Quartiles](https://plotly.com/r/box-plots/#choosing-the-algorithm-for-computing-quartiles)

```{r}
library(plotly)
fig <- plot_ly(y = list(1,2,3,4,5), type = "box", quartilemethod="linear", name="Linear Quartile Mode")
fig <- fig %>% add_trace(y = list(1,2,3,4,5), quartilemethod="inclusive", name="Inclusive Quartile Mode")
fig <- fig %>% add_trace(y = list(1,2,3,4,5), quartilemethod="exclusive", name="Exclusive Quartile Mode")
fig <- fig %>% layout(title = "Modifying The Algorithm For Computing Quartiles")

fig
```

### Box Plot With Precomputed Quartiles

You can specify precomputed quartile attributes rather than using a built-in quartile computation algorithm.

This could be useful if you have already pre-computed those values or if you need to use a different algorithm than the ones provided.

```{r}
library(plotly)
fig <- plot_ly(y = list(1,2,3,4,5,6,7,8,9), type = "box", q1=list(1, 2, 3), median=list(4, 5, 6),
                  q3=list(7, 8, 9 ), lowerfence=list(-1, 0, 1),
                  upperfence=list(5, 6, 7), mean=list(2.2, 2.8, 3.2 ),
                  sd=list(0.2, 0.4, 0.6), notchspan=list(0.2, 0.4, 0.6))

fig
```

### Horizontal Boxplot

```{r}
library(plotly)
fig <- plot_ly(x = ~rnorm(50), type = "box")
fig <- fig %>% add_trace(x = ~rnorm(50, 1))

fig
```

### Adding Jittered Points

```{r}
fig <- plot_ly(y = ~rnorm(50), type = "box", boxpoints = "all", jitter = 0.3,
        pointpos = -1.8)

fig
```


### Several Box Plots

```{r}
fig <- plot_ly(ggplot2::diamonds, y = ~price, color = ~cut, type = "box")

fig
```

### Grouped Box Plots

```{r}
fig <- plot_ly(ggplot2::diamonds, x = ~cut, y = ~price, color = ~clarity, type = "box")
fig <- fig %>% layout(boxmode = "group")

fig
```

### Styling Outliers

```{r}
library(plotly)

y1 <- c(0.75, 5.25, 5.5, 6, 6.2, 6.6, 6.80, 7.0, 7.2, 7.5, 7.5, 7.75, 8.15,
       8.15, 8.65, 8.93, 9.2, 9.5, 10, 10.25, 11.5, 12, 16, 20.90, 22.3, 23.25)
y2 <- c(0.75, 5.25, 5.5, 6, 6.2, 6.6, 6.80, 7.0, 7.2, 7.5, 7.5, 7.75, 8.15,
        8.15, 8.65, 8.93, 9.2, 9.5, 10, 10.25, 11.5, 12, 16, 20.90, 22.3, 23.25)
y3 <- c(0.75, 5.25, 5.5, 6, 6.2, 6.6, 6.80, 7.0, 7.2, 7.5, 7.5, 7.75, 8.15,
        8.15, 8.65, 8.93, 9.2, 9.5, 10, 10.25, 11.5, 12, 16, 20.90, 22.3, 23.25)
y4 <- c(0.75, 5.25, 5.5, 6, 6.2, 6.6, 6.80, 7.0, 7.2, 7.5, 7.5, 7.75, 8.15,
        8.15, 8.65, 8.93, 9.2, 9.5, 10, 10.25, 11.5, 12, 16, 20.90, 22.3, 23.25)

fig <- plot_ly(type = 'box')
fig <- fig %>% add_boxplot(y = y1, jitter = 0.3, pointpos = -1.8, boxpoints = 'all',
              marker = list(color = 'rgb(7,40,89)'),
              line = list(color = 'rgb(7,40,89)'),
              name = "All Points")
fig <- fig %>% add_boxplot(y = y2, name = "Only Whiskers", boxpoints = FALSE,
              marker = list(color = 'rgb(9,56,125)'),
              line = list(color = 'rgb(9,56,125)'))
fig <- fig %>% add_boxplot(y = y3, name = "Suspected Outlier", boxpoints = 'suspectedoutliers',
              marker = list(color = 'rgb(8,81,156)',
                            outliercolor = 'rgba(219, 64, 82, 0.6)',
                            line = list(outliercolor = 'rgba(219, 64, 82, 1.0)',
                                        outlierwidth = 2)),
              line = list(color = 'rgb(8,81,156)'))
fig <- fig %>% add_boxplot(y = y4, name = "Whiskers and Outliers", boxpoints = 'outliers',
              marker = list(color = 'rgb(107,174,214)'),
              line = list(color = 'rgb(107,174,214)'))
fig <- fig %>% layout(title = "Box Plot Styling Outliers")

fig
```

### Reference

See [https://plotly.com/r/reference/#box](https://plotly.com/r/reference/#box) for more information and chart attribute options!