---
name: Time Series
permalink: r/time-series/
description: How to plot date and time in R. An example of a time series plot with the POSIXct and Sys.Date classes.
layout: base
thumbnail: thumbnail/time-series.jpg
language: r
page_type: example_index
display_as: financial
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Dates

```{r}
library(plotly)
today <- Sys.Date()
tm <- seq(0, 600, by = 10)
x <- today - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "days from today"))

fig
```

### Hiding Weekends and Holidays
The `rangebreaks` attribute available on x- and y-axes of type date can be used to hide certain time-periods. In the example below, we create a plot of the last ten days, excluding Saturdays, Sundays, yesterday, and the day before yesterday. Check out the reference for more options: https://plotly.com/r/reference/#layout-xaxis-rangebreaks

```{r}
library(plotly)
today <- Sys.Date()
yesterday = today - 1
day_before_yesterday = yesterday - 1
tm <- seq(0, 10, by = 1)
x <- today - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'markers', text = paste(tm, "days from today"))
fig <- fig %>% layout(
  xaxis = list(rangebreaks = list( 
    list(bounds=c("sat", "sun")),
    list(values=c(yesterday,day_before_yesterday))
    ))
)
fig
```
### POSIXlt date time class with timezone

```{r}
library(plotly)
now_lt <- as.POSIXlt(Sys.time(), tz = "GMT")
tm <- seq(0, 600, by = 10)
x <- now_lt - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "seconds from now in GMT"))

fig
```

### POSIXct date time class without timezone

```{r}
library(plotly)
now_ct <- as.POSIXct(Sys.time())
tm <- seq(0, 600, by = 10)
x <- now_ct - tm
y <- rnorm(length(x))
fig <- plot_ly(x = ~x, y = ~y, mode = 'lines', text = paste(tm, "seconds from now in", Sys.timezone()))

fig
```
