---
name: Pie Charts
permalink: r/pie-charts/
description: How to make pie charts in R using plotly.
layout: base
thumbnail: thumbnail/pie-chart.jpg
language: r
page_type: example_index
display_as: basic
order: 4
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Basic Pie Chart

```{r}
library(plotly)

USPersonalExpenditure <- data.frame("Categorie"=rownames(USPersonalExpenditure), USPersonalExpenditure)
data <- USPersonalExpenditure[,c('Categorie', 'X1960')]

fig <- plot_ly(data, labels = ~Categorie, values = ~X1960, type = 'pie')
fig <- fig %>% layout(title = 'United States Personal Expenditures by Categories in 1960',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

### Styled Pie Chart

```{r}
library(plotly)

USPersonalExpenditure <- data.frame("Categorie" = rownames(USPersonalExpenditure), USPersonalExpenditure)
data <- USPersonalExpenditure[, c('Categorie', 'X1960')]

colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)')

fig <- plot_ly(data, labels = ~Categorie, values = ~X1960, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        text = ~paste('$', X1960, ' billions'),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
                      #The 'pull' attribute can also be used to create space between the sectors
        showlegend = FALSE)
fig <- fig %>% layout(title = 'United States Personal Expenditures by Categories in 1960',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

### Subplots
In order to create pie chart subplots, you need to use the [domain](https://plotly.com/javascript/reference/#pie-domain) attribute. It is important to note that the `X` array set the horizontal position whilst the `Y` array sets the vertical. For example, `x=[0,0.5], y=[0, 0.5]` would mean the bottom left position of the plot.
```{r}
library(plotly)
library(dplyr)

fig <- plot_ly()
fig <- fig %>% add_pie(data = count(diamonds, cut), labels = ~cut, values = ~n,
          name = "Cut", domain = list(x = c(0, 0.4), y = c(0.4, 1)))
fig <- fig %>% add_pie(data = count(diamonds, color), labels = ~color, values = ~n,
          name = "Color", domain = list(x = c(0.6, 1), y = c(0.4, 1)))
fig <- fig %>% add_pie(data = count(diamonds, clarity), labels = ~clarity, values = ~n,
          name = "Clarity", domain = list(x = c(0.25, 0.75), y = c(0, 0.6)))
fig <- fig %>% layout(title = "Pie Charts with Subplots", showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

### Subplots Using Grid
This example uses a plotly [grid](https://plotly.com/javascript/reference/#layout-grid) attribute for the suplots. Reference the row and column destination using the [domain](https://plotly.com/javascript/reference/#pie-domain) attribute.
```{r}
library(plotly)
library(dplyr)

fig <- plot_ly()
fig <- fig %>% add_pie(data = count(diamonds, cut), labels = ~cut, values = ~n,
                         name = "Cut", domain = list(row = 0, column = 0))
fig <- fig %>% add_pie(data = count(diamonds, color), labels = ~color, values = ~n,
                       name = "Color", domain = list(row = 0, column = 1))
fig <- fig %>% add_pie(data = count(diamonds, clarity), labels = ~clarity, values = ~n,
                       name = "Clarity", domain = list(row = 1, column = 0))
fig <- fig %>% add_pie(data = count(diamonds, cut), labels = ~cut, values = ~n,
                       name = "Clarity", domain = list(row = 1, column = 1))
fig <- fig %>% layout(title = "Pie Charts with Subplots", showlegend = F,
                      grid=list(rows=2, columns=2),
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

See more examples of subplots [here](https://plotly.com/r/subplots/).

#### Controlling text orientation inside sunburst sectors

The `insidetextorientation` attribute controls the orientation of text inside sectors. With "auto" the texts may automatically be rotated to fit with the maximum size inside the slice. Using "horizontal" (resp. "radial", "tangential") forces text to be horizontal (resp. radial or tangential). Note that `plotly` may reduce the font size in order to fit the text with the requested orientation.

```{r}
library(plotly)

labels = c('Oxygen','Hydrogen','Carbon_Dioxide','Nitrogen')
values = c(4500, 2500, 1053, 500)

fig <- plot_ly(type='pie', labels=labels, values=values, 
               textinfo='label+percent',
               insidetextorientation='radial')
fig
```
### Donut Chart

```{r}
library(plotly)
library(dplyr)

# Get Manufacturer
mtcars$manuf <- sapply(strsplit(rownames(mtcars), " "), "[[", 1)

df <- mtcars
df <- df %>% group_by(manuf)
df <- df %>% summarize(count = n())
fig <- df %>% plot_ly(labels = ~manuf, values = ~count)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "Donut charts using Plotly",  showlegend = F,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

#Reference

See [https://plotly.com/r/reference/#pie](https://plotly.com/r/reference/#pie) for more information and chart attribute options!
