---
description: How to add text labels and annotations to plots in R.
display_as: file_settings
language: r
layout: base
name: Text and Annotations
order: 9
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/text-and-annotations/
thumbnail: thumbnail/text-and-annotations.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Text Mode

```{r}
library(plotly)

Primates <- c('Potar monkey', 'Gorilla', 'Human', 'Rhesus monkey', 'Chimp')
Bodywt <- c(10.0, 207.0, 62.0, 6.8, 52.2)
Brainwt <- c(115, 406, 1320, 179, 440)
data <- data.frame(Primates, Bodywt, Brainwt)

fig <- plot_ly(data, x = ~Bodywt, y = ~Brainwt, type = 'scatter',
        mode = 'text', text = ~Primates, textposition = 'middle right',
        textfont = list(color = '#000000', size = 16))
fig <- fig %>% layout(title = 'Primates Brain and Body Weight',
         xaxis = list(title = 'Body Weight (kg)',
                      zeroline = TRUE,
                      range = c(0, 250)),
         yaxis = list(title = 'Brain Weight (g)',
                      range = c(0,1400)))

fig
```

See more options on the textposition argument [here](https://plotly.com/r/reference/#scatter-textposition).

### Styling Text

```{r}
library(plotly)

data <- mtcars[which(mtcars$am == 1 & mtcars$gear == 4),]

t <- list(
  family = "sans serif",
  size = 14,
  color = toRGB("grey50"))

fig <- plot_ly(data, x = ~wt, y = ~mpg, text = rownames(data))
fig <- fig %>% add_markers()
fig <- fig %>% add_text(textfont = t, textposition = "top right")
fig <- fig %>% layout(xaxis = list(range = c(1.6, 3.2)),
         showlegend = FALSE)

fig
```

### Controlling text fontsize with uniformtext

For the [pie](/r/pie-charts), [bar](/r/bar-charts), [sunburst](/r/sunburst-charts) and [treemap](/r/treemap-charts) traces, it is possible to force all the text labels to have the same size thanks to the `uniformtext` layout parameter. The `minsize` attribute sets the font size, and the `mode` attribute sets what happens for labels which cannot fit with the desired fontsize: either `hide` them or `show` them with overflow.

```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/gapminderDataFiveYear.csv", stringsAsFactors = FALSE)
df <- df[which(df$year==2007 & df$continent=='Europe' & df$pop > 2.e6),]

fig <- plot_ly(df, type='bar', x = ~country, y = ~pop, text = ~lifeExp, name="",
               hovertemplate = paste('%{x}', '<br>lifeExp: %{text:.2s}<br>'),
               texttemplate = '%{y:.2s}', textposition = 'outside')

fig <- fig %>% layout(uniformtext=list(minsize=8, mode='hide'))
fig
```

```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/gapminderDataFiveYear.csv", stringsAsFactors = FALSE)
df <- df[which(df$year==2007 & df$continent=='Asia'),]

fig <- plot_ly(df, type='pie', labels = ~country, values = ~pop, textposition = 'inside')
fig <- fig %>% layout(uniformtext=list(minsize=12, mode='hide'))
fig
````
### Adding Informations to Default Hover Text

```{r}
library(plotly)

fig <- plot_ly(iris, x = ~Petal.Length, y = ~Petal.Width, type = 'scatter', mode = 'markers',
        text = ~paste('Species: ', Species))

fig
```

### Custom Hover Text

```{r}
library(plotly)

fig <- plot_ly(iris, x = ~Petal.Length, y = ~Petal.Width, type = 'scatter', mode = 'markers',
        hoverinfo = 'text',
        text = ~paste('</br> Species: ', Species,
                      '</br> Petal Length: ', Petal.Length,
                      '</br> Petal Width: ', Petal.Width))

fig
```

### Single Annotation

```{r}
library(plotly)

m <- mtcars[which.max(mtcars$mpg), ]

a <- list(
  x = m$wt,
  y = m$mpg,
  text = rownames(m),
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 7,
  ax = 20,
  ay = -40
)

fig <- plot_ly(mtcars, x = ~wt, y = ~mpg)
fig <- fig %>% add_markers()
fig <- fig %>% layout(annotations = a)

fig
```

### Multiple Annotations

```{r}
library(plotly)

data <- mtcars[which(mtcars$am == 1 & mtcars$gear == 4),]

fig <- plot_ly(data, x = ~wt, y = ~mpg, type = 'scatter', mode = 'markers',
        marker = list(size = 10))
fig <- fig %>% add_annotations(x = data$wt,
                  y = data$mpg,
                  text = rownames(data),
                  xref = "x",
                  yref = "y",
                  showarrow = TRUE,
                  arrowhead = 4,
                  arrowsize = .5,
                  ax = 20,
                  ay = -40)

fig
```

### Subplot Annotations

```{r}
library(plotly)

m <- economics[which.max(economics$unemploy), ]
n <- economics[which.max(economics$uempmed), ]

# annotations
a <- list(
  x = m$date,
  y = m$unemploy,
  text = "annotation a",
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 7,
  ax = 20,
  ay = -40
)

b <- list(
  x = n$date,
  y = n$uempmed,
  text = "annotation b",
  xref = "x2",
  yref = "y2",
  showarrow = TRUE,
  arrowhead = 7,
  ax = 20,
  ay = -40
)

# figure labels
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f ")
x <- list(
  title = "x Axis",
  titlefont = f)
y <- list(
  title = "y Axis",
  titlefont = f)

fig1 <- plot_ly(economics, x = ~date, y = ~unemploy)
fig1 <- fig1 %>% add_lines(name = ~"unemploy")
fig1 <- fig1 %>% layout(annotations = a, xaxis = x, yaxis = y)
fig2 <- plot_ly(economics, x = ~date, y = ~uempmed)
fig2 <- fig2 %>% add_lines(name = ~"uempmed")
fig2 <- fig2 %>% layout(annotations = b, xaxis = x, yaxis = y)
fig <- subplot(fig1, fig2, titleX = TRUE, titleY = TRUE)
fig2 <- fig2 %>% layout(showlegend = FALSE)

fig
```

### 3D Annotations

```{r}
library(plotly)

fig <- plot_ly()
fig <- fig %>% add_trace(
    x = c("2017-01-01", "2017-02-10", "2017-03-20"), 
    y = c("A", "B", "C"), 
    z = c(1, 1000, 100000), 
    name = "z", 
    type = "scatter3d"
  )
fig <- fig %>% layout(
    scene = list(
      aspectratio = list(
        x = 1,
        y = 1,
        z = 1
      ),
      camera = list(
        center = list(
          x = 0,
          y = 0,
          z = 0
        ),
        eye = list(
          x = 1.96903462608,
          y = -1.09022831971,
          z = 0.405345349304
        ),
        up = list(
          x = 0,
          y = 0,
          z = 1
        )
      ),
      dragmode = "turntable",
      xaxis = list(
        title = "",
        type = "date"
      ),
      yaxis = list(
        title = "",
        type = "category"
      ),
      zaxis = list(
        title = "",
        type = "log"
      ),
      annotations = list(list(
        showarrow = F,
        x = "2017-01-01",
        y = "A",
        z = 0,
        text = "Point 1",
        xanchor = "left",
        xshift = 10,
        opacity = 0.7
      ), list(
        x = "2017-02-10",
        y = "B",
        z = 4,
        text = "Point 2",
        textangle = 0,
        ax = 0,
        ay = -75,
        font = list(
          color = "black",
          size = 12
        ),
        arrowcolor = "black",
        arrowsize = 3,
        arrowwidth = 1,
        arrowhead = 1
      ), list(
        x = "2017-03-20",
        y = "C",
        z = 5,
        ax = 50,
        ay = 0,
        text = "Point 3",
        arrowhead = 1,
        xanchor = "left",
        yanchor = "bottom"
      )
    )),
    xaxis = list(title = "x"),
    yaxis = list(title = "y")
  )

fig
```

### Styling Annotations

```{r}
library(plotly)

data <- mtcars[which(mtcars$am == 1 & mtcars$gear == 4),]

fig <- plot_ly(data, x = ~wt, y = ~mpg, type = 'scatter', mode = 'markers',
        marker = list(size = 10))
fig <- fig %>% add_annotations(x = data$wt,
                  y = data$mpg,
                  text = rownames(data),
                  xref = "x",
                  yref = "y",
                  showarrow = TRUE,
                  arrowhead = 4,
                  arrowsize = .5,
                  ax = 20,
                  ay = -40,
                  # Styling annotations' text:
                  font = list(color = '#264E86',
                              family = 'sans serif',
                              size = 14))

fig
```

### Set Annotation Text Anchors

```{r}
library(plotly)

fig <- plot_ly()
fig <- fig %>% add_markers(
    x = 1,
    y = 1,
    showlegend = F
  )
fig <- fig %>% add_markers(
    x = 1,
    y = 2,
    showlegend = F
  )
fig <- fig %>% add_markers(
    x = 1,
    y = 3,
    showlegend = F
  )
fig <- fig %>% add_annotations(
    x=1,
    y=1,
    xref = "x",
    yref = "y",
    text = "Right Anchor",
    xanchor = 'right',
    showarrow = F
  )
fig <- fig %>% add_annotations(
    x=1,
    y=2,
    xref = "x",
    yref = "y",
    text = "Center Anchor",
    xanchor = 'center',
    showarrow = F
  )
fig <- fig %>% add_annotations(
    x=1,
    y=3,
    xref = "x",
    yref = "y",
    text = "Left Anchor",
    xanchor = 'left',
    showarrow = F
  )

fig
```
### Customize Displayed Text with a Text Template

To show an arbitrary text in your chart you can use [texttemplate](https://plotly.com/r/reference/#pie-texttemplate), which is a template string used for rendering the information, and will override [textinfo](https://plotly.com/r/reference/#treemap-textinfo).
This template string can include `variables` in %{variable} format, `numbers` in [d3-format's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Formatting.md#d3_forma), and `date` in [d3-time-fomrat's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Time-Formatting.md#format).
`texttemplate` customizes the text that appears on your plot vs. [hovertemplate](https://plotly.com/r/reference/#pie-hovertemplate) that customizes the tooltip text.

```{r}
library(plotly)

fig <- plot_ly(
  type='pie',
  values=c(40000000, 20000000, 30000000, 10000000),
  labels=c("Wages", "Operating expenses", "Cost of sales", "Insurance"),
  texttemplate="%{label}: %{value:$,s} <br>(%{percent})",
  textposition="inside")
fig
```

### Customize Text Template

The following example uses [textfont](https://plotly.com/r/reference/#scatterternary-textfont) to customize the added text.

```{r}
library(plotly)

fig <- plot_ly(
  type='scatterternary',
  a = c(3, 2, 5),
  b = c(2, 5, 2),
  c = c(5, 2, 2),
  mode = "markers+text",
  text = c("A", "B", "C"),
  texttemplate = "%{text}<br>(%{a:.2f}, %{b:.2f}, %{c:.2f})",
  textposition = "bottom center",
  textfont = list(family= "Times", size= c(18, 21, 20), color= c("IndianRed", "MediumPurple", "DarkOrange"))
)
fig
```
### Set Date in Text Template

The following example shows how to show date by setting [axis.type](https://plotly.com/r/reference/#layout-yaxis-type) in [funnel charts](https://plotly.com/r/funnel-charts/).
As you can see [textinfo](https://plotly.com/r/reference/#funnel-textinfo) and [texttemplate](https://plotly.com/r/reference/#funnel-texttemplate) have the same functionality when you want to determine 'just' the trace information on the graph.

```{r}
fig <- plot_ly()

fig <- fig %>% add_trace(
  type='funnel',
  name = 'Montreal',
  orientation = "h",
  y = c("2018-01-01", "2018-07-01", "2019-01-01", "2020-01-01"),
  x = c(100, 60, 40, 20),
  textposition = "inside",
  texttemplate = "%{y| %a. %_d %b %Y}")

fig <- fig %>% add_trace(
  type='funnel',
  name = 'Vancouver',
  orientation = "h",
  y = c("2018-01-01", "2018-07-01", "2019-01-01", "2020-01-01"),
  x = c(90, 70, 50, 10),
  textposition = "inside",
  textinfo = "label")

fig <- fig %>% layout(yaxis = list(type= 'date'))
fig
```
### Set Annotation Coordinate References

```{r}
library(plotly)

fig <- plot_ly()
fig <- fig %>% add_markers(
    x = 0.5,
    y = 1,
    showlegend = F
  )
fig <- fig %>% add_annotations(
    x= 0.5,
    y= 1,
    xref = "paper",
    yref = "paper",
    text = "<b>paper reference = [0.5, 1]</b>",
    showarrow = F
  )
fig <- fig %>% add_annotations(
    x= 0.5,
    y= 1,
    xref = "x",
    yref = "y",
    text = "x + y reference = [0.5, 1]",
    showarrow = T,
    ax = 20,
    ay = -40
  )
fig <- fig %>% layout(
    xaxis = list(zeroline = F),
    yaxis = list(zeroline = F)
  )

fig
```

#Reference

See [https://plotly.com/r/reference/#layout-annotations](https://plotly.com/r/reference/#layout-annotations) for more information and chart attribute options!