---
name: Getting Started with Plotly
permalink: r/getting-started/
description: Get started with Plotly's R graphing library.
page_type: example_index
layout: base
language: r
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

# Getting Started with Plotly for R

`plotly` is an R package for creating interactive web-based graphs via the open source JavaScript graphing library [plotly.js](http://plot.ly/javascript).
As of version 2.0 (November 17, 2015), Plotly graphs are rendered *locally* through the [htmlwidgets](http://www.htmlwidgets.org/) framework.


#### Installation

[`plotly`](https://cran.r-project.org/web/packages/plotly/index.html) is now on [CRAN](https://cran.r-project.org/!

```r
install.packages("plotly")
```

Or install the latest development version (on GitHub) via the devtools R package:

```r
devtools::install_github("ropensci/plotly")
```

<br>
RStudio users should download the latest [RStudio release](https://www.rstudio.com/products/rstudio/download/) for compatibility with htmlwidgets.

#### Initialization for Offline Plotting

By default, the `plotly` R package runs locally in your web browser or in the R Studio viewer.

```{r}
library(plotly)
p <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
p
```

Simply printing the plot object will render the chart locally in your web browser or in the R Studio viewer.

Graphs created with the `plotly` R package are interactive. Click on legend entries to toggle traces, click-and-drag on the chart to zoom, double-click to autoscale, shift-and-drag to pan.