---
description: How to download Chart Studio users' public graphs and data into R.
display_as: chart_studio
language: r
layout: base
name: Dowloading Chart Studio Graphs
order: 6
output:
  html_document:
    keep_md: true
permalink: r/download-chart-studio-graphs/
redirect_from:
- r/get-requests/
thumbnail: thumbnail/get-requests.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### New to Plotly?

Plotly's R library is free and open source!<br>
[Get started](https://plot.ly/r/getting-started/) by downloading the client and [reading the primer](https://plot.ly/r/getting-started/).<br>
You can set up Plotly to work in [online](https://plot.ly/r/getting-started/#hosting-graphs-in-your-online-plotly-account) or [offline](https://plot.ly/r/offline/) mode.<br>
We also have a quick-reference [cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf) (new!) to help you get started!

### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.
```{r}
library(plotly)
packageVersion('plotly')
```

### Download Chart Studio Graphs Into R

Download Chart Studio figures directly into R with the `api_download_plot()` function. This takes the `plot_id` of the Chart Studio plot and the `username` of the plot's creator as arguments.

For example, to download [https://plot.ly/~cpsievert/559](https://plot.ly/~cpsievert/559) into R, call:

```{r}
library(plotly)
fig <- api_download_plot("559", "cpsievert")
fig
```

### Edit The Downloaded Graph
Once the figure is downloaded, you can edit it. This will create a new figure unless you specify the same filename as the figure that you downloaded.

```{r}
p <- layout(fig, title = paste("Modified on ", Sys.time()))
p
```

### Adding a Trace to a Subplot Figure

```{r}
fig <- api_download_plot("6343", "chelsea_lyn")

p <- add_lines(fig, x = c(1, 2), y = c(1, 2), xaxis = "x2", yaxis = "y2")
p
```

### Reference

See `help("api")`
