---
description: How to make a choropleth map in R. A choropleth map shades geographic
  regions by value.
display_as: maps
language: r
layout: base
name: Choropleth Maps
order: 1
output:
  html_document:
    keep_md: true
page_type: example_index
permalink: r/choropleth-maps/
thumbnail: thumbnail/choropleth.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

A [Choropleth Map](https://en.wikipedia.org/wiki/Choropleth_map) is a map composed of colored polygons. It is used to represent spatial variations of a quantity. This page documents how to build **outline** choropleth maps, but you can also build [choropleth **tile maps** using our Mapbox trace types](/r/mapbox-county-choropleth).

### Base Map Configuration

Plotly figures made with `plot_ly` have a `layout.geo` object which can be used to control the appearance of the base map onto which data is plotted.

### Introduction: main parameters for choropleth outline maps

Making choropleth maps requires two main types of input:

1. Geometry information:
   1. This can either be a supplied GeoJSON file where each feature has either an `id` field or some identifying value in `properties`; or
   2. one of the built-in geometries within `plot_ly`: US states and world countries (see below)
2. A list of values indexed by feature identifier.

The GeoJSON data is passed to the `geojson` argument, and the data is passed into the `z` argument of choropleth traces.

**Note** the `geojson` attribute can also be the URL to a GeoJSON file, which can speed up map rendering in certain cases.

### GeoJSON with `feature.id`

Here we load a GeoJSON file containing the geometry information for US counties, where `feature.id` is a [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
library(plotly)
library(rjson)

data <- fromJSON(file="https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json")
data$features[[1]]
```

#### Data indexed by `id`

Here we load unemployment data by county, also indexed by [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
df = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv", header = T, colClasses = c("fips"="character"))
head(df)
```

### Choropleth Map Using GeoJSON
```{r}
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
df <- read.csv(url2, colClasses=c(fips="character"))
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_ly()
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=counties,
    locations=df$fips,
    z=df$unemp,
    colorscale="Viridis",
    zmin=0,
    zmax=12,
    marker=list(line=list(
      width=0)
    )
  )
fig <- fig %>% colorbar(title = "Unemployment Rate (%)")
fig <- fig %>% layout(
    title = "2016 US Unemployment by County"
)

fig <- fig %>% layout(
    geo = g
  )

fig
```
**Note** In this example we set `layout.geo.scope` to usa to automatically configure the map to display USA-centric data in an appropriate projection.

### Indexing by GeoJSON Properties

If the GeoJSON you are using either does not have an `id` field or you wish you use one of the keys in the `properties` field, you may use the `featureidkey` parameter to specify where to match the values of locations.

In the following GeoJSON object/data-file pairing, the values of `properties.district` match the values of the `district` column:
```{r}
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/election.geojson'
geojson <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/election.csv"
df <- read.csv(url2)
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig <- plot_ly() 
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=geojson,
    locations=df$district,
    z=df$Bergeron,
    colorscale="Viridis",
    featureidkey="properties.district"
  )
fig <- fig %>% layout(
    geo = g
  )
fig <- fig %>% colorbar(title = "Bergeron Votes")
fig <- fig %>% layout(
    title = "2013 Montreal Election"
)
fig


```
### Using Built-in Country and State Geometries
Plotly comes with two built-in geometries which do not require an external GeoJSON file:

1. USA States
2. Countries as defined in the Natural Earth dataset.

**Note and disclaimer:** cultural (as opposed to physical) features are by definition subject to change, debate and dispute. Plotly includes data from Natural Earth "as-is" and defers to the [Natural Earth policy regarding disputed borders](https://www.naturalearthdata.com/downloads/50m-cultural-vectors/50m-admin-0-countries-2/) which read:

> Natural Earth Vector draws boundaries of countries according to defacto status. We show who actually controls the situation on the ground.

To use the built-in countries geometry, provide `locations` as [three-letter ISO country codes](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3).
```{r}
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv")
df

fig <- plot_ly(df, type='choropleth', locations=df$CODE, z=df$GDP..BILLIONS., text=df$COUNTRY, colorscale="Blues")

fig
```

To use the USA States geometry, set `locationmode='USA-states'` and provide `locations` as two-letter state abbreviations:
```{r}
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Millions USD")
fig <- fig %>% layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)'
  )

fig
```

### Customize choropleth chart
```{r}
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Millions USD")
fig <- fig %>% layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)',
    geo = g
  )

fig
```

**Note** In this example we set `layout.geo.scope` to `usa` to automatically configure the map to display USA-centric data in an appropriate projection.

### World Choropleth Map

```{r}
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
fig <- fig %>% add_trace(
    z = ~GDP..BILLIONS., color = ~GDP..BILLIONS., colors = 'Blues',
    text = ~COUNTRY, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'GDP Billions US$', tickprefix = '$')
fig <- fig %>% layout(
    title = '2014 Global GDP<br>Source:<a href="https://www.cia.gov/library/publications/the-world-factbook/fields/2195.html">CIA World Factbook</a>',
    geo = g
  )

fig
```

### Choropleth Inset Map

```{r}
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_ebola.csv')
# restrict from June to September
df <- subset(df, Month %in% 6:9)
# ordered factor variable with month abbreviations
df$abbrev <- ordered(month.abb[df$Month], levels = month.abb[6:9])
# September totals
df9 <- subset(df, Month == 9)

# common plot options
g <- list(
  scope = 'africa',
  showframe = F,
  showland = T,
  landcolor = toRGB("grey90")
)

g1 <- c(
  g,
  resolution = 50,
  showcoastlines = T,
  countrycolor = toRGB("white"),
  coastlinecolor = toRGB("white"),
  projection = list(type = 'Mercator'),
  list(lonaxis = list(range = c(-15, -5))),
  list(lataxis = list(range = c(0, 12))),
  list(domain = list(x = c(0, 1), y = c(0, 1)))
)

g2 <- c(
  g,
  showcountries = F,
  bgcolor = toRGB("white", alpha = 0),
  list(domain = list(x = c(0, .6), y = c(0, .6)))
)
 
fig <- df %>% plot_geo(
    locationmode = 'country names', sizes = c(1, 600), color = I("black")
  )
fig <- fig %>% add_markers(
    y = ~Lat, x = ~Lon, locations = ~Country,
    size = ~Value, color = ~abbrev, text = ~paste(Value, "cases")
  )
fig <- fig %>% add_text(
    x = 21.0936, y = 7.1881, text = 'Africa', showlegend = F, geo = "geo2"
  )
fig <- fig %>% add_trace(
    data = df9, z = ~Month, locations = ~Country,
    showscale = F, geo = "geo2"
  )
fig <- fig %>% layout(
    title = 'Ebola cases reported by month in West Africa 2014<br> Source: <a href="https://data.hdx.rwlabs.org/dataset/rowca-ebola-cases">HDX</a>',
    geo = g1, geo2 = g2
  )

fig
```