---
name: Buttons
permalink: r/custom-buttons/
description: How to add buttons in R with Plotly.
layout: base
thumbnail: thumbnail/custom-buttons.jpg
language: r
page_type: example_index
display_as: controls
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Methods

The [updatemenu method](https://plot.ly/r/reference/#layout-updatemenus-buttons-method) determines which [plotly.js](https://plot.ly/javascript/plotlyjs-function-reference/) function will be used to modify the chart. There are 4 possible methods:

* `"restyle"`: modify data or data attributes
* `"relayout"`: modify layout attributes
* `"update"`: modify data **and** layout attributes
* `"animate"`: start or pause an animation ([only available offline](https://cpsievert.github.io/plotly_book/key-frame-animations.html))

### Restyle Buttons

The `"restyle"` method should be used when modifying the data and data attributes of the graph   
**Update One Data Attribute** 
This example demostrates how to update a single data attribute: line `color` with the `"restyle"` method.

```{r}
library(plotly)

x <- seq(-2*pi, 2*pi, length.out = 1000)
df <- data.frame(x, y1 = sin(x))

fig <- plot_ly(df, x = ~x)
fig <- fig %>% add_lines(y = ~y1)


fig <- fig %>% layout(
  title = "Button Restyle",
  xaxis = list(domain = c(0.1, 1)),
  yaxis = list(title = "y"),
  updatemenus = list(
    list(
      type = "buttons",
      y = 0.8,
      buttons = list(
        
        list(method = "restyle",
             args = list("line.color", "blue"),
             label = "Blue"),
        
        list(method = "restyle",
             args = list("line.color", "red"),
             label = "Red")))
  ))

fig
```

**Update Several Data Attributes** 
This example demostrates how to update several data attributes: colorscale, chart `type`, and `colorscale` with the "restyle" method.

```{r}
library(plotly)

fig <- plot_ly(z = ~volcano, type = "heatmap", colorscale='Rainbow')

# chart option buttons
chart_types <- list(
  type = "buttons",
  direction = "right",
  xanchor = 'center',
  yanchor = "top",
  pad = list('r'= 0, 't'= 10, 'b' = 10),
  x = 0.5,
  y = 1.27,
  buttons = list(
    
    list(method = "restyle",
         args = list("type", "heatmap"),
         label = "Heatmap"),
    
    list(method = "restyle",
         args = list("type", "contour"),
         label = "Contour"),
    
    list(method = "restyle",
         args = list("type", "surface"),
         label = "Surface")
  ))

# color option buttons  
color_types <- list(
  type = "buttons",
  direction = "right",
  xanchor = 'center',
  yanchor = "top",
  pad = list('r'= 0, 't'= 10, 'b' = 10),
  x = 0.5,
  y = 1.17,
  buttons = list(
    
    list(method = "restyle",
         args = list("colorscale", "Rainbow"),
         label = "Rainbow"),
    
    list(method = "restyle",
         args = list("colorscale", "Jet"),
         label = "Jet"),
    
    list(method = "restyle",
         args = list("colorscale", "Earth"),
         label = "Earth"),
    
    list(method = "restyle",
         args = list("colorscale", "Electric"),
         label = "Electric")
  ))

annot <- list(list(text = "Chart<br>Type", x=0.2, y=1.25, xref='paper', yref='paper', showarrow=FALSE),
              list(text = "Color<br>Type", x=0.2, y=1.15, xref='paper', yref='paper', showarrow=FALSE))

# plot
fig <- fig %>% layout(
  xaxis = list(domain = c(0.1, 1)),
  yaxis = list(title = "y"),
  updatemenus = list(chart_types,color_types),
  annotations = annot)

fig
```

### Relayout Button

The `"relayout"` method should be used when modiying the layout attributes of the graph.   
**Update One Layout Attribute**
This example demostrated how to update a layout attribute: `shapes` with the '"relayout"' method.

```{r}
library(plotly)

x0 <- rnorm(400, mean=2, sd=0.4)
y0 <- rnorm(400, mean=2, sd=0.4)
x1 <- rnorm(400, mean=3, sd=0.6)
y1 <- rnorm(400, mean=6, sd=0.4)
x2 <- rnorm(400, mean=4, sd=0.2)
y2 <- rnorm(400, mean=4, sd=0.4)

# shapes components
cluster0 = list(
  type = 'circle',
  xref ='x', yref='y',
  x0=min(x0), y0=min(y0),
  x1=max(x0), y1=max(y0),
  opacity=0.25,
  line = list(color="#835AF1"),
  fillcolor="#835AF1")

cluster1 = list(
  type = 'circle',
  xref ='x', yref='y',
  x0=min(x1), y0=min(y1),
  x1=max(x1), y1=max(y1),
  opacity=0.25,
  line = list(color="#7FA6EE"),
  fillcolor="#7FA6EE")

cluster2 = list(
  type = 'circle',
  xref ='x', yref='y',
  x0=min(x2), y0=min(y2),
  x1=max(x2), y1=max(y2),
  opacity=0.25,
  line = list(color="#B8F7D4"),
  fillcolor="#B8F7D4")

# updatemenus component
updatemenus <- list(
  list(
    active = -1,
    type = 'buttons',
    buttons = list(
      
      list(
        label = "None",
        method = "relayout",
        args = list(list(shapes = c()))),
      
      list(
        label = "Cluster 0",
        method = "relayout",
        args = list(list(shapes = list(cluster0, c(), c())))),
      
      list(
        label = "Cluster 1",
        method = "relayout",
        args = list(list(shapes = list(c(), cluster1, c())))),
      
      list(
        label = "Cluster 2",
        method = "relayout",
        args = list(list(shapes = list(c(), c(), cluster2)))),
      
      list(
        label = "All",
        method = "relayout",
        args = list(list(shapes = list(cluster0,cluster1,cluster2))))
    )
  )
)

fig <- plot_ly(type = 'scatter', mode='markers') 
fig <- fig %>% add_trace(x=x0, y=y0, mode='markers', marker=list(color='#835AF1')) 
fig <- fig %>% add_trace(x=x1, y=y1, mode='markers', marker=list(color='#7FA6EE')) 
fig <- fig %>% add_trace(x=x2, y=y2, mode='markers', marker=list(color='#B8F7D4')) 
fig <- fig %>% layout(title = "Highlight Clusters", showlegend = FALSE,
         updatemenus = updatemenus)

fig
```

### Update Button

The '"update"' method should be used when modifying the data and layout sections of the graph.
This example demostrates how to update which traces are diplayed while simultaneously updating layout attributes such as the chart title and annotations. 

```{r}
library(plotly)
library(quantmod)

d <- quantmod::getSymbols("AAPL")

df <- data.frame(Date=index(AAPL),coredata(AAPL))

high_annotations <- list(
  x=df$Date[df$AAPL.High == max(df$AAPL.High)], 
  y=max(df$AAPL.High),
  xref='x', yref='y',
  text=paste0('High: $',max(df$AAPL.High)),
  ax=0, ay=-40
)

low_annotations <- list(
  x=df$Date[df$AAPL.Low == min(df$AAPL.Low)], 
  y=min(df$AAPL.Low),
  xref='x', yref='y',
  text=paste0('Low: $',min(df$AAPL.Low)),
  ax=0, ay=40
)

# updatemenus component
updatemenus <- list(
  list(
    active = -1,
    type= 'buttons',
    buttons = list(
      list(
        label = "High",
        method = "update",
        args = list(list(visible = c(FALSE, TRUE)),
                    list(title = "Apple High",
                         annotations = list(c(), high_annotations)))),
      list(
        label = "Low",
        method = "update",
        args = list(list(visible = c(TRUE, FALSE)),
                    list(title = "Apple Low",
                         annotations = list(low_annotations, c() )))),
      list(
        label = "Both",
        method = "update",
        args = list(list(visible = c(TRUE, TRUE)),
                    list(title = "Apple",
                         annotations = list(low_annotations, high_annotations)))),
      list(
        label = "Reset",
        method = "update",
        args = list(list(visible = c(TRUE, TRUE)),
                    list(title = "Apple",
                         annotations = list(c(), c())))))
  )
)

fig <- df 
fig <- fig %>% plot_ly(type = 'scatter', mode = 'lines') 
fig <- fig %>% add_lines(x=~Date, y=~AAPL.High, name="High",
            line=list(color="#33CFA5")) 
fig <- fig %>% add_lines(x=~Date, y=~AAPL.Low, name="Low",
            line=list(color="#F06A6A")) 
fig <- fig %>% layout(title = "Apple", showlegend=FALSE,
         xaxis=list(title="Date"),
         yaxis=list(title="Price ($)"),
         updatemenus=updatemenus)


fig
```

### Animate Button

Animations are currently only available in the [development package](https://plot.ly/r/getting-started/#installation) and [offline](https://cpsievert.github.io/plotly_book/key-frame-animations.html).

### Style Buttons

When adding buttons to Plotly charts, users have the option of styling the color, font, padding, and position of the buttons. The example below demostrates hot to apply different styling options. See all updatemenus styling attributes here: https://plot.ly/r/reference/#layout-updatemenus. 

#### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Mafig Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token

# Setting mapbox token for R environment 
Sys.setenv("MAPBOX_TOKEN" = mapboxToken)

# read in wind turbines and farms data
df_wind = read.csv('data/2805.csv')
df_farms = read.csv('https://plot.ly/~jackp/17256.csv', skipNul = TRUE)

# subset and clean data
df_sub <- subset(df_wind, manufac != "unknown", select=c(lat_DD, long_DD, manufac))
df_farms$Longitude = as.numeric(gsub("'", "", df_farms$Longitude))
df_farms$Wind.Farm = gsub("^((\\w+\\W+){2}).*","\\1",df_farms$Wind.Farm)
rm(df_wind)


# location buttons
usa <- list(method = "relayout",
            args = list(list(mapbox.center.lat = median(df_sub$lat_DD),
                             mapbox.center.lon = median(df_sub$long_DD),
                             mapbox.zoom = 3)),
            label = "USA")

buttons <- function(i) {
  list(method = "relayout",
       args = list(list(mapbox.center.lat = df_farms$Latitude[i],
                        mapbox.center.lon = df_farms$Longitude[i],
                        mapbox.zoom = 9)),
       label = df_farms$Wind.Farm[i]) 
}

# mafig style buttons
basic <- list(method = "relayout",
              args = list(list(mapbox.style = "basic")),
              label = "Basic")

dark <- list(method = "relayout",
             args = list(list(mapbox.style = "dark")),
             label = "Dark")

satellite <- list(method = "relayout",
                  args = list(list(mapbox.style = "satellite")),
                  label = "Satellite")  

# plot scattermapbox with buttons 
fig <- plot_mapbox(df_sub, lat = ~lat_DD, lon = ~long_DD, mode = 'scattermapbox',
                 split = ~manufac, size=3) 
fig <- fig %>% add_annotations(x = 0.05, y = 0.05,
                  text = "All US wind turbines (scroll to zoom)",
                  xref = "page",
                  yref = "page",
                  showarrow = FALSE,
                  font = list(color = 'magenta',
                              size = 14)) 
fig <- fig %>% layout(plot_bgcolor = 'black', 
         paper_bgcolor = 'black', 
         mapbox = list(center = list(lat = median(df_sub$lat_DD), 
                                     lon = median(df_sub$long_DD)), 
                       zoom = 3, style = 'dark'),
         updatemenus = list(
           
           list(type='buttons',
                direction = "right",
                xanchor = 'center',
                yanchor = "top",
                pad = list('r'= 0, 't'= 10, 'b' = 10),
                x = 0.5,
                y = 1,
                showactive = FALSE,
                buttons=list(usa, buttons(1),
                             buttons(2), buttons(3), 
                             buttons(4))),
           
           list(type='buttons',
                direction = "right",
                xanchor = 'center',
                yanchor = "top",
                pad = list('r'= 0, 't'= 10, 'b' = 10),
                x = 0.5,
                y = 0.935,
                showactive = FALSE,
                buttons=list(buttons(6), buttons(7),
                             buttons(8), buttons(9))),
           
           list(type='buttons',
                direction = "right",
                yanchor = "bottom",
                x = 1,
                y = 0,
                buttons=list(dark,basic,satellite)))) 
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

#Reference

See [https://plot.ly/r/reference/#layout-updatemenus](https://plot.ly/r/reference/#layout-updatemenus) for more information and options!
