---
description: How to create carpet plots in R with Plotly.
display_as: scientific
language: r
layout: base
name: Carpet Plot
order: 8
output:
  html_document:
    keep_md: true
permalink: r/carpet-plot/
thumbnail: thumbnail/carpet.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Set the Coordinates 

To set the `x` and `y` coordinates use `x` and `y` attributes. If `x` coorindate values are ommitted a cheater plot will be created.

```{r}
library(plotly)

fig <- plot_ly(
    type = 'carpet',
    y = c(2, 3.5, 4, 3, 4.5, 5, 5.5, 6.5, 7.5, 8, 8.5, 10))
```

### Add Parameter Values

To save parameter values use `a` and `b` attributes.

```{r}
library(plotly)

fig <- plot_ly(
    type = 'carpet',
    a = c(4, 4, 4, 4.5, 4.5, 4.5, 5, 5, 5, 6, 6, 6),
    b = c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3),
    y = c(2, 3.5, 4, 3, 4.5, 5, 5.5, 6.5, 7.5, 8, 8.5, 10))

fig
```

### Add Carpet Axes

Use `aaxis` or `baxis` lists to make changes to the axes. For a more detailed list of attributes refer to [R reference](https://plot.ly/r/reference/#carpet-aaxis).

```{r}
library(plotly)

fig <- plot_ly(
    type = 'carpet',
    a = c(4, 4, 4, 4.5, 4.5, 4.5, 5, 5, 5, 6, 6, 6),
    b = c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3),
    y = c(2, 3.5, 4, 3, 4.5, 5, 5.5, 6.5, 7.5, 8, 8.5, 10),
    aaxis = list(
      tickprefix = 'a = ',
      ticksuffix = 'm',
      smoothing = 1,
      minorgridcount = 9
      ),
    baxis = list(
      tickprefix = 'b = ',
      ticksuffix = 'Pa',
      smoothing = 1,
      minorgridcount = 9
      )
    )

fig
```

### Style Carpet Axes

```{r}
library(plotly)

fig <- plot_ly(
  type = 'carpet',
  a = c(4, 4, 4, 4.5, 4.5, 4.5, 5, 5, 5, 6, 6, 6),
  b = c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3),
  y = c(2, 3.5, 4, 3, 4.5, 5, 5.5, 6.5, 7.5, 8, 8.5, 10),
  aaxis = list(
    tickprefix = 'a = ',
    ticksuffix = 'm',
    smoothing = 1,
    minorgridcount = 9,
    minorgridwidth = 0.6,
    minorgridcolor = 'white',
    gridcolor = 'white',
    color = 'white'
  ),
  baxis = list(
    tickprefix = 'b = ',
    ticksuffix = 'Pa',
    smoothing = 1,
    minorgridcount = 9,
    minorgridwidth = 0.6,
    gridcolor = 'white',
    minorgridcolor = 'white',
    color = 'white'
  )
) 
fig <- fig %>% layout(
    plot_bgcolor = 'black', paper_bgcolor = 'black',
    xaxis = list(showgrid = F, showticklabels = F),
    yaxis = list(showgrid = F, showticklabels = F)
  )

fig
```

### Add Points and Contours

To add points and lines to see [Carpet Scatter Plots](https://plot.ly/r/carpet-scatter) or to add contours see [Carpet Contour Plots](https://plot.ly/r/carpet-contour)

### Reference

See [https://plot.ly/r/reference/#carpet](https://plot.ly/r/reference/#carpet) for more information and options!