---
name: Aggregations
permalink: r/aggregations/
description: How to use aggregates in R with Plotly.
layout: base
thumbnail: thumbnail/aggregations.jpg
language: r
display_as: transforms
page_type: example_index
order: 3
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Introduction

Aggregates are a type of transform that can be applied to values in a given expression. Available aggregations are:

Function | Description
------------- | -------------
`count` | Returns the quantity of items for each group.
`sum` | Returns the summation of all numeric values.
`avg` | Returns the average of all numeric values.
`median` | Returns the median of all numeric values.
`mode` | Returns the mode of all numeric values.
`rms` | Returns the rms of all numeric values.
`stddev` | Returns the standard deviation of all numeric values.
`min` | Returns the minimum numeric value for each group.
`max` | Returns the maximum numeric value for each group.
`first` | Returns the first numeric value for each group.
`last` | Returns the last numeric value for each group.

#### Basic Example

```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = diamonds$cut,
  y = diamonds$price,
  mode = 'markers',
  transforms = list(
    list(
      type = 'aggregate',
      groups = diamonds$cut,
      aggregations = list(
        list(
          target = 'y', func = 'sum', enabled = T
        )
      )
    )
  )
)

fig
```

#### Aggregate Functions

``` {r}
library(plotly)

s <- schema()
agg <- s$transforms$aggregate$attributes$aggregations$items$aggregation$func$values


l = list()
for (i in 1:length(agg)) {
  ll = list(method = "restyle",
            args = list('transforms[0].aggregations[0].func', agg[i]),
            label = agg[i]) 
  l[[i]] = ll
}

fig <- plot_ly(
  type = 'scatter',
  x = diamonds$cut,
  y = diamonds$price,
  mode = 'markers',
  marker = list(
    size = 10,
    color = 'blue',
    opacity = 0.8
  ),
  transforms = list(
    list(
      type = 'aggregate',
      groups = diamonds$cut,
      aggregations = list(
        list(
          target = 'y', func = 'avg', enabled = T
        )
      )
    )
  )
)
fig <- fig %>% layout(
    title = '<b>Plotly Aggregations</b><br>use dropdown to change aggregation',
    xaxis = list(title = 'Cut'),
    yaxis = list(title = 'Price ($)'),
    updatemenus = list(
      list(
        x = 0.25,
        y = 1.04,
        xref = 'paper',
        yref = 'paper',
        yanchor = 'top',
        buttons = l
      )
    )
  )

fig
```


#### Histogram Binning

``` {r}
library(plotly)

df <- read.csv("https://plot.ly/~public.health/17.csv")

labels <- function(size, label) {
  list(
    args = c("xbins.size", size), 
    label = label, 
    method = "restyle"
  )
}

fig <- df 
fig <- fig %>%
  plot_ly(
    x = ~date,
    autobinx = FALSE, 
    autobiny = TRUE, 
    marker = list(color = "rgb(68, 68, 68)"), 
    name = "date", 
    type = "histogram", 
    xbins = list(
      end = "2016-12-31 12:00", 
      size = "M1", 
      start = "1983-12-31 12:00"
    )
  )
  fig <- fig %>% layout(
    paper_bgcolor = "rgb(240, 240, 240)", 
    plot_bgcolor = "rgb(240, 240, 240)", 
    title = "<b>Shooting Incidents</b><br>use dropdown to change bin size",
    xaxis = list(
      type = 'date'
    ),
    yaxis = list(
      title = "Incidents"
    ),
    updatemenus = list(
      list(
        x = 0.1, 
        y = 1.15,
        active = 1, 
        showactive = TRUE,
        buttons = list(
          labels("D1", "Day"),
          labels("M1", "Month"),
          labels("M6", "Half Year"),
          labels("M12", "Year")
        )
      )
    )
  )

fig
```

#### Mapping with Aggregations

``` {r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/worldhappiness.csv")

s <- schema()
agg <- s$transforms$aggregate$attributes$aggregations$items$aggregation$func$values


l = list()
for (i in 1:length(agg)) {
  ll = list(method = "restyle",
            args = list('transforms[0].aggregations[0].func', agg[i]),
            label = agg[i]) 
  l[[i]] = ll
}

fig <- df 
fig <- fig %>%
  plot_ly(
    type = 'choropleth',
    locationmode = 'country names',
    locations = ~Country,
    z = ~HappinessScore,
    autocolorscale = F,
    reversescale = T,
    colorscale = 'Portland', 
    transforms = list(list(
      type = 'aggregate',
      groups = ~Country,
      aggregations = list(
        list(target = 'z', func = 'sum', enabled = T)
      )
    ))
  )
fig <- fig %>% layout(
    title = "<b>World Happiness</b>",
    geo = list(
      showframe = F,
      showcoastlines = F
    ),
    updatemenus = list(
      list(
        x = 0.25,
        y = 1.04,
        xref = 'paper',
        yref = 'paper',
        yanchor = 'top',
        buttons = l
      )
    )
  )

fig
```

#### Reference

See [https://plot.ly/r/reference/](https://plot.ly/r/reference/) for more information and options!
