---
description: How to create county-level choropleths in R with Plotly.
display_as: maps
language: r
layout: base
name: County Level Choropleth
order: 8
output:
  html_document:
    keep_md: true
permalink: r/county-level-choropleth/
thumbnail: thumbnail/county-level-choropleth.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

### New to Plotly?

Plotly's R library is free and open source!<br>
[Get started](https://plot.ly/r/getting-started/) by downloading the client and [reading the primer](https://plot.ly/r/getting-started/).<br>
You can set up Plotly to work in [online](https://plot.ly/r/getting-started/#hosting-graphs-in-your-online-plotly-account) or [offline](https://plot.ly/r/offline/) mode.<br>
We also have a quick-reference [cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf) (new!) to help you get started!

### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.
```{r}
library(plotly)
packageVersion('plotly')
```

### Mapbox Access Tokens and When You Need Them

  The word "mapbox" in the trace names and `layout.mapbox` refers to the Mapbox.js open-source library. If your basemap in `layout.mapbox.style` uses data from the Mapbox *service*, then you will need to register for a free account at https://mapbox.com/ and obtain a Mapbox Access token. This token should be provided either in `mapboxAccessToken`  in `setPlotConfig` function, or as a variable that would be passed as an argument of `newPlot`.
    If your `layout.mapbox.style` does not use data from the Mapbox service, you do *not* need to register for a Mapbox account.
    <h6>Base Maps in `layout.mapbox.style`</h6>
    The accepted values for `layout.mapbox.style` are one of the following tiles.
    <ol>
        <li> `"white-bg"` yields an empty white canvas which results in no external HTTP requests </li>
        <li> `"open-street-map"`, `"carto-positron"`, `"carto-darkmatter"`, `"stamen-terrain"`, `"stamen-toner"` or `"stamen-watercolor"` yeild maps composed of *raster* tiles from various public tile servers which do not require signups or access tokens </li>
        <li> `"basic"`, `"streets"`, `"outdoors"`, `"light"`, `"dark"`, `"satellite"`, or `"satellite-streets"` yeild maps composed of *vector* tiles from the Mapbox service, and *do* require a Mapbox Access Token or an on-premise Mapbox installation. </li>
        <li> A Mapbox service style URL, which requires a Mapbox Access Token or an on-premise Mapbox installation. </li>
        <li> A Mapbox Style object as defined at https://docs.mapbox.com/mapbox-gl-js/style-spec/ </li>
    </ol>

### Creating Polygon Boundaries

```{r}
library(plotly)

blank_layer <- list(
  title = "",
  showgrid = F,
  showticklabels = F,
  zeroline = F)

p <- map_data("county") %>%
  filter(region == 'california') %>%
  group_by(group) %>%
  plot_ly(
    x = ~long,
    y = ~lat,
    fillcolor = 'white',
    hoverinfo = "none") %>%
  add_polygons(
    line = list(color = 'black', width = 0.5)) %>%
  layout(
    xaxis = blank_layer,
    yaxis = blank_layer)

p
```

### Add County-Level Data

```{r}
library(tidyverse)
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/californiaPopulation.csv")

cali <- map_data("county") %>%
  filter(region == 'california')

pop <- df %>%
  group_by(County.Name) %>%
  summarise(Pop = sum(Population))

pop$County.Name <- tolower(pop$County.Name) # matching string

cali_pop <- merge(cali, pop, by.x = "subregion", by.y = "County.Name")

cali_pop$pop_cat <- cut(cali_pop$Pop, breaks = c(seq(0, 11000000, by = 500000)), labels=1:22)

p <- cali_pop %>%
  group_by(group) %>%
  plot_ly(x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
          text = ~subregion, hoverinfo = 'text') %>%
  add_polygons(line = list(width = 0.4)) %>%
  add_polygons(
    fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) %>%
  layout(
    title = "California Population by County",
    titlefont = list(size = 10),
    xaxis = list(title = "", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE),
    yaxis = list(title = "", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE)
  )

p
```

### Add Polygon to a Map Projection

```{r}
library(plotly)

geo <- list(
  scope = 'usa',
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80")
)

p <- cali_pop %>%
  group_by(group) %>%
  plot_geo(
    x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
    text = ~subregion, hoverinfo = 'text') %>%
  add_polygons(line = list(width = 0.4)) %>%
  add_polygons(
    fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) %>%
  layout(
    title = "California Population by County",
    geo = geo)

p
```

### Add Polygon to Mapbox
```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token

# Setting mapbox token for R environment 
Sys.setenv("MAPBOX_TOKEN" = mapboxToken)

p <- cali_pop %>%
  group_by(group) %>%
  plot_mapbox(x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
          text = ~subregion, hoverinfo = 'text', showlegend = FALSE) %>%
  add_polygons(
    line = list(width = 0.4)
  ) %>%
  add_polygons(fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) %>%
  layout(
    xaxis = list(title = "", showgrid = FALSE, showticklabels = FALSE),
    yaxis = list(title = "", showgrid = FALSE, showticklabels = FALSE),
    mapbox = list(
      style = 'light',
      zoom = 4,
      center = list(lat = ~median(lat), lon = ~median(long))),
    margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)
  ) %>%
  config(mapboxAccessToken = mapboxToken)

p
```