---
description: How to create scattermapbox plots in R with Plotly.
display_as: maps
language: r
layout: base
name: Scattermapbox
order: 7
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/scattermapbox/
thumbnail: thumbnail/scatter-mapbox.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Mafig Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Basic Example

```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca


dat <- map_data("world", "canada") 
dat <- dat %>% group_by(group)

fig <- plot_mapbox(dat, x = ~long, y = ~lat)
fig <- fig %>% add_paths(size = I(2))
fig <- fig %>% add_segments(x = -100, xend = -50, y = 50, 75)
fig <- fig %>% layout(mapbox = list(zoom = 0,
                       center = list(lat = ~median(lat),
                                     lon = ~median(long))
  ))
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

### Multiple Markers

```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

df = read.csv('https://raw.githubusercontent.com/bcdunbar/datasets/master/meteorites_subset.csv')

fig <- df 
fig <- fig %>% plot_mapbox(lat = ~reclat, lon = ~reclong,
              split = ~class, size=2,
              mode = 'scattermapbox', hoverinfo='name') 
fig <- fig %>% layout(title = 'Meteorites by Class',
         font = list(color='white'),
         plot_bgcolor = '#191A1A', paper_bgcolor = '#191A1A',
         mapbox = list(style = 'dark'),
         legend = list(orientation = 'h',
                       font = list(size = 8)),
         margin = list(l = 25, r = 25,
                       b = 25, t = 25,
                       pad = 2)) 
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

### Adding lines to Mapbox

```{r}
library(plotly)
library(dplyr)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

# airport locations
air <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2011_february_us_airport_traffic.csv')

# flights between airports
flights <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2011_february_aa_flight_paths.csv')
flights$id <- seq_len(nrow(flights))

fig <- plot_mapbox(mode = 'scattermapbox') 
fig <- fig %>% add_markers(
    data = air, x = ~long, y = ~lat, text=~airport, color=I("red"),
    size = ~cnt, hoverinfo = "text", alpha = 0.5) 
fig <- fig %>% add_segments(
    data = group_by(flights, id),
    x = ~start_lon, xend = ~end_lon,
    y = ~start_lat, yend = ~end_lat,
    alpha = 0.3, size = I(1), hoverinfo = "none",
    color=I("red")) 
fig <- fig %>% layout(
    plot_bgcolor = '#191A1A', paper_bgcolor = '#191A1A',
    mapbox = list(style = 'dark',
                  zoom = 1.5,
                  center = list(lat = median(air$lat),
                                lon = median(air$long))),
    margin = list(l = 0, r = 0,
                  b = 0, t = 0,
                  pad = 0),
    showlegend=FALSE) 
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

#Reference

See [https://plot.ly/r/reference/#scattermapbox](https://plot.ly/r/reference/#scattermapbox) for more information and options!