---
name: Colorscales
permalink: r/colorscales/
description: How to create colorscales in R with Plotly.
layout: base
language: r
page_type: u-guide
display_as: file_settings
order: 15
thumbnail: thumbnail/heatmap_colorscale.jpg
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Colorscale for Scatter Plots

```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  mode='markers',
  y=rep(5, 40),
  marker=list(
    size=seq(0, 39),
    color=seq(0, 39),
    colorbar=list(
      title='Colorbar'
    ),
    colorscale='Viridis',
    reversescale =T
  )
  )
fig <- fig %>% layout(
    xaxis = list(
      showgrid = F,
      zeroline = F
    ),
    yaxis = list(
      showgrid = F,
      zeroline = F
    )
  )
```

### Colorscale Contour

```{r}
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z=matrix(c(10, 10.625, 12.5, 15.625, 20,
      5.625, 6.25, 8.125, 11.25, 15.625,
      2.5, 3.125, 5., 8.125, 12.5,
      0.625, 1.25, 3.125, 6.25, 10.625,
      0, 0.625, 2.5, 5.625, 10),
      nrow=5, ncol=5)
)

fig
```

### Share Color Axis

This example shows how to specify the color scale and color bar per trace. To share colorscale information in multiple subplots, you can use [coloraxis](https://plot.ly/r/reference/#scatter-marker-line-coloraxis).
Below we show how to set a reference coloraxis1 to a shared coloraxis, which are set in the layout. Note that multiple color scales can be linked to the same color.

```{r}

library(plotly)

fig1 <- plot_ly(
    type = "heatmap",
    x = c(1,2,3,4),
    z = list(c(1,2,3,4), c(4,-3,-1,1)),
    coloraxis = 'coloraxis')
fig2 <- plot_ly(
    type = "heatmap",
    x = c(3,4,5,6),
    z = list(c(10,2,1,0), c(4,3,5,6)),
    coloraxis = 'coloraxis')
fig <- subplot(fig1, fig2)
fig <- fig %>% layout(coloraxis=list(colorscale='Jet'))

fig
```

#Reference

See [https://plot.ly/r/reference/#heatmap-colorscale](https://plot.ly/r/reference/#heatmap-colorscale) for more information and options!