---
name: Intro to Animations
permalink: r/animations/
description: How to create animations in R with Plotly.
layout: base
thumbnail: thumbnail/animations.gif
language: r
page_type: example_index
display_as: animations
order: 1
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Frames

Now, along with `data` and `layout`, `frames` is added to the keys that `figure` allows. Your `frames` key points to a list of figures, each of which will be cycled through upon instantiation of the plot.

### Basic Example

```{r}
library(plotly)

df <- data.frame(
  x = c(1,2,1), 
  y = c(1,2,1), 
  f = c(1,2,3)
)

fig <- df 
fig <- fig %>%
  plot_ly(
    x = ~x,
    y = ~y,
    frame = ~f,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
  )

fig
```

### Mulitple Trace Animations

```{r}
library(plotly)
library(gapminder)

fig <- gapminder 
fig <- fig %>%
  plot_ly(
    x = ~gdpPercap, 
    y = ~lifeExp, 
    size = ~pop, 
    color = ~continent, 
    frame = ~year, 
    text = ~country, 
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
fig <- fig %>% layout(
    xaxis = list(
      type = "log"
    )
  )

fig
```

### Add Animation Options

```{r}
library(plotly)



fig <- fig %>%
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  )

fig
```

### Add Button Options

```{r}
library(plotly)



fig <- fig %>%
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  )

fig
```

### Add Slider Options

```{r}
library(plotly)


fig <- fig %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )

fig
```

### Advanced Example

```{r}
library(plotly)

fig <- gapminder 
fig <- fig %>%
  plot_ly(
    x = ~gdpPercap, 
    y = ~lifeExp, 
    size = ~pop, 
    color = ~continent, 
    frame = ~year, 
    text = ~country, 
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
fig <- fig %>% layout(
    xaxis = list(
      type = "log"
    )
  )
fig <- fig %>% animation_opts(
    1000, easing = "elastic", redraw = FALSE
  )
fig <- fig %>% animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  )
fig <- fig %>% animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(color="red"))
  )

fig
```

### Reference 

To read more on animations see [The Plotly Book](https://cpsievert.github.io/plotly_book/key-frame-animations.html).
