---
description: How to create county-level choropleths in R with Plotly.
display_as: maps
language: r
layout: base
name: County Level Choropleth
order: 8
output:
  html_document:
    keep_md: true
permalink: r/county-level-choropleth/
thumbnail: thumbnail/county-level-choropleth.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Mafig Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Creating Polygon Boundaries

```{r}
library(plotly)

blank_layer <- list(
  title = "",
  showgrid = F,
  showticklabels = F,
  zeroline = F)

fig <- map_data("county") 
fig <- fig %>% filter(region == 'california') 
fig <- fig %>% group_by(group) 
fig <- fig %>% plot_ly(
    x = ~long,
    y = ~lat,
    fillcolor = 'white',
    hoverinfo = "none") 
fig <- fig %>% add_polygons(
    line = list(color = 'black', width = 0.5)) 
fig <- fig %>% layout(
    xaxis = blank_layer,
    yaxis = blank_layer)

fig
```

### Add County-Level Data

```{r}
library(tidyverse)
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/californiaPopulation.csv")

cali <- map_data("county") 
cali <- cali %>% filter(region == 'california')

pop <-  df 
pop <- pop %>% group_by(County.Name) 
pop <- pop %>% summarise(Pop = sum(Population))

pop$County.Name <- tolower(pop$County.Name) # matching string

cali_pop <-  merge(cali, pop, by.x = "subregion", by.y = "County.Name")

cali_pop$pop_cat <- cut(cali_pop$Pop, breaks = c(seq(0, 11000000, by = 500000)), labels=1:22)

fig <- cali_pop 
fig <- fig %>% group_by(group) 
fig <- fig %>% plot_ly(x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
          text = ~subregion, hoverinfo = 'text') 
fig <- fig %>% add_polygons(line = list(width = 0.4)) 
fig <- fig %>% add_polygons(
    fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) 
fig <- fig %>% layout(
    title = "California Population by County",
    titlefont = list(size = 10),
    xaxis = list(title = "", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE),
    yaxis = list(title = "", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE)
  )

fig
```

### Add Polygon to a Map Projection

```{r}
library(plotly)

geo <- list(
  scope = 'usa',
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("gray80")
)

fig <- cali_pop 
fig <- fig %>% group_by(group) 
fig <- fig %>% plot_geo(
    x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
    text = ~subregion, hoverinfo = 'text') 
fig <- fig %>% add_polygons(line = list(width = 0.4)) 
fig <- fig %>% add_polygons(
    fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) 
fig <- fig %>% layout(
    title = "California Population by County",
    geo = geo)

fig
```

### Add Polygon to Mapbox
```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

fig <- cali_pop
fig <- fig %>% group_by(group) 
fig <- fig %>% plot_mapbox(x = ~long, y = ~lat, color = ~pop_cat, colors = c('#ffeda0','#f03b20'),
          text = ~subregion, hoverinfo = 'text', showlegend = FALSE) 
fig <- fig %>% add_polygons(
    line = list(width = 0.4)
  ) 
fig <- fig %>% add_polygons(fillcolor = 'transparent',
    line = list(color = 'black', width = 0.5),
    showlegend = FALSE, hoverinfo = 'none'
  ) 
fig <- fig %>% layout(
    xaxis = list(title = "", showgrid = FALSE, showticklabels = FALSE),
    yaxis = list(title = "", showgrid = FALSE, showticklabels = FALSE),
    mapbox = list(
      style = 'light',
      zoom = 4,
      center = list(lat = ~median(lat), lon = ~median(long))),
    margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)
  ) 
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```