---
description: How to customize 3d hover options for plots in R with Plotly.
display_as: file_settings
language: r
layout: base
name: 3D Hover Options
order: 8
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/3d-hover/
thumbnail: thumbnail/hover-text.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Customize Hover for Spikelines

By default, Plotly's 3D plots display lines called "spikelines" while hovering over a point.
These lines project from the hover point to each of the three axes' normal planes and
then extend from those projection data points to the planes' wall boundaries.

```{r}
library(plotly)

mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

fig <-  plot_ly(mtcars, x = ~wt, y = ~hp, z = ~qsec, opacity = 0.8, color = ~am, colors = c('#BF382A', '#0C4B8E')) 
fig <- fig %>%
  add_markers() 
fig <- fig %>%
  layout(
    scene = list(
      xaxis = list(
          spikecolor = '#a009b5',
          spikesides = FALSE,
  		  spikethickness = 6
  	 	),
  	  yaxis = list(
  		  spikecolor = '#a009b5',
  		  spikesides = FALSE,
  		  spikethickness = 6  		
  		),
  	  zaxis = list(
  		  spikecolor = '#a009b5',
  		  spikethickness = 6
  	  )
    )
  )

fig
```

### Customize Hover for Surface Contours

In addition to spikelines, Plotly 3D Surface plots also display surface contours on hover by default.
These are customized by styling the [`contours`](https://plot.ly/r/reference/#surface-contours)
attribute in the surface trace.

```{r}
library(plotly)

fig <-  plot_ly(z = ~volcano) %>% add_surface(
  contours = list(
    x = list( 
      highlight = TRUE,
	  highlightcolor = "#41a7b3"
	  ),
      y = list(highlight = FALSE),
      z = list(highlight = FALSE)
  )
) 
fig <- fig %>%
  layout(
    scene = list(
      xaxis = list(showspikes=FALSE),
  	  yaxis = list(showspikes=FALSE),
  	  zaxis = list(showspikes=FALSE)
    )
  )


fig
```

# Reference

See [https://plot.ly/r/reference/#layout-scene-xaxis](https://plot.ly/r/reference/#layout-scene-xaxis) and [https://plot.ly/r/reference/#surface-contours](https://plot.ly/r/reference/#surface-contours) for more information and options!