---
description: How to style markers in R.
display_as: file_settings
language: r
layout: base
name: Marker Styling
order: 16
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/marker-style/
thumbnail: thumbnail/marker-style.gif
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Add Marker Border
In order to make markers distinct, you can add a border to the markers. This can be achieved by adding the line dict to the marker dict. For example, `marker:{..., line: {...}}`.

```{r}
library(plotly)

x <- runif(500, min=3, max=6)
y <- runif(500, min=3, max=6)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = x,
    y = y,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 20,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    showlegend = F
  ) 
fig <- fig %>%
  add_trace(
    x = c(2),
    y = c(4.5),
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 120,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 12
      )
    ),
    showlegend = F
  )

fig
```

#### Fully Opaque
Fully opaque, the default setting, is useful for non-overlapping markers. When many points overlap it can be hard to observe density.  

```{r}
library(plotly)

x <- runif(500, min=3, max=6)
y <- runif(500, min=3, max=6)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = x,
    y = y,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 20,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    showlegend = F
  ) 
fig <- fig %>%
  add_trace(
    x = c(2,2),
    y = c(4.25,4.75),
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 120,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 12
      )
    ),
    showlegend = F
  )

fig
```

#### Opacity
Setting opacity outside the marker will set the opacity of the trace. Thus, it will allow greater visbility of additional traces but like fully opaque it is hard to distinguish density.

```{r}
library(plotly)

x <- runif(500, min=3, max=6)
y <- runif(500, min=3, max=4.5)
x2 <- runif(500, min=3, max=6)
y2 <- runif(500, min=4.5, max=6)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = x,
    y = y,
    opacity = 0.5,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 20,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    name = 'Opacity 0.5'
  ) 
fig <- fig %>%
  add_trace(
    x = x2,
    y = y2,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 20,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    name = 'Opacity 1.0'
  ) 
fig <- fig %>%
  add_trace(
    x = c(2,2),
    y = c(4.25,4.75),
    opacity = 0.5,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 120,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 12
      )
    ),
    showlegend = F
  )

fig
```

#### Marker Opacity
To maximise visibility of density, it is recommended to set the opacity inside the marker `marker:{opacity:0.5}`. If mulitple traces exist with high density, consider using marker opacity in conjunction with trace opacity.

```{r}
library(plotly)

x <- runif(500, min=3, max=6)
y <- runif(500, min=3, max=6)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = x,
    y = y,
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 20,
      opacity = 0.5,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    showlegend = F
  ) 
fig <- fig %>%
  add_trace(
    x = c(2,2),
    y = c(4.25,4.75),
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 120,
      opacity = 0.5,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 12
      )
    ),
    showlegend = F
  )

fig
```

#### Color Opacity
To maximise visibility of each point, set the color opacity by using alpha: `marker:{color: 'rgba(0,0,0,0.5)'}`. Here, the marker line will remain opaque.

```{r}
library(plotly)

x <- runif(500, min=3, max=6)
y <- runif(500, min=3, max=6)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = x,
    y = y,
    marker = list(
      color = 'rgba(17, 157, 255,0.5)',
      size = 20,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    showlegend = F
  ) 
fig <- fig %>%
  add_trace(
    x = c(2,2),
    y = c(4.25,4.75),
    marker = list(
      color = 'rgba(17, 157, 255,0.5)',
      size = 120,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 12
      )
    ),
    showlegend = F
  )

fig
```

### Reference

See [https://plot.ly/r/reference/](https://plot.ly/r/reference/) for more information and chart attribute options!