---
description: How to set locales with Plotly.
display_as: file_settings
language: r
layout: base
name: Locales
order: 17
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/locales/
thumbnail: thumbnail/mapbox-layers.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```



### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.

```{r}
library(plotly)
packageVersion('plotly')
```

### Setting Locale
To change the localization, you can use the `locale` argument in the `config()` function. 

``` {r, results = 'hide'}
library(plotly)

x <- seq.Date(Sys.Date(), Sys.Date() + 360, by = "day")

p <- plot_ly(x = x, y = rnorm(length(x))) %>%
   add_lines() %>%
  config(locale = 'zh-CN')
```

### Reference

See `?plotly::config` for more information about config options or [click here](https://plot.ly/r/configuration-options/). For more information regarding localization, [click here](https://github.com/plotly/plotly.js/tree/master/dist#to-include-localization).