---
name: Hover Text and Formatting
permalink: r/hover-text-and-formatting/
description: How to use hover text and formatting in R with Plotly.
layout: base
thumbnail: thumbnail/hover-text.jpg
language: r
display_as: file_settings
order: 7
page_type: u-guide
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Add Hover Text

```{r}
library(plotly)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = c(1:5), 
    y = rnorm(5, mean = 5),
    text = c("Text A", "Text B", "Text C", "Text D", "Text E"),
    hoverinfo = 'text',
    marker = list(color='green'),
    showlegend = F
  )

fig
```

#### Format Hover Text 

```{r}
library(plotly)

fig <-  plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = c(1:100), 
    y = rnorm(100, mean = 5), 
    marker = list(color='green'),
    hoverinfo = 'y',
    showlegend = F
  ) 
fig <- fig %>%
  layout(
    title = list(text="Set hover text formatting<br><a href= https://github.com/d3/d3-time-format/blob/master/README.md#locale_format>https://github.com/d3/d3-time-format/blob/master/README.md#locale_format</a>",
      size = 10),
    xaxis = list(zeroline = F),
    yaxis = list(hoverformat = '.2f'))

fig
```

#### Customize Tooltip Text with a Hovertemplate

To customize the tooltip on your graph you can use [hovertemplate](https://plot.ly/r/reference/#pie-hovertemplate), which is a template string used for rendering the information that appear on hoverbox.
This template string can include `variables` in %{variable} format, `numbers` in [d3-format's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Formatting.md#d3_forma), and `date` in [d3-time-fomrat's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Time-Formatting.md#format).
Hovertemplate customize the tooltip text vs. [texttemplate](https://plot.ly/r/reference/#pie-texttemplate) which customizes the text that appears on your chart. <br>
Set the horizontal alignment of the text within tooltip with [hoverlabel.align](https://plot.ly/r/reference/#layout-hoverlabel-align).

```{r}
library(plotly)

fig <-  plot_ly() 
fig <- fig %>%
  add_trace(
    type = 'scatter',
    mode = 'lines+markers',
    x = c(1,2,3,4,5),
    y = c(2.02825,1.63728,6.83839,4.8485,4.73463),
    text = c("Text A", "Text B", "Text C", "Text D", "Text E"),
    hovertemplate = paste('<i>Price</i>: $%{y:.2f}',
                        '<br><b>X</b>: %{x}<br>',
                        '<b>%{text}</b>'),
    showlegend = FALSE
  ) 
fig <- fig %>%
  add_trace(
    type = 'scatter',
    mode = 'lines+markers',
    x = c(1,2,3,4,5),
    y = c(3.02825,2.63728,4.83839,3.8485,1.73463),
    hovertemplate = 'Price: %{y:$.2f}<extra></extra>',
    showlegend = FALSE
  )

fig
```

```{r}
library(plotly)
fig <-   plot_ly() 
fig <- fig %>%
  add_trace(
    type = "pie",
    name = "",
    values = c(2, 1, 3, 2.5),
    labels = c("R", "Python", "Java Script", "Matlab"),
    text = c("textA", "TextB", "TextC", "TextD"),
    hovertemplate = "%{label}: <br>Popularity: %{percent} </br> %{text}")


fig
```

#### Advanced Hovertemplate

```{r}
library(plotly)

d <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/job-automation-probability.csv')

fig <-  d 

fig <- fig %>%
  plot_ly(
    type = 'scatter',
    mode = 'markers',
    x = ~prob,
    y = ~Average.annual.wage,
    marker = list(size = ~numbEmployed, sizeref = 4000, sizemode = 'area'),
    color = ~education,
    text = ~short.occupation,
    hovertemplate = paste(
      "<b>%{text}</b><br><br>",
      "%{yaxis.title.text}: %{y:$,.0f}<br>",
      "%{xaxis.title.text}: %{x:.0%}<br>",
      "Number Employed: %{marker.size:,}",
      "<extra></extra>"
      )
    ) 
fig <- fig %>%
  layout(legend = list(orientation = 'h', y = -0.3))

fig
```

#### Reference

See [https://plot.ly/r/reference/](https://plot.ly/r/reference/) for more information and options!
