---
description: How to make maps with Plotly and simple features.
display_as: maps
language: r
layout: base
name: Simple Features
order: 11
output:
  html_document:
    keep_md: true
permalink: r/maps-sf/
thumbnail: thumbnail/sf.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Introduction

In order to complete the examples below, you'll require installing additional packages (`install.packages("packageName")`):
- [sf](https://github.com/r-spatial/sf)

The examples below use the library [simple features](https://r-spatial.github.io/sf/) to read in the shape files before plotting the features with Plotly.

### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Map Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Basic sf

```{r}
library(plotly)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

fig <-  plot_ly(nc)

fig
```

You can also use `plot_geo`:

```{r}
library(plotly)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

fig <-  plot_geo(nc)

fig
```

Or `plot_mapbox`:

```{r}
library(plotly)
library(sf)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

fig <-  plot_mapbox(nc, split=~NAME) 
fig <- fig %>%
  layout(
    mapbox = list(
      zoom = 6
    )
  ) 
fig <- fig %>%
  config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

### Using ggplot2

Alternatively, if working in ggplot2 you can use [`ggplotly`](https://plot.ly/ggplot2/maps-sf/).

### Reference

See [https://plot.ly/r/reference/](https://plot.ly/r/reference/) for more information and chart attribute options! If you would like to read more on visualizing geo-spatial data with sf and plotly click [here](https://blog.cpsievert.me/2018/03/30/visualizing-geo-spatial-data-with-sf-and-plotly/).