---
description: How to make maps with Plotly and simple features.
display_as: maps
language: r
layout: base
name: Simple Features
order: 11
output:
  html_document:
    keep_md: true
permalink: r/maps-sf/
thumbnail: thumbnail/sf.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

### New to Plotly?

Plotly's R library is free and open source!<br>
[Get started](https://plot.ly/r/getting-started/) by downloading the client and [reading the primer](https://plot.ly/r/getting-started/).<br>
You can set up Plotly to work in [online](https://plot.ly/r/getting-started/#hosting-graphs-in-your-online-plotly-account) or [offline](https://plot.ly/r/offline/) mode.<br>
We also have a quick-reference [cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf) (new!) to help you get started!

### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.

```{r}
library(plotly)
packageVersion('plotly')
```

### Introduction

In order to complete the examples below, you'll require installing additional packages (`install.packages("packageName")`):
- [sf](https://github.com/r-spatial/sf)

The examples below use the library [simple features](https://r-spatial.github.io/sf/) to read in the shape files before plotting the features with Plotly.

### Mapbox Access Tokens and When You Need Them

  The word "mapbox" in the trace names and `layout.mapbox` refers to the Mapbox.js open-source library. If your basemap in `layout.mapbox.style` uses data from the Mapbox *service*, then you will need to register for a free account at https://mapbox.com/ and obtain a Mapbox Access token. This token should be provided either in `mapboxAccessToken`  in `setPlotConfig` function, or as a variable that would be passed as an argument of `newPlot`.
    If your `layout.mapbox.style` does not use data from the Mapbox service, you do *not* need to register for a Mapbox account.
    <h6>Base Maps in `layout.mapbox.style`</h6>
    The accepted values for `layout.mapbox.style` are one of the following tiles.
    <ol>
        <li> `"white-bg"` yields an empty white canvas which results in no external HTTP requests </li>
        <li> `"open-street-map"`, `"carto-positron"`, `"carto-darkmatter"`, `"stamen-terrain"`, `"stamen-toner"` or `"stamen-watercolor"` yeild maps composed of *raster* tiles from various public tile servers which do not require signups or access tokens </li>
        <li> `"basic"`, `"streets"`, `"outdoors"`, `"light"`, `"dark"`, `"satellite"`, or `"satellite-streets"` yeild maps composed of *vector* tiles from the Mapbox service, and *do* require a Mapbox Access Token or an on-premise Mapbox installation. </li>
        <li> A Mapbox service style URL, which requires a Mapbox Access Token or an on-premise Mapbox installation. </li>
        <li> A Mapbox Style object as defined at https://docs.mapbox.com/mapbox-gl-js/style-spec/ </li>
    </ol>

### Basic sf 

``` {r}
library(plotly)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

p <- plot_ly(nc)

p
```

You can also use `plot_geo`:

``` {r}
library(plotly)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

p <- plot_geo(nc)

p
```

Or `plot_mapbox`:

``` {r}
library(plotly)
library(sf)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token

# Setting mapbox token for R environment 
Sys.setenv("MAPBOX_TOKEN" = mapboxToken)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

p <- plot_mapbox(nc, split=~NAME) %>%
  layout(
    mapbox = list(
      zoom = 6
    )
  ) %>%
  config(mapboxAccessToken = mapboxToken)

p
```

### Using ggplot2

Alternatively, if working in ggplot2 you can use [`ggplotly`](https://plot.ly/ggplot2/maps-sf/).

### Reference

See [https://plot.ly/r/reference/](https://plot.ly/r/reference/) for more information and chart attribute options! If you would like to read more on visualizing geo-spatial data with sf and plotly click [here](https://blog.cpsievert.me/2018/03/30/visualizing-geo-spatial-data-with-sf-and-plotly/).