---
description: How to create streamtube plots with Plotly.
display_as: 3d_charts
language: r
layout: base
name: 3D Streamtube Plots
order: 7
output:
  html_document:
    keep_md: true
permalink: r/streamtube-plot/
thumbnail: thumbnail/streamtube.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Introduction

In streamtube plots, attributes inlcude `x`, `y`, and `z`, which set the coorindates of the vector field, and `u`, `v`, and `w`, which sets the x, y, and z components of the vector field. Additionally, you can use `starts` to determine the streamtube's starting position. Lastly, `maxdisplayed` determines the maximum segments displayed in a streamtube.  

#### Basic Streamtube Plot

```{r}
library(plotly)

df = read.csv('https://raw.githubusercontent.com/plotly/datasets/master/streamtube-basic.csv')

fig <-  df 
fig <- fig %>%
  plot_ly(
    type = 'streamtube',
    x = ~x,
    y = ~y,
    z = ~z,
    u = ~u,
    v = ~v,
    w = ~w,
    sizeref = 0.5,
    cmin = 0,
    cmax = 3
  ) 
fig <- fig %>%
  layout(
    scene = list(
      camera = list(
        eye = list(
          x = -0.7243612458865182,
          y = 1.9269804254717962,
          z = 0.6704828299861716
        )
      )
    )
  )

fig
```

#### Starting Position and Segments

```{r}
library(plotly)

df = read.csv('https://raw.githubusercontent.com/plotly/datasets/master/streamtube-wind.csv')

fig <-  df 
fig <- fig %>%
  plot_ly(
    type = 'streamtube',
    x = ~x,
    y = ~y,
    z = ~z,
    u = ~u,
    v = ~v,
    w = ~w,
    starts = list(
      x = rep(80, 16),
      y = rep(c(20,30,40,50), 4),
      z = c(rep(0,4),rep(5,4),rep(10,4),rep(15,4))
    ),
    sizeref = 0.3,
    showscale = F,
    maxdisplayed = 3000
  ) 
fig <- fig %>%
  layout(
    scene = list(
      aspectratio = list(
        x = 2,
        y = 1,
        z = 0.3
      )
    ),
    margin = list(
      t = 20, b = 20, l = 20, r = 20
    )
  )

fig
```


#### Reference

See our [reference page](https://plot.ly/r/reference/) for more information and chart attribute options!.