---
description: How to use hover text and formatting in R with Plotly.
display_as: file_settings
language: r
layout: base
name: Hover Text and Formatting
order: 6
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/hover-text-and-formatting/
thumbnail: thumbnail/hover-text.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
#### Add Hover Text

```{r}
library(plotly)

fig <- plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = c(1:5), 
    y = rnorm(5, mean = 5),
    text = c("Text A", "Text B", "Text C", "Text D", "Text E"),
    hoverinfo = 'text',
    marker = list(color='green'),
    showlegend = F
  )

fig
```

#### Unified hovermode

If you set the `hovermode` attribute of your figure's layout to `x unified` or `y unified`), a single hover label will appear, describing one point per trace, for points at the same x (or y) value as the cursor. 

If multiple points in a given trace exist at the same coordinate, only one will get an entry in the hover label. In the line plot below we have forced markers to appear, to make it clearer what can be hovered over:

```{r}
library(plotly)

trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x, y = ~trace_0, name = 'trace 0', type = 'scatter', mode = 'lines') 
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers') 
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')
fig <- fig %>%
  layout(hovermode = "x unified")
fig
```

#### Format Hover Text 

```{r}
library(plotly)

fig <- plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = c(1:100), 
    y = rnorm(100, mean = 5), 
    marker = list(color='green'),
    hoverinfo = 'y',
    showlegend = F
  ) 
fig <- fig %>%
  layout(
    title = list(text="Set hover text formatting<br><a href= https://github.com/d3/d3-time-format/blob/master/README.md#locale_format>https://github.com/d3/d3-time-format/blob/master/README.md#locale_format</a>",
      size = 10),
    xaxis = list(zeroline = F),
    yaxis = list(hoverformat = '.2f'))

fig
```

#### Customize Tooltip Text with a Hovertemplate

To customize the tooltip on your graph you can use [hovertemplate](https://plotly.com/r/reference/#pie-hovertemplate), which is a template string used for rendering the information that appear on hoverbox.
This template string can include `variables` in %{variable} format, `numbers` in [d3-format's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Formatting.md#d3_forma), and `date` in [d3-time-fomrat's syntax](https://github.com/d3/d3-3.x-api-reference/blob/master/Time-Formatting.md#format).
Hovertemplate customize the tooltip text vs. [texttemplate](https://plotly.com/r/reference/#pie-texttemplate) which customizes the text that appears on your chart. <br>
Set the horizontal alignment of the text within tooltip with [hoverlabel.align](https://plotly.com/r/reference/#layout-hoverlabel-align).

```{r}
library(plotly)

fig <- plot_ly() 
fig <- fig %>%
  add_trace(
    type = 'scatter',
    mode = 'lines+markers',
    x = c(1,2,3,4,5),
    y = c(2.02825,1.63728,6.83839,4.8485,4.73463),
    text = c("Text A", "Text B", "Text C", "Text D", "Text E"),
    hovertemplate = paste('<i>Price</i>: $%{y:.2f}',
                        '<br><b>X</b>: %{x}<br>',
                        '<b>%{text}</b>'),
    showlegend = FALSE
  ) 
fig <- fig %>%
  add_trace(
    type = 'scatter',
    mode = 'lines+markers',
    x = c(1,2,3,4,5),
    y = c(3.02825,2.63728,4.83839,3.8485,1.73463),
    hovertemplate = 'Price: %{y:$.2f}<extra></extra>',
    showlegend = FALSE
  )

fig
```

```{r}
library(plotly)
fig <- plot_ly() 
fig <- fig %>%
  add_trace(
    type = "pie",
    name = "",
    values = c(2, 1, 3, 2.5),
    labels = c("R", "Python", "Java Script", "Matlab"),
    text = c("textA", "TextB", "TextC", "TextD"),
    hovertemplate = "%{label}: <br>Popularity: %{percent} </br> %{text}")


fig
```

#### Advanced Hovertemplate

```{r}
library(plotly)

df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/job-automation-probability.csv')

fig <- df %>%
  plot_ly(
    type = 'scatter',
    mode = 'markers',
    x = ~prob,
    y = ~Average.annual.wage,
    marker = list(size = ~numbEmployed, sizeref = 4000, sizemode = 'area'),
    color = ~education,
    text = ~short.occupation,
    hovertemplate = paste(
      "<b>%{text}</b><br><br>",
      "%{yaxis.title.text}: %{y:$,.0f}<br>",
      "%{xaxis.title.text}: %{x:.0%}<br>",
      "Number Employed: %{marker.size:,}",
      "<extra></extra>"
      )
    ) 
fig <- fig %>%
  layout(legend = list(orientation = 'h', y = -0.3))

fig
```

#### Reference

See [https://plotly.com/r/reference/](https://plotly.com/r/reference/) for more information and options!