---
description: How to make a Mapbox Density Heatmap in R
display_as: maps
language: r
layout: base
name: Mapbox Density
order: 3
page_type: example_index
output:
  html_document:
    keep_md: true
permalink: r/mapbox-density-heatmaps/
thumbnail: thumbnail/mapbox-density.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Map Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).


#### Stamen Terrain Tile, no Token Needed

```{r}
library(plotly)

quakes = read.csv('https://raw.githubusercontent.com/plotly/datasets/master/earthquakes-23k.csv')

fig <-  quakes 
fig <- fig %>%
  plot_ly(
    type = 'densitymapbox',
    lat = ~Latitude,
    lon = ~Longitude,
    coloraxis = 'coloraxis',
    radius = 10) 
fig <- fig %>%
  layout(
    mapbox = list(
      style="stamen-terrain",
      center= list(lon=180)), coloraxis = list(colorscale = "Viridis"))

fig
```

#Reference

See [https://plot.ly/r/reference/#densitymapbox](https://plot.ly/r/reference/#densitymapbox) for more information and options!