---
name: Sunburst Charts
permalink: r/sunburst-charts/
description: How to make sunburst charts in R with Plotly.
layout: base
thumbnail: thumbnail/sunburst.gif
language: r
display_as: basic
order: 12
output:
  html_document:
    keep_md: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
### Basic Sunburst Chart

```{r}
library(plotly)

fig <-  plot_ly(
  labels = c("Eve", "Cain", "Seth", "Enos", "Noam", "Abel", "Awan", "Enoch", "Azura"),
  parents = c("", "Eve", "Eve", "Seth", "Seth", "Eve", "Eve", "Awan", "Eve"),
  values = c(10, 14, 12, 10, 2, 6, 6, 4, 4),
  type = 'sunburst'
)

fig
```

### Branchvalues
With branchvalues "total", the value of the parent represents the width of its wedge. 
In the example below, "Enoch" is 4 and "Awan" is 6 and so Enoch's width is 4/6ths of Awans.
With branchvalues "remainder", the parent's width is determined by its own value plus those
of its children. So, Enoch's width is 4/10ths of Awan's (4 / (6 + 4)).


Note that this means that the sum of the values of the children cannot exceed the
value of their parent when branchvalues "total". When branchvalues "relative" (the default), children will
not take up all of the space below their parent (unless the parent is the root and it has a value of 0).

```{r}
library(plotly)

fig <-  plot_ly(
  labels = c("Eve", "Cain", "Seth", "Enos", "Noam", "Abel", "Awan", "Enoch", "Azura"),
  parents = c("", "Eve", "Eve", "Seth", "Seth", "Eve", "Eve", "Awan", "Eve"),
  values = c(65, 14, 12, 10, 2, 6, 6, 4, 4),
  type = 'sunburst',
  branchvalues = 'total'
)

fig
```

### Sunburst with Repeated Labels

```{r}
library(plotly)

d <- data.frame(
    ids = c(
    "North America", "Europe", "Australia", "North America - Football", "Soccer",
    "North America - Rugby", "Europe - Football", "Rugby",
    "Europe - American Football","Australia - Football", "Association",
    "Australian Rules", "Autstralia - American Football", "Australia - Rugby",
    "Rugby League", "Rugby Union"
  ),
  labels = c(
    "North<br>America", "Europe", "Australia", "Football", "Soccer", "Rugby",
    "Football", "Rugby", "American<br>Football", "Football", "Association",
    "Australian<br>Rules", "American<br>Football", "Rugby", "Rugby<br>League",
    "Rugby<br>Union"
  ),
  parents = c(
    "", "", "", "North America", "North America", "North America", "Europe",
    "Europe", "Europe","Australia", "Australia - Football", "Australia - Football",
    "Australia - Football", "Australia - Football", "Australia - Rugby",
    "Australia - Rugby"
  ),
  stringsAsFactors = FALSE
)

fig <-  plot_ly(d, ids = ~ids, labels = ~labels, parents = ~parents, type = 'sunburst')

fig
```

### Subplots
In order to create sunburst chart subplots, we use the [domain](https://plot.ly/r/reference/#sunburst-domain) attribute and the layout [grid](https://plot.ly/r/reference/#layout-grid) attribute.
```{r}
library(plotly)

d1 <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/coffee-flavors.csv')
d2 <- read.csv('https://raw.githubusercontent.com/plotly/datasets/718417069ead87650b90472464c7565dc8c2cb1c/sunburst-coffee-flavors-complete.csv')
fig <-  plot_ly() 
fig <- fig %>%
  add_trace(
    ids = d1$ids,
    labels = d1$labels,
    parents = d1$parents,
    type = 'sunburst',
    maxdepth = 2,
    domain = list(column = 0)
    ) 
fig <- fig %>%
  add_trace(
    ids = d2$ids,
    labels = d2$labels,
    parents = d2$parents,
    type = 'sunburst',
    maxdepth = 3,
    domain = list(column = 1)
  ) 
fig <- fig %>%
    layout(
      grid = list(columns =2, rows = 1),
      margin = list(l = 0, r = 0, b = 0, t = 0),
      sunburstcolorway = c(
        "#636efa","#EF553B","#00cc96","#ab63fa","#19d3f3",
        "#e763fa", "#FECB52","#FFA15A","#FF6692","#B6E880"
      ),
      extendsunburstcolors = TRUE)
fig
```


#Reference

See [https://plot.ly/r/reference/#sunburst](https://plot.ly/r/reference/#sunburst) for more information and chart attribute options!
