---
description: How to draw a line on Map in R with plotly.
display_as: maps
language: r
layout: base
name: Lines on Mapbox
order: 10
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/lines-on-mapbox/
thumbnail: thumbnail/line_mapbox.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### Mapbox Access Token

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Map Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Lines on Mapbox maps using "Scattermapbox" traces

To draw a line on your map, you either can use [Scattermapbox](https://plot.ly/r/reference/#scattermapbox) or [scattergeo](https://plot.ly/r/reference/#scattergeo) trace type in plotly. This example uses scattermapbox and defines the drawing [mode](https://plot.ly/python/reference/#scattermapbox-mode) to the combination of markers and line.

```{r}
library(plotly)

fig <-  plot_ly(
  type = 'scattermapbox',
  mode = "markers+lines",
  lon = c(10, 20, 30),
  lat = c(10, 20,30),
  marker = list(size = 10)) 
fig <- fig %>%
  add_trace(
    type = 'scattermapbox',
    mode = "markers+lines",
    lon = c(-50, -60,40),
    lat = c(30, 10, -20),
    marker = list(size = 10)) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = "stamen-terrain",
      center = list(lon = 10, lat= 10),
      zoom = 1),
    margin =list(l=0,t=0,b=0,r=0))

fig
```
#Reference

See [https://plot.ly/r/reference/#scattermapbox](https://plot.ly/r/reference/#scattermapbox) for more information and options!