---
description: How to make Mapbox maps in R with various base layers, with or without
  needing a Mapbox Access token.
display_as: maps
language: r
layout: base
name: Mapbox Layers
order: 6
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/mapbox-layers/
thumbnail: thumbnail/mapbox-layers.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### How Layers Work in Mapbox Maps

If your figure contains one or more traces of type `scattermapbox`, `choroplethmapbox` or `densitymapbox`, the `layout` of your figure contains configuration information for the map itself. 

The map is composed of layers of three different types.
<ol>
  <li> `layout.mapbox.style` defines the lowest layers, also known as your "base map"</li>
  <li> the various traces in the `plot_ly` call signature are by default rendered above the base map (this can be controlled via the use of `below` attribute).</li>
  <li> the `layout.mapbox.layers` attribute accepts an array that defines layers that are by default rendered above the traces in the `plot_ly` call signature (this can be controlled via the `below` attribute).</li>
</ol>

### Mapbox Access Tokens and When You Need Them

The word "mapbox" in trace names and the `layout.mapbox` attribute refers to the Mapbox.js open-source library. If the basemap you define using `layout.mapbox.style` uses data from the Mapbox *service* that requires API authentication, then you will need to register for a free account at https://mapbox.com/ and obtain a Mapbox Access token.

If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

#### - The following values of `layout.mapbox.style` **DO** require an Access Token:
<ol>
  <li> `"basic"`, `"streets"`, `"outdoors"`, `"light"`, `"dark"`, `"satellite"`, or `"satellite-streets"` yeild maps composed of *vector* tiles from the Mapbox service, and *do* require a Mapbox Access Token or an on-premise Mapbox installation. </li>
  <li> A Mapbox service style URL, which requires a Mapbox Access Token or an on-premise Mapbox installation. </li>
  <li> A Mapbox Style object as defined at https://docs.mapbox.com/mapbox-gl-js/style-spec/ </li>
</ol>

- We recommend setting your MapBox Access Token as an environment variable in your R environment. That way, your token will not be committed to version control. See [Basemap Requiring MapBox Access Token](#basemap-requiring-mapbox-access-token)

#### - The following values of `layout.mapbox.style` **DO NOT** require an Access Token:
<ol>
  <li> `"white-bg"` yields an empty white canvas which results in no external HTTP requests </li>
  <li> `"open-street-map"`, `"carto-positron"`, `"carto-darkmatter"`, `"stamen-terrain"`, `"stamen-toner"` or `"stamen-watercolor"` yeild maps composed of *raster* tiles from various public tile servers which do not require signups or access tokens </li>
</ol>

#### Using `layout.mapbox.layers` to Specify a Base Map

If you have access to your own private tile servers, or wish to use a tile server not included in the list above, the recommended approach is to first set `layout.mapbox.style` to `"white-bg"` and then set `layout.mapbox.layers` to `below` to specify a custom base map.

> If you omit setting `layout.mapbox.layers` to `below`, when using this approach, your data will likely be hidden by fully-opaque raster tiles!

### OpenStreetMap Tiles, no Token Needed

This chart demonstrates how to use "open-street-map" tiles as the base map for a chart. A Mapbox Access Token is not required.

```{r}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <-  us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 

fig
```

### Base Map Below A Trace: No Token Needed

This chart demonstrates how to use a public USGS imagery map as a base map using `layout.mapbox.layers`. The map is rendered *below* the trace.

```{r}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <-   us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    type = "scattermapbox",
    hovertext = us_cities[,"City"],
    marker = list(color = "fuchsia")) 
fig <- fig %>%
  layout(mapbox= list(
    style = "white-bg",
    zoom = 3,
    center = list(lon = -93 ,lat= 41),
    layers = list(list(
      below = 'traces',
      sourcetype = "raster",
      source = list("https://basemap.nationalmap.gov/arcgis/rest/services/USGSImageryOnly/MapServer/tile/{z}/{y}/{x}")))))

fig
```

#### Maps Below and Above A Trace: No Token Needed

This example uses the same base map *below* the trace as the one above. 

In addition, a WMS layer map from Environment Canada (which displays near-real-time radar imagery in partly-transparent raster tiles) is rendered *above* the `scattermapbox` trace, as is the default:

```{r}
library(plotly)
fig <-  us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    type = "scattermapbox",
    hovertext = us_cities[,"City"],
    marker = list(color = "fuchsia")) 
fig <- fig %>%
  layout(
    mapbox= list(
      style = "white-bg",
      sourcetype = 'raster',
      zoom = 3,
      center = list(lon = -93 ,lat= 41),
      layers = list(list(
        below = 'traces',
        sourcetype = "raster",
        source = list("https://basemap.nationalmap.gov/arcgis/rest/services/USGSImageryOnly/MapServer/tile/{z}/{y}/{x}")),
        list(
          sourcetype = "raster",
          source = list("https://geo.weather.gc.ca/geomet/?SERVICE=WMS&VERSION=1.3.0&REQUEST=GetMap&BBOX={bbox-epsg-3857}&CRS=EPSG:3857&WIDTH=1000&HEIGHT=1000&LAYERS=RADAR_1KM_RDBR&TILED=true&FORMAT=image/png")))))

fig
```

#### Map Requiring MapBox Access Token

This example demonstrates how to use a MapBox Access token in order to use a "dark" base map from MapBox, which requires API authentication. 

In this case, the Access Token is stored in an environment variable which is then passed to the chart configuration call signature. 

```{r}
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <-  us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'dark',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 
fig <- fig %>%
  config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```

#Reference

See [https://plot.ly/r/reference/#scattermapbox](https://plot.ly/r/reference/#scattermapbox) for more information and options!
