---
description: How to make a Mapbox Choropleth Map of US Counties in R with Plotly.
display_as: maps
language: r
layout: base
name: Mapbox Choropleth Maps
order: 7
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/mapbox-county-choropleth/
redirect_from: r/county-level-choropleth/
thumbnail: thumbnail/mapbox-choropleth.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

A [Choropleth Map](https://en.wikipedia.org/wiki/Choropleth_map) is a map composed of colored polygons. It is used to represent spatial variations of a quantity. This page documents how to build **tile-map** choropleth maps, but you can also build [**outline** choropleth maps using our non-Mapbox trace types](/r/choropleth-maps).

Below we show how to create Choropleth Maps using Plotly `Choroplethmapbox` graph object.

### Mapbox Access Token and Base Map Configuration

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Map Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Introduction: main parameters for choropleth tile maps

Making choropleth Mapbox maps requires two main types of input:

1. GeoJSON-formatted geometry information where each feature has either an `id` field or some identifying value in `properties`.
2. A list of values indexed by feature identifier.

The GeoJSON data is passed to the `geojson` argument, and the data is passed into the `z` argument of `plot_geo`.

**Note** the `geojson` attribute can also be the URL to a GeoJSON file, which can speed up map rendering in certain cases.

#### GeoJSON with `feature.id`

Here we load a GeoJSON file containing the geometry information for US counties, where `feature.id` is a [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
library(rjson)
url = 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
json_file <- rjson::fromJSON(file=url)
json_file$features[[1]]$id
```

### Data indexed by `id`

Here we load unemployment data by county, also indexed by [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
df = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv", header = T, colClasses = c("fips"="character"))
head(df)
```
### Mapbox Choropleth Map Using GeoJSON

With `choroplethmapbox`, each row of the DataFrame is represented as a region of the choropleth.

```{r}
library(rjson)
library(plotly)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
df <- read.csv(url2, colClasses=c(fips="character"))
fig <- plot_ly() 
fig <- fig %>% add_trace(
    type="choroplethmapbox",
    geojson=counties,
    locations=df$fips,
    z=df$unemp,
    colorscale="Viridis",
    zmin=0,
    zmax=12,
    marker=list(line=list(
      width=0),
      opacity=0.5
    )
  )
fig <- fig %>% layout(
    mapbox=list(
      style="carto-positron",
      zoom =2,
      center=list(lon= -95.71, lat=37.09))
  )
fig
```

### Mapbox Choropleth Map Using GeoJSON with `featureidkey`
If the GeoJSON you are using either does not have an `id` field or you wish you use one of the keys in the `properties` field, you may use the `featureidkey` parameter to specify where to match the values of locations.

In the following GeoJSON object/data-file pairing, the values of `properties.district` match the values of the `district` column:
```{r}
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/election.geojson'
geojson <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/election.csv"
df <- read.csv(url2)
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig <- plot_ly() 
fig <- fig %>% add_trace(
    type="choroplethmapbox",
    geojson=geojson,
    locations=df$district,
    z=df$Bergeron,
    colorscale="Viridis",
    featureidkey="properties.district"
  )
fig <- fig %>% colorbar(title = "Bergeron Votes")
fig <- fig %>% layout(
    mapbox=list(
      style="carto-positron",
      zoom =9,
      center=list(lon=-73.7073, lat=45.5517))
  )
fig
```

#### Mapbox Light base map: free token needed

```{r}
library(rjson)
library(plotly)

mapboxToken <- paste(readLines("../.mapbox_token"), collapse="")    # You need your own token
Sys.setenv("MAPBOX_TOKEN" = mapboxToken) # for Orca

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file = url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
df <- read.csv(url2,colClasses = c(fips = "character"))

fig <- plot_ly()
fig <- fig %>% add_trace(
  type = "choroplethmapbox",
  geojson = counties,
  locations = df$fips,
  z=df$unemp,
  colorscale="Viridis",
  zmin=0,
  zmax=12,
  marker=list(line = list(
    width = 0),
    opacity=0.5
  )
)
fig <- fig %>% layout(
  mapbox = list(
    style = "light",
    zoom =3,
    center = list(lon = -95.7129, lat = 37.0902))
) 
fig <- fig %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
fig
```
#Reference

See [https://plotly.com/r/reference/#scattermapbox](https://plotly.com/r/reference/#choroplethmapbox) for more information and options!