---
description: How to make a Mapbox Choropleth Map of US Counties in R with Plotly.
display_as: maps
language: r
layout: base
name: Choropleth mapbox
order: 12
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/mapbox-county-choropleth/
thumbnail: thumbnail/mapbox-choropleth.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

A [Choropleth Map](https://en.wikipedia.org/wiki/Choropleth_map) is a map composed of colored polygons. It is used to represent spatial variations of a quantity. This page documents how to build **tile-map** choropleth maps, but you can also build [**outline** choropleth maps using our non-Mapbox trace types](/r/choropleth-maps).

Below we show how to create Choropleth Maps using Plotly `Choroplethmapbox` graph object.

### Mapbox Access Token and Base Map Configuration

To plot on Mapbox maps with Plotly you *may* need a Mapbox account and a public [Mapbox Access Token](https://www.mapbox.com/studio). See our [Mapbox Map Layers](/r/mapbox-layers/) documentation for more information. If you're using a Chart Studio Enterprise server, please see additional instructions [here](https://help.plot.ly/mapbox-atlas).

### Introduction: main parameters for choropleth tile maps

Making choropleth Mapbox maps requires two main types of input:

1. GeoJSON-formatted geometry information where each feature has either an `id` field or some identifying value in `properties`.
2. A list of values indexed by feature identifier.

The GeoJSON data is passed to the `geojson` argument, and the data is passed into the `color` argument of `px.choropleth_mapbox` (`z` if using `graph_objects`), in the same order as the IDs are passed into the `location` argument.

**Note** the `geojson` attribute can also be the URL to a GeoJSON file, which can speed up map rendering in certain cases.

#### GeoJSON with `feature.id`

Here we load a GeoJSON file containing the geometry information for US counties, where `feature.id` is a [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
library(jsonlite)
json_file <- fromJSON('https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json')
counties <- as.data.frame(json_file)
counties <- counties[,-1]
a <- counties[1,]
```

### Data indexed by `id`

Here we load unemployment data by county, also indexed by [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).

```{r}
df = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv", header = T, colClasses = c("fips"="character"))
head(df)
```

### Choropleth map using carto base map (no token needed)

With `choroplethmapbox`, each row of the DataFrame is represented as a region of the choropleth.

```{r}
library(rjson)
library(plotly)
library(data.table)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file = url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
mydat <- fread(url2,colClasses = c(fips = "character"))
fig <- plot_ly() %>%
  add_trace(
    type = "choroplethmapbox",
    mydat,
    geojson = counties,
    locations = mydat$fips,
    z=mydat$unemp,
    colorscale="Viridis",
    zmin=0,
    zmax=12,
    marker=list(line = list(
      width = 0),
      opacity=0.5
    )
  ) %>%
  layout(
    mapbox = list(
      style = "carto-positron",
      zoom =3,
      center = list(lon = -95.7129, lat = 37.0902))
  )
fig
```
### Indexing by GeoJSON Properties

If the GeoJSON you are using either does not have an `id` field or you wish you use one of the keys in the `properties` field, you may use the `featureidkey` parameter to specify where to match the values of `locations`.

In the following GeoJSON object/data-file pairing, the values of `properties.district` match the values of the `district` column:

```{r}

```

#Reference

See [https://plot.ly/r/reference/#scattermapbox](https://plot.ly/r/reference/#scattermapbox) for more information and options!