---
name: Imshow
description: How to display image data in Python with R.
display_as: scientific
layout: base
language: r
order: 15
output:
  html_document:
    keep_md: true
permalink: r/imshow/
thumbnail: thumbnail/imshow.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
This tutorial shows how to display and explore image data. If you would like instead a logo or static image, use `layout.Image` as explained [here](https://plot.ly/r/logos/).

### Display RGB Image Data with Image Trace

Note that `Image` trace only accepts multichannel images. For single images, use [`Heatmap`](https://plot.ly/r/heatmaps/). `Image` trace is different from the `layout.Image` class, which can be used for adding background images or logos to figures.
```{r}
library(plotly)

img_rgb = list(list(c(255, 0, 0),c(0, 255, 0),c(0, 0, 255)),
               list(c(0,255, 0),c(0, 0, 255),c(255, 0, 0)))
fig <- plot_ly(type="image", z=img_rgb)
fig
```
### Read image arrays from image files

In order to create a numerical array to be passed to `Image` trace, you can use a third-party library like [EBImage](https://www.rdocumentation.org/packages/EBImage/versions/4.14.2) to open an image from a URL.

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')

fig <- plot_ly(type="image", z=img*255) 
fig
```
### Define the data range covered by the color range with zmin and zmax

The data range and color range are mapped together using the parameters `zmin` and `zmax`, which correspond respectively to the data values mapped to black `[0, 0, 0]` and white `[255, 255, 255]`.
The default value of `zmin` and [zmax](https://plot.ly/r/reference/#image-zmax) depends on the `colormodal` value. In this example `colormodel is "rgb"`(by default), so the default value of `zmin is [0, 0, 0]` and `zmax is [250, 250, 250]`.

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')
# Stretch the contrast of the red channel only, resulting in a more red image
fig <- plot_ly(type="image", z=img*250, 
               zmin=c(10, 0, 0), zmax=c(200, 250, 250))
fig
```
### Set Ticks and Margins 

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')
fig <- plot_ly(type="image", z=img*250)
fig <- fig %>% layout(margin=list(l=10, r=10, b=0, t=0),
                      xaxis=list(showticklabels=FALSE, ticks=""),
                      yaxis=list(showticklabels=FALSE, ticks=""))
fig
```

### Combine image charts and other traces

```{r}
library(plotly)
library(EBImage)
img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')

fig <- plot_ly(type="image", z=img*250)
fig <- fig %>% add_trace(
  type='bar', y=c(50, 60), x=c(40, 50), 
  marker=list(color='pink', size=10))
fig
```
### Reference
See [https://plot.ly/r/reference/#image](https://plot.ly/r/reference/#area) for more information and chart attribute options!
