---
description: How to display image data with R.
display_as: scientific
language: r
layout: base
name: Display Image Data
order: 14
output:
  html_document:
    keep_md: true
permalink: r/displaying-images/
redirect_from: r/imshow/
thumbnail: thumbnail/imshow.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
This tutorial shows how to display and explore image data. If you would like instead a logo or static image, use `layout.Image` as explained [here](https://plotly.com/r/logos/).

### Display RGB Image Data with Image Trace

Note that `Image` trace only accepts multichannel images. For single images, use [`Heatmap`](https://plotly.com/r/heatmaps/). `Image` trace is different from the `layout.Image` class, which can be used for adding background images or logos to figures.
```{r}
library(plotly)
library(EBImage)

img_rgb = list(list(c(255, 0, 0),c(0, 255, 0),c(0, 0, 255)),
               list(c(0,255, 0),c(0, 0, 255),c(255, 0, 0)))
fig <- plot_ly(type="image", z=img_rgb)
fig
```
### Read image arrays from image files

In order to create a numerical array to be passed to `Image` trace, you can use a third-party library like [EBImage](https://www.rdocumentation.org/packages/EBImage/versions/4.14.2) to open an image from a URL.

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')

fig <- plot_ly(type="image", z=img*255)
fig
```
### Define the data range covered by the color range with zmin and zmax

The data range and color range are mapped together using the parameters `zmin` and `zmax`, which correspond respectively to the data values mapped to black `[0, 0, 0]` and white `[255, 255, 255]`.
The default value of `zmin` and [zmax](https://plotly.com/r/reference/#image-zmax) depends on the `colormodal` value. In this example `colormodel is "rgb"`(by default), so the default value of `zmin is [0, 0, 0]` and `zmax is [250, 250, 250]`.

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')
# Stretch the contrast of the red channel only, resulting in a more red image
fig <- plot_ly(type="image", z=img*250,
               zmin=c(10, 0, 0), zmax=c(200, 250, 250))
fig
```
### Set Ticks and Margins

```{r}
library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')
fig <- plot_ly(type="image", z=img*250)
fig <- fig %>% layout(margin=list(l=10, r=10, b=0, t=0),
                      xaxis=list(showticklabels=FALSE, ticks=""),
                      yaxis=list(showticklabels=FALSE, ticks=""))
fig
```

### Drawing Shapes on Images

You can create layout shapes programatically, but you can also draw shapes manually by setting the `dragmode` to one of the shape-drawing modes: `drawline`,`drawopenpath`, `drawclosedpath`, `drawcircle`, or `drawrect`. If you need to switch between different shape-drawing or other dragmodes (panning, selecting, etc.), modebar buttons can be added in the `config` of your figure to select the dragmode.

If you switch to a different dragmode such as pan or zoom, you will need to select the drawing tool in the modebar to go back to shape drawing.

This shape-drawing feature is particularly interesting for annotating graphs, in particular [image traces](https://plotly.com/r/displaying-images/).

Once you have drawn shapes, you can select and modify an existing shape by clicking on its boundary (note the arrow pointer). Its fillcolor turns to pink to highlight the activated shape and then you can

- drag and resize it for lines, rectangles and circles/ellipses
- drag and move individual vertices for closed paths
- move individual vertices for open paths.

An activated shape is deleted by cliking on the `eraseshape` button.

Drawing or modifying a shape triggers a `relayout` event, which [can be captured by a callback inside a Dash For R application](https://dash.plotly.com/interactive-graphing).

```{r}
library(plotly)

library(plotly)
library(EBImage)

img = readImage('https://upload.wikimedia.org/wikipedia/commons/thumb/0/00/Crab_Nebula.jpg/240px-Crab_Nebula.jpg')

fig <- plot_ly(type="image", z=img*255)

fig <- layout(fig, dragmode="drawrect", xaxis = list(title = 'Click and drag inside the figure to draw a rectangle or select another shape in the modebar'))

fig <- fig %>%
  config(modeBarButtonsToAdd = list("drawine", "drawopenpath", "drawclosedpath", "drawcircle", "drawrect", "eraseshape" ) )

fig

```

### Reference
See [https://plotly.com/r/reference/#image](https://plotly.com/r/reference/#area) for more information and chart attribute options!
