---
description: How to export R graphs as static images.
display_as: file_settings
language: r
layout: base
name: Exporting Graphs as Static Images
order: 21
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/image-export-orca/
thumbnail: thumbnail/sizing.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### New to Plotly?

Plotly's R library is free and open source!<br>
[Get started](https://plot.ly/r/getting-started/) by downloading the client and [reading the primer](https://plot.ly/r/getting-started/).<br>
You can set up Plotly to work in [online](https://plot.ly/r/getting-started/#hosting-graphs-in-your-online-plotly-account) or [offline](https://plot.ly/r/offline/) mode.<br>
We also have a quick-reference [cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf) (new!) to help you get started!

### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.
```{r}
library(plotly)
packageVersion('plotly')
```

### Export R Graphs As Static Images Using `orca()`

[Orca](https://github.com/plotly/orca) is an open source command line tool for generating static images of graphs created with [Plotly's graphing libraries](https://plot.ly/graphing-libraries). 

To use Orca to export static images of the graphs you create with the `plotly` R package, you can use the built-in `orca()` funciton in versions `4.7.900` and above.

In order to use the `orca()` function, you need to have the [`processx`](https://github.com/r-lib/processx) R package installed. 

The `orca()` function accepts two parameters. The frist is the plot to be exported and second is the filename. 

For example, the running the following commands in an R session would export the graph stored in `p` in a file called `surface-plot.svg`:

```{r, eval = FALSE}
if (!require("processx")) install.packages("processx")
p <- plot_ly(z = ~volcano) %>% add_surface()
orca(p, "surface-plot.svg")
```
