---
description: How to export R graphs as static images.
display_as: file_settings
language: r
layout: base
name: Exporting Graphs as Static Images
order: 21
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/static-image-export/
thumbnail: thumbnail/sizing.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
### New to Plotly?

Plotly's R library is free and open source!<br>
[Get started](https://plot.ly/r/getting-started/) by downloading the client and [reading the primer](https://plot.ly/r/getting-started/).<br>
You can set up Plotly to work in [online](https://plot.ly/r/getting-started/#hosting-graphs-in-your-online-plotly-account) or [offline](https://plot.ly/r/offline/) mode.<br>
We also have a quick-reference [cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf) (new!) to help you get started!

### Version Check

Version 4 of Plotly's R package is now [available](https://plot.ly/r/getting-started/#installation)!<br>
Check out [this post](http://moderndata.plot.ly/upgrading-to-plotly-4-0-and-above/) for more information on breaking changes and new features available in this version.
```{r}
library(plotly)
packageVersion('plotly')
```
### Supported File Formats

With the `plotly` R package, you can export graphs you create as static images in the `.png`, `.jpg`/`.jpeg`, `.eps`, `.svg`, and/or `.pdf` formats using[Orca](https://github.com/plotly/orca), an open source command line tool for generating static images of graphs created with [Plotly's graphing libraries](https://plot.ly/graphing-libraries).  

**Note:** It is important to be aware that R graphs containing WebGL-based traces (i.e. of type `scattergl`, `heatmapgl`, `contourgl`, `scatter3d`, `surface`, `mesh3d`, `scatterpolargl`, `cone`, `streamtube`, `splom`, and/or `parcoords`) will include encapsulated rasters instead of vectors for some parts of the image if they are exported as static images in a vector format like `.eps`, `.svg`, and/or `.pdf`.

**Note** Orca runs entirely locally and does not require internet access. No network requests are made to the Chart Studio or any other web service when you invoke the `orca()` function to export static images in your R session. 

### Install Orca

Please follow the installation instructions which can be found on [Orca's GitHub Page](https://github.com/plotly/orca#installation).

### Export R Graphs As Static Images Using `orca()`

To use Orca to export static images of the graphs you create with the `plotly` R package, you can use the built-in `orca()` function in versions `4.7.900` and above. 

You need to have the [`processx`](https://github.com/r-lib/processx) R package installed as well. 

The `orca()` function accepts two parameters. The first is the plot to be exported and second is the filename. 

For example, running the following commands in an R session would export the graph stored in `p` in a file called `surface-plot.svg`:

```{r, eval = FALSE}
if (!require("processx")) install.packages("processx")
p <- plot_ly(z = ~volcano) %>% add_surface()
orca(p, "surface-plot.svg")
```
