---
description: Visualize t-SNE and UMAP in R with Plotly.
display_as: ai_ml
language: r
layout: base
name: t-SNE and UMAP projections
order: 5
output:
  html_document:
    keep_md: true
permalink: r/t-sne-and-umap-projections/
thumbnail: thumbnail/tsne-umap-projections.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
## t-SNE and UMAP projections in R


This page presents various ways to visualize two popular dimensionality reduction techniques, namely the [t-distributed stochastic neighbor embedding](https://lvdmaaten.github.io/tsne/) (t-SNE) and [Uniform Manifold Approximation and Projection](https://umap-learn.readthedocs.io/en/latest/index.html) (UMAP). They are needed whenever you want to visualize data with more than two or three features (i.e. dimensions).

We first show how to visualize data with more than three features using the [scatter plot matrix](https://plotly.com/r/splom/#:~:text=The%20Plotly%20splom%20trace%20implementation,array%2Fvariable%20represents%20a%20dimension), then we apply dimensionality reduction techniques to get 2D/3D representation of our data, and visualize the results with [scatter plots](https://plotly.com/r/line-and-scatter/) and [3D scatter plots](https://plotly.com/r/3d-scatter-plots/).


## Basic t-SNE projections

t-SNE is a popular dimensionality reduction algorithm that arises from probability theory. Simply put, it projects the high-dimensional data points (sometimes with hundreds of features) into 2D/3D by inducing the projected data to have a similar distribution as the original data points by minimizing something called the [KL divergence](https://towardsdatascience.com/light-on-math-machine-learning-intuitive-guide-to-understanding-kl-divergence-2b382ca2b2a8).

Compared to a method like Principal Component Analysis (PCA), it takes significantly more time to converge, but present significantly better insights when visualized. For example, by projecting features of flowers, it will be able to distinctly group 



### Visualizing high-dimensional data with `splom`

First, let's try to visualize every feature of the [Iris dataset](https://archive.ics.uci.edu/ml/datasets/iris), and color everything by the species. We will use the Scatter Plot Matrix ([splom](https://plotly.com/r/splom/#:~:text=The%20Plotly%20splom%20trace%20implementation,array%2Fvariable%20represents%20a%20dimension)), which lets us plot each feature against everything else, which is convenient when your dataset has more than 3 dimensions.

```{r}
library(plotly) 
library(stats) 
data(iris) 
X <- subset(iris, select = -c(Species)) 
axis = list(showline=FALSE, 
            zeroline=FALSE, 
            gridcolor='#ffff', 
            ticklen=4)
fig <- iris %>%  
  plot_ly()  %>%  
  add_trace(  
    type = 'splom',  
    dimensions = list( 
      list(label = 'sepal_width',values=~Sepal.Width),  
      list(label = 'sepal_length',values=~Sepal.Length),  
      list(label ='petal_width',values=~Petal.Width),  
      list(label = 'petal_length',values=~Petal.Length)),  
    color = ~Species, colors = c('#636EFA','#EF553B','#00CC96') 
  ) 
fig <- fig %>% 
  layout( 
    legend=list(title=list(text='species')), 
    hovermode='closest', 
    dragmode= 'select', 
    plot_bgcolor='rgba(240,240,240,0.95)', 
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    xaxis2=axis, 
    xaxis3=axis, 
    xaxis4=axis, 
    yaxis2=axis, 
    yaxis3=axis, 
    yaxis4=axis 
  ) 
fig

```

### Project data into 2D with t-SNE and `px.scatter`

Now, let's use the t-SNE algorithm to project the data shown above into two dimensions. Notice how each of the species is physically separate from each other.

```{r}
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

set.seed(0)
tsne <- tsne(features, initial_dims = 2)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, type = 'scatter', mode = 'markers', split = ~iris$Species)

fig <- fig %>%
  layout(
    plot_bgcolor = "#e5ecf6"
  )

fig

```

### Project data into 3D with t-SNE and `px.scatter_3d`

t-SNE can reduce your data to any number of dimensions you want! Here, we show you how to project it to 3D and visualize with a 3D scatter plot.

```{r}
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

#set.seed(0)
tsne <- tsne(features, initial_dims = 3, k =3)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96') ) %>% 
  add_markers(size = 8) %>%
  layout( 
    xaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'), 
    yaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'),
    scene =list(bgcolor = "#e5ecf6"))
fig

```

## Projections with UMAP

Just like t-SNE, [UMAP](https://umap-learn.readthedocs.io/en/latest/index.html) is a dimensionality reduction specifically designed for visualizing complex data in low dimensions (2D or 3D). As the number of data points increase, UMAP becomes more time efficient compared to TSNE.

In the example below, we see how easy it is to use UMAP in R.

```{r}

library(plotly) 
library(umap) 
iris.data = iris[, grep("Sepal|Petal", colnames(iris))] 
iris.labels = iris[, "Species"] 
iris.umap = umap(iris.data, n_components = 2, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 
 
fig <- plot_ly(final, x = ~X1, y = ~X2, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96'), type = 'scatter', mode = 'markers')%>%  
  layout(
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='species')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 
 
iris.umap = umap(iris.data, n_components = 3, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 
 
fig2 <- plot_ly(final, x = ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96')) 
fig2 <- fig2 %>% add_markers() 
fig2 <- fig2 %>% layout(scene = list(xaxis = list(title = '0'), 
                                     yaxis = list(title = '1'), 
                                     zaxis = list(title = '2'))) 
 
fig 
fig2
```

## Visualizing image datasets

In the following example, we show how to visualize large image datasets using UMAP. 

Although there's over 1000 data points, and many more dimensions than the previous example, it is still extremely fast. This is because UMAP is optimized for speed, both from a theoretical perspective, and in the way it is implemented. Learn more in [this comparison post](https://umap-learn.readthedocs.io/en/latest/benchmarking.html).

```{r}
library(rsvd) 
library(plotly) 
library(umap) 
data('digits') 
digits.data = digits[, grep("pixel", colnames(digits))] 
digits.labels = digits[, "label"] 
digits.umap = umap(digits.data, n_components = 2, k = 10) 
layout <- digits.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, digits[,'label']) 
colnames(final) <- c('X1', 'X2', 'label') 

fig <- plot_ly(final, x = ~X1, y = ~X2, split = ~label,  type = 'scatter', mode = 'markers')%>%  
  layout(  
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='digit')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 
fig

```

<!-- #region -->
## Reference

Plotly figures:
* https://plotly.com/r/line-and-scatter/

* https://plotly.com/r/3d-scatter-plots/

* https://plotly.com/r/splom/


Details about algorithms:
* UMAP library: https://umap-learn.readthedocs.io/en/latest/

* t-SNE User guide: https://cran.r-project.org/web/packages/tsne/tsne.pdf

* t-SNE paper: https://www.jmlr.org/papers/volume9/vandermaaten08a/vandermaaten08a.pdf

* MNIST: http://yann.lecun.com/exdb/mnist/
<!-- #endregion -->
