
## ROC and PR Curves in R

Interpret the results of your classification using Receiver Operating Characteristics (ROC) and Precision-Recall (PR) Curves in R with Plotly.

## Preliminary plots

Before diving into the receiver operating characteristic (ROC) curve, we will look at two plots that will give some context to the thresholds mechanism behind the ROC and PR curves.

In the histogram, we observe that the score spread such that most of the positive labels are binned near 1, and a lot of the negative labels are close to 0. When we set a threshold on the score, all of the bins to its left will be classified as 0's, and everything to the right will be 1's. There are obviously a few outliers, such as **negative** samples that our model gave a high score, and *positive* samples with a low score. If we set a threshold right in the middle, those outliers will respectively become **false positives** and *false negatives*.

As we adjust thresholds, the number of positive positives will increase or decrease, and at the same time the number of true positives will also change; this is shown in the second plot. As you can see, the model seems to perform fairly well, because the true positive rate and the false positive rate decreases sharply as we increase the threshold. Those two lines each represent a dimension of the ROC curve.

```{r}
library(plotly) 
library(tidymodels) 
set.seed(0) 
X <- matrix(rnorm(10000),nrow=500) 
y <- sample(0:1, 500, replace=TRUE) 
data <- data.frame(X,y) 
data$y <- as.factor(data$y) 
X <- subset(data,select = -c(y)) 
logistic_glm <- 
  logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification") %>% 
  fit(y ~ ., data = data) 
 
y_scores <- logistic_glm %>% 
  predict(X, type = 'prob') 
 
y_score <- y_scores$.pred_1 
db <- data.frame(data$y, y_score) 
 
z <- roc_curve(data = db, 'data.y', 'y_score') 
z$specificity <- 1 - z$specificity 
colnames(z) <- c('threshold', 'tpr', 'fpr') 
 
fig1 <- plot_ly(x= y_score, color = data$y, colors = c('blue', 'red'), type = 'histogram', alpha = 0.5, nbinsx = 50) %>%  
  layout(barmode = "overlay") 
fig1 
 
fig2 <- plot_ly(data = z, x = ~threshold) %>%  
  add_trace(y = ~fpr, mode = 'lines', name = 'false positive rate', type = 'scatter')%>%  
  add_trace(y = ~tpr, mode = 'lines', name = 'true positive rate', type = 'scatter')%>%  
  layout(title = 'TPR and FPR at every threshold')
fig2
```

## Basic binary ROC curve

We  display the area under the ROC curve (ROC AUC), which is fairly high, thus consistent with our interpretation of the previous plots.

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(pROC)

set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
db <- data.frame(X,y)
db$y <- as.factor(db$y)
test_data = db[1:20]

model<- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  # Fit the model
  fit(y ~., data = db)


ypred <- predict(model,
                  new_data = test_data,
                  type = "prob")

yscore <- data.frame(ypred$.pred_0)
rdb <- cbind(db$y,yscore)
colnames(rdb) = c('y','yscore')


pdb <- roc_curve(rdb, y, yscore)
pdb$specificity <- 1 - pdb$specificity 
auc = roc_auc(rdb, y, yscore)
auc = auc$.estimate

tit = paste('ROC Curve (Auc = ',toString(round(auc,2)),')',sep = '')

fig <-  plot_ly(data = pdb ,x =  ~specificity, y = ~sensitivity, type = 'scatter', mode = 'lines', fill = 'tozeroy') %>%
  layout(title = tit,xaxis = list(title = "False Positive Rate"), yaxis = list(title = "True Positive Rate")) %>%
add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'),inherit = FALSE, showlegend = FALSE)
fig
```



## Multiclass ROC Curve

When you have more than 2 classes, you will need to plot the ROC curve for each class separately. Make sure that you use a [one-versus-rest](https://cran.r-project.org/web/packages/multiclassPairs/vignettes/Tutorial.html) model, or make sure that your problem has a multi-label format; otherwise, your ROC curve might not return the expected results.
                                                                                                                   
```{r}
library(plotly) 
library(tidymodels) 
library(fastDummies) 
 
data(iris) 
ind <- sample.int(150, 50) 
samples <- sample(x = iris$Species, size = 50) 
iris[ind,'Species'] = samples 
 
X <- subset(iris, select = -c(Species)) 
iris$Species <- as.factor(iris$Species) 
 
logistic <- 
  multinom_reg() %>% 
  set_engine("nnet") %>% 
  set_mode("classification") %>% 
  fit(Species ~ ., data = iris) 
 
y_scores <- logistic %>% 
  predict(X, type = 'prob') 
 
y_onehot <- dummy_cols(iris$Species) 
colnames(y_onehot) <- c('drop', 'setosa', 'versicolor', 'virginica') 
y_onehot <- subset(y_onehot, select = -c(drop)) 
 
z = cbind(y_scores, y_onehot) 
 
z$setosa <- as.factor(z$setosa) 
roc_setosa <- roc_curve(data = z, setosa, .pred_setosa) 
roc_setosa$specificity <- 1 - roc_setosa$specificity 
colnames(roc_setosa) <- c('threshold', 'tpr', 'fpr') 
auc_setosa <- roc_auc(data = z, setosa, .pred_setosa) 
auc_setosa <- auc_setosa$.estimate 
setosa <- paste('setosa (AUC=',toString(round(1-auc_setosa,2)),')',sep = '') 
 
z$versicolor <- as.factor(z$versicolor) 
roc_versicolor <- roc_curve(data = z, versicolor, .pred_versicolor) 
roc_versicolor$specificity <- 1 - roc_versicolor$specificity 
colnames(roc_versicolor) <- c('threshold', 'tpr', 'fpr') 
auc_versicolor <- roc_auc(data = z, versicolor, .pred_versicolor) 
auc_versicolor <- auc_versicolor$.estimate 
versicolor <- paste('versicolor (AUC=',toString(round(1-auc_versicolor,2)),')', sep = '') 
 
z$virginica <- as.factor(z$virginica) 
roc_virginica <- roc_curve(data = z, virginica, .pred_virginica) 
roc_virginica$specificity <- 1 - roc_virginica$specificity 
colnames(roc_virginica) <- c('threshold', 'tpr', 'fpr') 
auc_virginica <- roc_auc(data = z, virginica, .pred_virginica) 
auc_virginica <- auc_virginica$.estimate 
virginica <- paste('virginica (AUC=',toString(round(1-auc_virginica,2)),')',sep = '') 
 
fig <- plot_ly()%>% 
  add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'), showlegend = FALSE) %>%  
  add_trace(data = roc_setosa,x = ~fpr, y = ~tpr, mode = 'lines', name = setosa, type = 'scatter')%>%  
  add_trace(data = roc_versicolor,x = ~fpr, y = ~tpr, mode = 'lines', name = versicolor, type = 'scatter')%>%  
  add_trace(data = roc_virginica,x = ~fpr, y = ~tpr, mode = 'lines', name = virginica, type = 'scatter')%>%  
  layout(xaxis = list( 
    title = "False Positive Rate" 
  ), yaxis = list( 
    title = "True Positive Rate" 
  ),legend = list(x = 100, y = 0.5)) 
fig

```
                                                                                                                  
                                                                                                                   
## Precision-Recall Curves                                                                                        

Plotting the PR curve is very similar to plotting the ROC curve. The following examples are slightly modified from the previous examples:                                                                                                            

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(pROC)

set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
db <- data.frame(X,y)
db$y <- as.factor(db$y)
test_data = db[1:20]

model<- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  # Fit the model
  fit(y ~., data = db)

ypred <- predict(model,
                 new_data = test_data,
                 type = "prob")

yscore <- data.frame(ypred$.pred_0)
rdb <- cbind(db$y,yscore)
colnames(rdb) = c('y','yscore')

pdb <- pr_curve(rdb, y, yscore)
auc = roc_auc(rdb, y, yscore)
auc = auc$.estimate

tit = paste('ROC Curve (Auc = ',toString(round(auc,2)),')',sep = '')

fig <-  plot_ly(data = pdb ,x =  ~recall, y = ~precision, type = 'scatter', mode = 'lines', fill = 'tozeroy') %>%
  add_segments(x = 0, xend = 1, y = 1, yend = 0, line = list(dash = "dash", color = 'black'),inherit = FALSE, showlegend = FALSE) %>%
  layout(title = tit, xaxis = list(title = "Recall"), yaxis = list(title = "Precision") ) 

fig
```
                                                                                                                   
In this example, we use the average precision metric, which is an alternative scoring method to the area under the PR curve.

```{r}
library(plotly)
library(tidymodels)
library(fastDummies)

data(iris)
ind <- sample.int(150, 50)
samples <- sample(x = iris$Species, size = 50)
iris[ind,'Species'] = samples

X <- subset(iris, select = -c(Species))
iris$Species <- as.factor(iris$Species)

logistic <-
  multinom_reg() %>%
  set_engine("nnet") %>%
  set_mode("classification") %>%
  fit(Species ~ ., data = iris)

y_scores <- logistic %>%
  predict(X, type = 'prob')

y_onehot <- dummy_cols(iris$Species)
colnames(y_onehot) <- c('drop', 'setosa', 'versicolor', 'virginica')
y_onehot <- subset(y_onehot, select = -c(drop))

z = cbind(y_scores, y_onehot)

z$setosa <- as.factor(z$setosa)
pr_setosa <- pr_curve(data = z, setosa, .pred_setosa)
aps_setosa <- mean(pr_setosa$precision)
setosa <- paste('setosa (AP =',toString(round(aps_setosa,2)),')',sep = '')


z$versicolor <- as.factor(z$versicolor)
pr_versicolor <- pr_curve(data = z, versicolor, .pred_versicolor)
aps_versicolor <- mean(pr_versicolor$precision)
versicolor <- paste('versicolor (AP = ',toString(round(aps_versicolor,2)),')',sep = '')

z$virginica <- as.factor(z$virginica)
pr_virginica <- pr_curve(data = z, virginica, .pred_virginica)
aps_virginica <- mean(pr_virginica$precision)
virginica <- paste('virginica (AP = ',toString(round(aps_virginica,2)),')',sep = '')


fig <- plot_ly()%>%
  add_segments(x = 0, xend = 1, y = 1, yend = 0, line = list(dash = "dash", color = 'black'), showlegend = FALSE) %>% 
  add_trace(data = pr_setosa,x = ~recall, y = ~precision, mode = 'lines', name = setosa, type = 'scatter')%>% 
  add_trace(data = pr_versicolor,x = ~recall, y = ~precision, mode = 'lines', name = versicolor, type = 'scatter')%>% 
  add_trace(data = pr_virginica,x = ~recall, y = ~precision, mode = 'lines', name = virginica, type = 'scatter')%>% 
  layout(xaxis = list(
    title = "Recall"
  ), yaxis = list(
    title = "Precision"
  ),legend = list(x = 100, y = 0.5))
fig
```
                           

## References


Learn more about histograms, filled area plots and line charts:

* https://plot.ly/r/histograms/

* https://plot.ly/r/filled-area-plots/

* https://plot.ly/r/line-charts/                   

                                                                                                                                            
                                                                                                           
                          