---
description: How to make subplots in with Plotly's R graphing library. Examples of stacked, custom-sized, gridded, and annotated subplots.
display_as: file_settings
language: r
layout: base
name: Subplots
order: 12
output:
  html_document:
    keep_md: true
page_type: u-guide
permalink: r/subplots/
thumbnail: thumbnail/subplots.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

### Subplots with Plotly

Plotly’s R graphing library makes it easy to create interactive, publication-quality graphs.

Plotly also has subplot capabilities. This page documents the usage of the lower-level `subplot` module. 

#### Simple Subplot

Figures with subplots are created using the `subplot` function.

Here is an example of creating a figure that includes two `scatter` traces which are side-by-side since there are 2 columns and 1 row in the subplot layout.

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3))) %>%
  layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))


fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3))) %>%
  layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))

fig <- subplot(fig1, fig2) %>% 
  layout(title = 'Side By Side Subplots')
fig

```

#### Stacked Subplots

Here is an example of creating a figure with subplots that are stacked on top of each other since there are 3 rows and 1 column in the subplot layout.

```{r}
library(plotly) 
fig1 <- plot_ly(x = c(3, 4, 5), y = c(1000, 1100, 1200), type = 'scatter', mode = 'lines+markers') 
fig2 <- plot_ly(x = c(2, 3, 4), y = c(100, 110, 120), type = 'scatter', mode = 'lines+markers') 
fig3 <- plot_ly(x = c(0, 1, 2), y = c(10, 11, 12), type = 'scatter', mode = 'lines+markers') 
fig <- subplot(fig1, fig2, fig3, nrows = 3) %>% 
  layout(title = list(text = "Stacked Subplots"),
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff')) 
fig

```

#### Multiple Subplots

Here is an example of creating a 2 x 2 subplot grid and populating each subplot with  `scatter` trace.

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))


fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))


fig3 <- plot_ly(x = c(300,400,500), y = c(600,700,800), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))


fig4 <- plot_ly(x = c(4000,5000,6000), y = c(7000,8000,9000), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))

fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2) %>%
  layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))
fig
```

#### Multiple Subplots with Titles
The `annotations` argument can be used to position text annotations as titles for each subplot.

Here is an example of adding subplot titles to a 2 x 2 subplot grid of scatter traces.

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))


fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))


fig3 <- plot_ly(x = c(300,400,500), y = c(600,700,800), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))


fig4 <- plot_ly(x = c(4000,5000,6000), y = c(7000,8000,9000), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))

fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2)%>% 
  layout(title = 'Multiple Subplots with Titles',
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))

annotations = list( 
  list( 
    x = 0.2,  
    y = 1.0,  
    text = "Plot 1",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.8,  
    y = 1,  
    text = "Plot 2",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.2,  
    y = 0.45,  
    text = "Plot 3",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.8,  
    y = 0.45,  
    text = "Plot 4",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))

fig <- fig %>%layout(annotations = annotations) 
#options(warn = -1)
fig
```

#### Subplots with Annotations

```{r}
library(plotly) 
fig1 <- plot_ly(x = c(1, 2, 3), y = c(4, 5, 6), type = 'scatter', mode = 'markers+text' 
                ,text = list("Text D", "Text E", "Text F"), textposition = "bottom center" 
                ,texttemplate = "%{text}") 
 
fig2 <- plot_ly(x = c(20, 30, 40), y = c(50, 60, 70), type = 'scatter', mode = 'markers+text' 
                ,text = list("Text D", "Text E", "Text F"), textposition = "bottom center" 
                ,texttemplate = "%{text}") 
 
fig <- subplot(fig1, fig2)%>% 
  layout(title = list(text = "Subplots with Annotations"),
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff')) 
fig

```

#### Customize Subplot Column Widths and Row Heights
The `widths` argument can be used to customize the relative widths of the columns in a subplot grid. It should be set in a list of numbers so that they sum to 1, and used to compute the relative widths of the subplot grid columns. The `heights` argument serves the same purpose for controlling the relative heights of rows in the subplot grid.

Here is an example of creating a figure with two scatter traces in side-by-side subplots. The left subplot is set to be wider than the right one.

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))


fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))

fig <- subplot(fig1, fig2, widths = c(0.7, 0.3)) %>%
               layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))
fig
```

#### Subplots in Dash

[Dash for R](https://dashr.plotly.com) is an open-source framework for building analytical applications, with no Javascript required, and it is tightly integrated with the Plotly graphing library.

Learn about how to install Dash for R at https://dashr.plot.ly/installation.

Everywhere in this page that you see fig, you can display the same figure in a Dash for R application by passing it to the figure argument.


```{r eval=FALSE}

library(dash)
library(dashCoreComponents)
library(dashHtmlComponents)
library(plotly)

app <- Dash$new()

app$layout(
  htmlDiv(
    list(
      dccGraph(id = 'graph-with-slider'),
      htmlLabel('Subplots Width:'),
      dccSlider(
        id='slider',
        min = 0,
        max = 1,
        value = 0.5, 
        step=0.01
      )
      
    )
  )
)
app$callback(
  output(id = 'graph-with-slider', property='figure'),
  params=list(input(id='slider', property='value')),
  function(value) {
    fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+marker', 
                    marker = list(line = list(width = 3)))
    
    
    fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+marker',
                    marker = list(line = list(width = 3)))
    
    fig <- subplot(fig1, fig2, widths = c(value, 1 - value))
    return(fig)
  })
```

After executing this code, give app$run_server() in the console to start the dash.

#### Customizing Subplot Axes
After a figure with subplots is created using the `subplot` function, its axis properties (title, font, range, grid style, etc.) can be customized using the `xaxis` and `yaxis` graph object figure methods.  By default, these methods apply to all of the x axes or y axes in the figure. The `row` and `col` arguments can be used to control which axes are targeted by the update.

Here is an example that creates a figure with a 2 x 2 subplot grid, populates each subplot with a scatter trace, and then updates the x and y axis titles for each subplot individually.

```{r}

library(plotly)

#Initialize figures 
fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))%>%
  layout(xaxis = list(title = 'xaxis1 title'), yaxis = list(title = 'yaxis1 title'))

fig2 <- plot_ly(x = c(20,30,40), y = c(50,60,70), type = 'scatter', mode = 'lines+markers',
                marker = list(line = list(width = 3)))%>%
  layout(xaxis = list(title = 'xaxis2 title', range = c(10,50)), yaxis = list(title = 'yaxis2 title', range = c(40,80)))


fig3 <- plot_ly(x = c(300,400,500), y = c(600,700,800), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))%>%
  layout(xaxis = list(title = 'xaxis3 title', showgrid = FALSE), yaxis = list(title = 'yaxis3 title', showgrid = FALSE))


fig4 <- plot_ly(x = c(4000,5000,6000), y = c(7000,8000,9000), type = 'scatter', mode = 'lines+markers', 
                marker = list(line = list(width = 3)))%>%
  layout(xaxis = list(title = 'xaxis4 title', type = 'log'), yaxis = list(title = 'yaxis4 title'))

#creating subplot
fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2, titleY = TRUE, titleX = TRUE, margin = 0.1 )
fig <- fig %>%layout(title = 'Customizing Subplot Axes',
                     plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))

# Update title
annotations = list( 
  list( 
    x = 0.2,  
    y = 1.0,  
    text = "Plot 1",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.8,  
    y = 1,  
    text = "Plot 2",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.2,  
    y = 0.4,  
    text = "Plot 3",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.8,  
    y = 0.4,  
    text = "Plot 4",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))

fig <- fig %>%layout(annotations = annotations) 
fig
```

#### Subplots with Shared X-Axes
The `shareX_x` argument  can be used to link the x axes of subplots in the resulting figure. The `margin` argument is used to control the vertical spacing between rows in the subplot grid.

Here is an example that creates a figure with 3 vertically stacked subplots with linked x axes. A small margin value is used to reduce the spacing between subplot rows.

```{r}

library(plotly) 
fig1 <- plot_ly(x = c(3, 4, 5), y = c(1000, 1100, 1200), type = 'scatter', mode = 'lines+markers') 
fig2 <- plot_ly(x = c(2, 3, 4), y = c(100, 110, 120), type = 'scatter', mode = 'lines+markers') 
fig3 <- plot_ly(x = c(0, 1, 2), y = c(10, 11, 12), type = 'scatter', mode = 'lines+markers') 
fig <- subplot(fig1, fig2, fig3, nrows = 3, shareX = TRUE) %>% 
  layout(title = list(text = "Stacked Subplots with Shared X-Axes"),
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff')) 
fig
```

#### Subplots with Shared Y-Axes
The `shareY` argument can be used to link the y axes of subplots in the resulting figure.

Here is an example that creates a figure with a 2 x 2 subplot grid, where the y axes of each row are linked.


```{r}
library(plotly) 
fig1 <- plot_ly(x = c(1, 2, 3), y = c(2, 3, 4), type = 'scatter', mode = 'lines+markers') 
fig2 <- plot_ly(x = c(20, 30, 40), y = c(5, 5, 5), type = 'scatter', mode = 'lines+markers') 
fig3 <- plot_ly(x = c(2, 3, 4), y = c(600, 700, 800), type = 'scatter', mode = 'lines+markers') 
fig4 <- plot_ly(x = c(4000, 5000, 6000), y = c(7000, 8000, 9000), type = 'scatter', mode = 'lines+markers') 
fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2, shareY = TRUE) %>% 
  layout(title = list(text = "Multiple Subplots with Shared Y-Axes"),
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff')) 
fig

```

### Subplots with Shared Colorscale

To share colorscale information in multiple subplots, you can use [coloraxis](https://plotly.com/r/reference/layout/coloraxis/).

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2,3), y = c(4,5,6), type = 'bar', 
                marker = list(color = c(4,5,6), coloraxis="coloraxis"))

fig2 <- plot_ly(x = c(1,2,3), y = c(2,3,5), type = 'bar', 
                marker = list(color = c(2,3,5), coloraxis="coloraxis"))

fig <- subplot(fig1, fig2, shareY = TRUE)
fig <- fig %>%layout(showlegend = FALSE, coloraxis=list(colorscale='RdBu'),
                     plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))
fig
```

#### Custom Sized Subplot with Subplot Titles
The heights and weights argument in subplot function is used to custom size the individual plots in the subplot.

Here is an example that creates a 2 by 2 subplot grid containing 3 subplots. 

```{r}

library(plotly) 
fig1 <- plot_ly(x = c(1, 2), y = c(1, 2), type = 'scatter', mode = 'lines+markers') 
fig2 <- plot_ly(x = c(1, 2), y = c(1, 2), type = 'scatter', mode = 'lines+markers') 
fig3 <- plot_ly(x = c(1, 2, 3), y = c(2, 1, 2), type = 'scatter', mode = 'lines+markers')%>%  
  layout(annotations = list( 
    list(x = 0.52 , y = 1.05, text = "Third Subplot", showarrow = F, xref='paper', yref='paper')) 
  ) 

s1 <- subplot(fig1, fig2)%>%  
  layout(annotations = list( 
    list(x = 0.15 , y = 1.2, text = "First Subplot", showarrow = F, xref='paper', yref='paper'), 
    list(x = 0.85 , y = 1.2, text = "Second Subplot", showarrow = F, xref='paper', yref='paper')) 
  ) 
fig <- subplot(s1, fig3, nrows = 2, margin = 0.07) %>%  
  layout(title = "Specs with Subplot Title", 
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'),
         showlegend=FALSE,showlegend2=FALSE, 
         margin = 0.01) 

fig
```

#### Multiple Custom Sized Subplots

Here is an example that uses the `heights` and `weights` subplot options to create a custom subplot layout with subplots of mixed sizes. 

```{r}

library(plotly)

fig1 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(1,1)',
                marker = list(line = list(width = 3)))

fig2 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(2,1)',
                marker = list(line = list(width = 3)))

fig3 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(1,2)',
                marker = list(line = list(width = 3)))

fig4 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(3,1)',
                marker = list(line = list(width = 3)))

fig5 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(5,1)',
                marker = list(line = list(width = 3)))

fig6 <- plot_ly(x = c(1,2), y = c(1,2), type = 'scatter', mode = 'lines+markers', name = '(5,2)',
                marker = list(line = list(width = 3)))

s1 <- subplot(fig1, fig2, nrows = 2)
s2 <- subplot(s1, fig3)
s3 <- subplot(s2, fig4, nrows = 2)
s4 <- subplot(fig5, fig6)
s5 <- subplot(s3, s4, nrows = 2, heights = c(0.8,0.2)) %>%
  layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))
s5
```

#### Subplots Types
By default, the `subplots` function assumes that the traces that will be added to all subplots are 2-dimensional cartesian traces (e.g. `scatter`, `bar`, `histogram`, `violin`, etc.).  


 - trace type: A trace type name (e.g. `"bar"`, `"contour"`, `"density"`, `"scatter"`, etc.) which will be used to determine the appropriate subplot type for that trace.

Here is an example that creates and populates a 2 x 2 subplot grid containing 4 different subplot types.

```{r}
library(plotly) 

fig1 <- plot_ly(y = c(2, 3, 1), type = 'bar') 
 
data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv") 
fig2 <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers', color = ~Gap, colors = 'Reds', 
                marker = list(size = ~Gap, opacity = 0.5))%>%  
  layout(xaxis = list(showgrid = FALSE), 
         yaxis = list(showgrid = FALSE)) 
fig2 <- hide_colorbar(fig2) 
 
fig3<- plot_ly(z = ~volcano, type = "contour") 
fig3 <- hide_colorbar(fig3) 
 
density <- density(diamonds$carat) 
fig4 <- plot_ly(x = ~density$x, y = ~density$y, type = 'scatter', mode = 'lines', fill = 'tozeroy') 
fig4 <- fig4 %>% layout(xaxis = list(title = 'Carat'), 
                      yaxis = list(title = 'Density'), showlegend=FALSE) 
 
fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2, margin = 0.05) %>%
  layout(plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))
 
fig
```


#### Reference
All of the axis properties are found here: https://plotly.com/r/axes/#


```{r}
tools:::Rd2ex(utils:::.getHelpFile(as.character(help(subplot))))
```
