---
description: How to add images to charts as background images or logos in R
display_as: file_settings
language: r
layout: base
name: Images
order: 30
permalink: r/images/
thumbnail: thumbnail/images.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

### Add a Background Image

In this page we explain how to add static, non-interactive images as background, logo or annotation images to a figure. For exploring image data in interactive charts, see the [tutorial on displaying image data](https://plotly.com/r/displaying-images/).

A background image can be added to the layout of a figure by setting the `images` parameter of `plot_ly.layout`. The
`source` attribute of a `layout.images` can be the URL of an image, or a PIL
Image object (`from PIL import Image; img = Image.open('filename.png')`).

```{r}
library('plotly')
# Create figure
plot_ly(x = c(0, 0.5, 1, 2, 2.2), y = c(1.23, 2.5, 0.42, 3, 1), type = 'scatter', mode = 'lines+markers') %>% 
  # Add trace
  layout(
    images = list(
      list(
        # Add images
        source =  "https://images.plot.ly/language-icons/api-home/r-logo.png?raw=true",
        xref = "x",
        yref = "y",
        x = 0.2,
        y = 3,
        sizex = 2,
        sizey = 2,
        sizing = "stretch",
        opacity = 0.4,
        layer = "below"
      )
    )
  )%>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
```

### Add a Logo
See more examples of [adding logos to charts](https://plotly.com/r/logos/)!

```{r}
library(plotly) 
x= c("-35.3", "-15.9", "-15.8", "-15.6", "-11.1", 
     "-9.6", "-9.2", "-3.5", "-1.9", "-0.9", 
     "1.0", "1.4", "1.7", "2.0", "2.8", "6.2", 
     "8.1", "8.5", "8.5", "8.6", "11.4", "12.5", 
     "13.3", "13.7", "14.4", "17.5", "17.7", 
     "18.9", "25.1", "28.9", "41.4") 
 
y = c("Designers, musicians, artists, etc.", 
      "Secretaries and administrative assistants", 
      "Waiters and servers", "Archivists, curators, and librarians", 
      "Sales and related", "Childcare workers, home car workers, etc.", 
      "Food preparation occupations", "Janitors, maids, etc.", 
      "Healthcare technicians, assistants. and aides", 
      "Counselors, social and religious workers", 
      "Physical, life and social scientists", "Construction", 
      "Factory assembly workers", "Machinists, repairmen, etc.", 
      "Media and communications workers", "Teachers", 
      "Mechanics, repairmen, etc.", "Financial analysts and advisers", 
      "Farming, fishing and forestry workers", 
      "Truck drivers, heavy equipment operator, etc.", "Accountants and auditors", 
      "Human resources, management analysts, etc.", "Managers", 
      "Lawyers and judges", "Engineers, architects and surveyors", 
      "Nurses", "Legal support workers", 
      "Computer programmers and system admin.", "Police officers and firefighters", 
      "Chief executives", "Doctors, dentists and surgeons") 
 
df = data.frame(x,y,stringsAsFactors = FALSE) 

m = list(r=20, l=300, b=75, t=125) 

fig <- plot_ly(data = df, x = ~x, y = ~y, type = 'bar', orientation = 'h', 
               marker = list(color = 'rgb(253, 240, 54)',  
               line = list(width = 2, color = 'rgb(0, 0, 0)'))) %>%  
  layout( xaxis = list(title = ""), yaxis = list(title = ""),
    images = list(  
      list(  
        source =  "https://raw.githubusercontent.com/cldougl/plot_images/add_r_img/vox.png",  
        xref = "paper",  
        yref = "paper",  
        x = 1.05,  
        y = 1.05,  
        sizex = 0.2,  
        sizey = 0.2,  
        xanchor="right",  
        yanchor="bottom" 
      )  
    ) ) 
 
fig <- fig %>% layout(autosize = F, margin = m, 
                      title=(paste("Moving Up, Moving Down<br>" , 
                               "<i>Percentile change in income between childhood and adulthood</i>")), 
                      hovermode="x" 
                      ) %>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
fig
```

### Zoom on Static Images

```{r}
library(plotly) 
 
#Constants 
img_width = 1600 
img_height = 900 
scale_factor = 0.5 
 
 
# Add invisible scatter trace. 
# This trace is added to help the autoresize logic work. 
fig <- plot_ly(width=img_width * scale_factor, 
               height=img_height * scale_factor 
) %>% 
  add_trace( x= c(0, img_width * scale_factor), 
             y= c(0, img_height * scale_factor), 
             type = 'scatter',  mode = 'markers', alpha = 0) 
 
# Configure axes 
xconfig <- list( 
  title = "", 
  zeroline = FALSE, 
  showline = FALSE, 
  showticklabels = FALSE, 
  showgrid = FALSE, 
  range = c(0, img_width * scale_factor) 
) 
 
yconfig <- list( 
  title = "", 
  zeroline = FALSE, 
  showline = FALSE, 
  showticklabels = FALSE, 
  showgrid = FALSE, 
  range = c(0, img_height * scale_factor), 
  scaleanchor="x" 
) 
 
fig <- fig %>% layout(xaxis = xconfig, yaxis = yconfig) 
 
# Add image 
 
fig <- fig %>% layout( 
  images = list(  
    list(  
      source =  "https://raw.githubusercontent.com/michaelbabyn/plot_data/master/bridge.jpg",  
      x=0, 
      sizex=img_width * scale_factor, 
      y=img_height * scale_factor, 
      sizey=img_height * scale_factor, 
      xref="x", 
      yref="y", 
      opacity=1.0, 
      layer="below", 
      sizing="stretch" 
    )  
  )) 
 
# Configure other layout 
 
m = list(r=0, l=0, b=0, t=0) 
fig <- fig %>% layout(margin = m) %>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
fig
```

### Annotating layout image with shapes
It can be useful to add shapes to a layout image, for highlighting an object, drawing bounding boxes as part of a machine learning training set, or identifying seeds for a segmentation algorithm.

In order to enable shape drawing, you need to

+ define a dragmode corresponding to a drawing tool (`'drawline'`,`'drawopenpath'`, `'drawclosedpath'`, `'drawcircle'`, or `'drawrect'`)

+ add [modebar buttons](https://plotly-r.com/control-modebar.html) corresponding to the drawing tools you wish to use.

The style of new shapes is specified by the `newshape` layout attribute. Shapes can be selected and modified after they have been drawn. More details and examples are given in the [tutorial on shapes](https://plotly.com/r/shapes/#drawing-shapes-on-cartesian-plots).

Drawing or modifying a shape triggers a `relayout` event, which [can be captured by a callback inside a Dash application](https://dashr.plotly.com/interactive-graphing).

```{r}
library(plotly) 
 
#Constants 
img_width = 1600 
img_height = 900 
scale_factor = 0.5 
 
fig <- plot_ly() %>% 
  add_trace( x= c(0, img_width ), 
             y= c(0, img_height ), 
             type = 'scatter',  mode = 'markers', alpha = 0)%>% 
  layout(images = list(  
    list(  
      source =  "https://raw.githubusercontent.com/michaelbabyn/plot_data/master/bridge.jpg",  
      x=0, 
      sizex=img_width, 
      y=0, 
      sizey=img_height, 
      xref="x", 
      yref="y", 
      opacity=1.0, 
      layer="below" 
    )  
  )) 
 
xconfig <- list( 
  title = "", 
  showgrid = FALSE, 
  range = c(0, img_width) 
) 
 
yconfig <- list( 
  title = "", 
  showgrid = FALSE, 
  range = c(img_height,0), 
  scaleanchor="x" 
) 
 
fig <- fig %>% layout(xaxis = xconfig, yaxis = yconfig) 
 
#Add lineshape 
fig <- fig %>%  
  add_segments(x = 650, xend = 1080, y = 380, yend = 180, line = list( color = 'cyan'),inherit = FALSE, showlegend = FALSE) 
 
fig <- fig %>% layout(dragmode='drawrect', 
                      newshape=list(line = list(color='cyan')), 
                      title = 'Drag to add annotations - use modebar to change drawing tool') 
 
#Add modebar buttons 
fig <- fig %>%  
  config(modeBarButtonsToAdd = c('drawline', 
                                 'drawopenpath', 
                                 'drawclosedpath', 
                                 'drawcircle', 
                                 'drawrect', 
                                 'eraseshape')) %>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
 
fig
```


### Images Placed Relative to Axes

Using `xref='x domain'` or `yref='y domain'`, images can be placed relative to
axes. As an example, the following shows how to put an image in the top corner
of a subplot (try panning and zooming the resulting figure):

```{r}
library(plotly) 

db1 <- iris[iris$Species == "setosa", ] 
db2 <- iris[iris$Species == "versicolor", ] 
db3 <- iris[iris$Species == "virginica", ] 

fig1 <- plot_ly(data = db1, x = ~Sepal.Length, y = ~Sepal.Width, type = 'scatter', mode = 'markers') %>% 
  layout(xaxis = list(range = c(4,8))) 
# add images
fig1 <- fig1 %>% layout( 
  images = list(  
    list(  
      # sources of images
      source =  "https://upload.wikimedia.org/wikipedia/commons/thumb/f/fe/Iris_setosa_var._setosa_%282595031014%29.jpg/360px-Iris_setosa_var._setosa_%282595031014%29.jpg",  
      row=1, 
      col=1, 
      source=1, 
      xref="x domain", 
      yref="y domain", 
      x=1, 
      y=1, 
      xanchor="right", 
      yanchor="top", 
      sizex=0.2, 
      sizey=0.2 
    )  
  )) 

fig2 <- plot_ly(data = db2, x = ~Sepal.Length, y = ~Sepal.Width, type = 'scatter', mode = 'markers') %>% 
  layout(xaxis = list(range = c(4,8))) 
# add images
fig2 <- fig2 %>% layout( 
  images = list(  
    list(  
      # sources of images
      source =  "https://upload.wikimedia.org/wikipedia/commons/thumb/3/38/Iris_versicolor_quebec_1.jpg/320px-Iris_versicolor_quebec_1.jpg",  
      row=1, 
      col=2, 
      source=2, 
      xref="x domain", 
      yref="y domain", 
      x=2.05, 
      y=1, 
      xanchor="right", 
      yanchor="top", 
      sizex=0.2, 
      sizey=0.2 
    )  
  )) 

fig3 <- plot_ly(data = db3, x = ~Sepal.Length, y = ~Sepal.Width, type = 'scatter', mode = 'markers') %>% 
  layout(xaxis = list(range = c(4,8), title = 'Sepal..Length')) 
# add images  
fig3 <- fig3 %>% layout( 
  images = list(  
    list(  
      # sources of images
      source =  "https://upload.wikimedia.org/wikipedia/commons/thumb/f/f8/Iris_virginica_2.jpg/480px-Iris_virginica_2.jpg",  
      row=1, 
      col=3, 
      source=2, 
      xref="x domain", 
      yref="y domain", 
      x=3.15, 
      y=1, 
      xanchor="right", 
      yanchor="top", 
      sizex=0.2, 
      sizey=0.2 
    )  
  )) 

fig <- subplot(fig1, fig2, fig3, shareY = TRUE, shareX = TRUE) %>% layout(showlegend = FALSE) 

annotations = list(  
  list(  
    x = 0.2,   
    y = 1.0,   
    font = list(size = 10),   
    text = "species=setosa",   
    xref = "paper",   
    yref = "paper",   
    xanchor = "center",   
    yanchor = "bottom",   
    showarrow = FALSE  
  ),   
  list(  
    x = 0.8,   
    y = 1,   
    font = list(size = 10),   
    text = "species=versicolor",   
    xref = "paper",   
    yref = "paper",   
    xanchor = "center",   
    yanchor = "bottom",   
    showarrow = FALSE  
  ),   
  list(  
    x = 0.5,   
    y = 1,   
    font = list(size = 10),   
    text = "species=virginica",   
    xref = "paper",   
    yref = "paper",   
    xanchor = "center",   
    yanchor = "bottom",   
    showarrow = FALSE  
  )) 

fig <- fig %>%
  layout(annotations = annotations)  %>%layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
options(warn = -1) 
fig
```

#### Reference
See https://plotly.com/r/reference/layout/images/ for more information and chart attribute options!