---
description: A reference for the built-in named continuous (sequential, diverging
      and cyclical) color scales in Plotly.
display_as: file_settings
language: r
layout: base
name: Built-in Continuous Color Scales
order: 28
permalink: r/builtin-colorscales/
thumbnail: thumbnail/heatmap_colorscale.jpg
v4upgrade: true
---

### Using Built-In Continuous Color Scales

Plotly R comes with a large number of built-in continuous color scales. It can be viewed using 'RColorBrewer' package. 

When using continuous color scales, you will often want to [configure various aspects of its range and colorbar](https://plotly.com/r/colorscales/).

### Discrete Color Sequences

Plotly also comes with some built-in discrete color sequences.

### Named Built-In Continuous Color Scales

You can use any of the following names as string values to set `colorscale` arguments.
These strings are case-sensitive.

```{r}

library("RColorBrewer") 
brewer.pal.info

```

Built-in color scales are stored as a string of CSS colors:

```{r}

library("RColorBrewer") 
brewer.pal(n = 8, name = "YlGn")
```



### Built-In Sequential Color scales

A collection of predefined sequential colorscales is provided in the 'RColorBrewer' package. Sequential color scales are appropriate for most continuous data, but in some cases it can be helpful to use a diverging or cyclical color scale (see below).

Here are all the built-in scales in the 'RColorBrewer' package:

```{r}

library("RColorBrewer") 
display.brewer.all(type = 'seq')
```


### Built-In Diverging Color scales

A collection of predefined diverging color scales is provided in the 'RColorBrewer' package.
Diverging color scales are appropriate for continuous data that has a natural midpoint or an  otherwise informative special value, such as 0 altitude, or the boiling point
of a liquid. These scales are intended to be used when explicitly setting the midpoint of the scale.

Here are all the built-in scales in the 'RColorBrewer' package:

```{r}

library("RColorBrewer") 
display.brewer.all(type = 'div')
```

