---
description: How to design figures with multiple chart types in Plotly for R.
display_as: file_settings
language: r
layout: base
name: Multiple Chart Types
order: 13
output:
  html_document:
    keep_md: true
permalink: r/graphing-multiple-chart-types/
thumbnail: thumbnail/multiple-chart-type.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
## Multiple Chart Types in R

How to design figures with multiple chart types in R.

### Chart Types versus Trace Types

Plotly's figure data structure supports defining [subplots](https://plotly.com/r/subplots/) of [various types](https://plotly.com/r/mixed-subplots/) (e.g. [cartesian](https://plotly.com/r/axes/), [polar](https://plotly.com/r/polar-chart/), [3-dimensional](https://plotly.com/r/3d-charts/), [maps](https://plotly.com/r/maps/) etc) with attached traces of various compatible types (e.g. scatter, bar, choropleth, surface etc). This means that **Plotly figures are not constrained to representing a fixed set of "chart types"** such as scatter plots only or bar charts only or line charts only: any subplot can contain multiple traces of different types.


### Multiple Trace Types with Plotly

Figures produced with Plotly have the add_trace() method, so it is easy to start with a Plotly figure containing only traces of a given type, and add traces of another type.

```{r}
library(plotly)
data <- data.frame(
  Fruits = c ("apples", "bananas", "oranges"),
  Line = c(1,3,2),
  Bar = c(2,1,3))

fig <- plot_ly(data , x = ~Fruits, y = ~Bar, type = 'bar', name = 'Last Year') %>%
  add_trace(data , x = ~Fruits, y = ~Line, type = 'scatter',  mode = 'lines', name = 'This year')

fig <- fig %>% layout(yaxis = list(title = "Amount"))
fig <- fig %>% layout(legend=list(title=list(text='<b> Time Period </b>')))
fig
```

#### Line Chart and a Bar Chart

```{r}
library(plotly)
data <- data.frame(
  X = c (0, 1, 2, 3, 4, 5),
  Line = c(1.5, 1, 1.3, 0.7, 0.8, 0.9),
  Bar = c(1, 0.5, 0.7, -1.2, 0.3, 0.4))

fig <- plot_ly(data , x = ~X, y = ~Bar, type = 'bar') %>%
  add_trace(data , x = ~X, y = ~Line, type = 'scatter',  mode = 'lines+markers')

fig
```

#### A Contour and Scatter Plot of the Method of Steepest Descent

```{r}
library(plotly)
library(jsonlite)
urlfile<-'https://raw.githubusercontent.com/plotly/datasets/master/steepest.json'
data<-fromJSON(url(urlfile))
X <- data[["contour_x"]][,]
Y <- data[["contour_y"]][,]
Z <- data[["contour_z"]][,,]
fig <- plot_ly() %>%
  add_trace(x = X, y= Y, z = Z, type = "contour") %>%
  hide_colorbar()%>% layout(showlegend = FALSE) %>%
  add_trace(x = data$trace_x, y = data$trace_y, type = "scatter",
            mode = "lines+markers", name = 'steepest', inherit =  FALSE,
            marker = list(color = 'black'), line = list(color = 'black'))
fig
```

#### Reference
See https://plotly.com/r/reference/ for more information and attribute options!
