---
description: How to set the global font, title, legend-entries, and axis-titles in R.
display_as: file_settings
language: r
layout: base
name: Setting the Font, Title, Legend Entries, and Axis Titles
order: 24
output:
  html_document:
    keep_md: true
permalink: r/figure-labels/
redirect_form: r/font/
thumbnail: thumbnail/figure-labels.png
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
## Setting the Font, Title, Legend Entries, and Axis Titles in R

How to set the global font, title, legend-entries, and axis-titles in for plots in R.


### Automatic Labelling with Plotly 

When using Plotly, your axes is automatically labelled, and it's easy to override the automation for a customized figure using the `labels` keyword argument. The title of your figure is up to you though!

Here's a figure with automatic labels and then the same figure with overridden labels. Note the fact that when overriding labels, the axes, legend title *and hover labels* reflect the specified labels automatically.

```{r}

library(plotly)

data("iris")

fig1 <-  plot_ly(data = iris ,x =  ~Sepal.Length, y = ~Sepal.Width, color = ~Species, type = 'scatter', mode = 'markers')%>%
        layout(title = 'Automatic Labels Based on Data Frame Column Names', plot_bgcolor = "#e5ecf6")
fig1

#Manually specifying labels

fig2 <-  plot_ly(data = iris ,x =  ~Sepal.Length, y = ~Sepal.Width, color = ~Species, type = 'scatter', mode = 'markers')%>%
  layout(title = 'Manually Specified Labels', plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Sepal Length (cm)'), 
         yaxis = list(title = 'Sepal Width (cm)'), legend = list(title=list(text='<b> Species of Iris </b>')))
fig2
```


### Global and Local Font Specification

You can set the figure-wide font with the `layout.font.family` attribute, which will apply to all titles and tick labels, but this can be overridden for specific plot items like individual axes and legend titles etc. In the following figure, we set the figure-wide font to Courier New in blue, and then override this for certain parts of the figure.

```{r}
library(plotly)
data(iris)

t <- list(
  family = "Courier New",
  size = 14,
  color = "blue")
t1 <- list(
  family = "Times New Roman",
  color = "red"
)
t2 <- list(
  family = "Courier New",
  size = 14,
  color = "green")
t3 <- list(family = 'Arial')

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Sepal.Width, color = ~Species, 
               type = 'scatter', mode = 'markers')%>% 
  layout(title= list(text = "Playing with Fonts",font = t1), font=t, 
         legend=list(title=list(text='Species',font = t2)), 
         xaxis = list(title = list(text ='Sepal.Length', font = t3)),
         plot_bgcolor='#e5ecf6')
fig
```

### Fonts and Labels in Dash
[Dash for R](https://dashr.plotly.com) is an open-source framework for building analytical applications, with no Javascript required, and it is tightly integrated with the Plotly graphing library.

Learn about how to install Dash for R at https://dashr.plot.ly/installation.

Everywhere in this page that you see fig, you can display the same figure in a Dash for R application by passing it to the figure argument.

```{r}

library(dash)
library(dashCoreComponents)
library(dashHtmlComponents)
library(devtools)
library(plotly)
library(dashDaq)
data("iris")


app <- Dash$new()

app$layout(
  htmlDiv(
    list(
      dccGraph(id = 'graph'),
      daqColorPicker(id = 'font',
                     label = 'Font Color',
                     value = list(hex = "#119DFF")),
      daqColorPicker(id = 'title',
                     label = 'Title Color',
                     value = list(hex = "#2A0203"))
      
    )
  )
)

app$callback(
  output(id = 'graph', property='figure'),
  params=list(input(id='font', property='value'), 
              input(id='title', property='value')),
  function(font_color, title_color) {
    t <- list(
      family = "Courier New",
      size = 14,
      color = strsplit(toString(font_color), split = ",")[[1]][1]
      )
    t1 <- list(
      family = "Times New Roman",
      color = strsplit(toString(title_color), split = ",")[[1]][1]
    )
    t2 <- list(
      family = "Courier New",
      size = 14,
      color = strsplit(toString(font_color), split = ",")[[1]][1])
    t3 <- list(family = 'Arial')
    
    fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Sepal.Width, color = ~Species, 
                   type = 'scatter', mode = 'markers')%>% 
      layout(title= list(text = "Playing with Fonts",font = t1), font=t, 
             legend=list(title=list(text='Species',font = t2)), 
             xaxis = list(title = list(text ='Sepal.Length', font = t3)))
  
    return(fig)
  })


```

After executing this code, give app$run_server() in the console to start the dash.

### Manual Labelling in Plotly 

Explicitly Labeling traces and axes in Plotly.

```{r}

library(plotly)
t <- list(
  family = "Courier New, monospace",
  size = 15,
  color = "RebeccaPurple")
x1 = c(0, 1, 2, 3, 4, 5, 6, 7, 8)
y1 = c(0, 1, 2, 3, 4, 5, 6, 7, 8)

x2 = c(0, 1, 2, 3, 4, 5, 6, 7, 8)
y2 = c(1, 0, 3, 2, 5, 4, 7, 6, 8)

df = data.frame(x1, y1, x2, y2)

fig <-  plot_ly()%>%
  add_trace(df, x = ~x1, y = ~y1, type = 'scatter', mode = 'lines+markers', name = 'Name of Trace 1')%>%
  add_trace(df, x = ~x2, y = ~y2, type = 'scatter', mode = 'lines+markers', name = 'Name of Trace 2')%>%
  layout(title = 'Plot Title', xaxis = list(title = 'X Axis Title'), font=t, plot_bgcolor = "#e5ecf6",
         yaxis = list(title = 'Y Axis Title'), legend = list(title=list(text='Legend Title')))

fig
```

The configuration of the legend is discussed in detail in the [Legends](https://plotly.com/r/legend/) page.

### Align Plot Title
The following example shows how to align the plot title in [layout.title](https://plotly.com/r/reference/layout/#layout-title). `x` sets the x position with respect to `xref` from "0" (left) to "1" (right), and `y` sets the y position with respect to `yref` from "0" (bottom) to "1" (top). Moreover, you can define `xanchor` to `left`,`right`, or `center` for setting the title's horizontal alignment with respect to its x position, and/or `yanchor` to `top`, `bottom`, or `middle` for setting the title's vertical alignment with respect to its y position.

```{r}
library(plotly)

fig <- plot_ly(x= c('Mon', 'Tue', 'Wed'), y= c(3,1,4), type= 'scatter', mode= 'lines+markers')%>% 
  layout(title = list(text='Plot Title', y = 0.95, x = 0.5, xanchor = 'center', yanchor =  'top'), plot_bgcolor = "#e5ecf6")
fig
```

#### Reference
See https://plotly.com/r/reference/layout/ for more information!
