---
description: How to make a graph with multiple axes (dual y-axis plots, plots with secondary axes) in R.
display_as: file_settings
language: r
layout: base
name: Multiple Axes
order: 28
page_type: example_index
permalink: r/multiple-axes/
thumbnail: thumbnail/multiple-axes.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

### Multiple Y Axes and Plotly

> *Note*: At this time, Plotly does not support multiple Y axes on a single figure. To make such a figure, use the [`subplot()`](https://plotly.com/r/subplot-charts/) function as documented below.


## Two Y Axes

```{r}
library(plotly)   
 
fig <- plot_ly()   
# Add traces
fig <- fig %>% add_trace(x = ~1:3, y = ~10*(4:6), name = "yaxis data", mode = "lines+markers", type = "scatter")   
 
ay <- list(   
  tickfont = list(color = "red"),   
  overlaying = "y",   
  side = "right",   
  title = "<b>secondary</b> yaxis title")   
 
fig <- fig %>% add_trace(x = ~2:4, y = ~4:6, name = "yaxis 2 data", yaxis = "y2", mode = "lines+markers", type = "scatter")   

# Set figure title, x and y-axes titles
fig <- fig %>% layout(   
  title = "Double Y Axis Example", yaxis2 = ay,  
  xaxis = list(title="xaxis title "),   
  yaxis = list(title="<b>primary</b> yaxis title")   
)%>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )   
 
fig
```

## Multiple axes in Dash

[Dash for R](https://dashr.plotly.com) is an open-source framework for building analytical applications, with no Javascript required, and it is tightly integrated with the Plotly graphing library.

Learn about how to install Dash for R at https://dashr.plot.ly/installation.

Everywhere in this page that you see fig, you can display the same figure in a Dash for R application by passing it to the figure argument.

```{r}
library(dash)
library(dashCoreComponents)
library(dashHtmlComponents)
library(plotly)

app <- Dash$new()

app$layout(
  htmlDiv(
    list(
      dccGraph(id = 'graph'),
      htmlLabel("Red line's axis:"),
      dccRadioItems(
        id='radio',
        options = list(list(label = "Primary", value = "Primary"),
                       list(label = "Secondary", value = "Secondary")),
        value = 'Secondary'
      )
    )
  )
)
app$callback(
  output(id = 'graph', property='figure'),
  params=list(input(id='radio', property='value')),
  function(value) {
    if(value == 'Primary'){
      fig <- plot_ly()   
      fig <- fig %>% add_trace(x = ~1:3, y = ~10*(4:6), name = "yaxis data", mode = "lines+markers", type = "scatter")   
      fig <- fig %>% add_trace(x = ~2:4, y = ~4:6, name = "yaxis 2 data", mode = "lines+markers", type = "scatter")   
      fig <- fig %>% layout(   
        title = "Double Y Axis Example",   
        xaxis = list(title="xaxis title"),   
        yaxis = list(title="<b>primary</b> yaxis title")   
      )%>%
        layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )   
      return(fig)
    }
    else{
      fig <- plot_ly()   
      fig <- fig %>% add_trace(x = ~1:3, y = ~10*(4:6), name = "yaxis data", mode = "lines+markers", type = "scatter")   
      
      ay <- list( 
        overlaying = "y",   
        side = "right",   
        title = "<b>secondary</b> yaxis title")   
      
      fig <- fig %>% add_trace(x = ~2:4, y = ~4:6, name = "yaxis 2 data", yaxis = "y2", mode = "lines+markers", type = "scatter")   
      
      fig <- fig %>% layout(   
        title = "Double Y Axis Example", yaxis2 = ay,   
        xaxis = list(title="xaxis title"),   
        yaxis = list(title="<b>primary</b> yaxis title")   
      )%>%
        layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )   
      return(fig)
    }
  })
#app$run_server()
```

Use `app$run_server()` to run the dash file.

## Multiple Y-Axes Subplots

```{r}
library(plotly)
# Top left
p1 <- plot_ly() %>% add_trace(x = c(1, 2, 3), y = c(2, 52, 62),type="scatter",mode="lines+markers",yaxis="y", name="yaxis data") %>%
  add_trace(x = c(1, 2, 3), y = c(40, 50, 60),type="scatter",mode="lines+markers",yaxis="y2", name="yaxis2 data") %>%
  layout(yaxis=list(side="left"),
         yaxis2=list(side="right",overlaying="y"),
         showlegend=TRUE)
# Top right
p2 <-plot_ly() %>% add_trace(x = c(1, 2, 3), y = c(2, 52, 62),type="scatter",mode="lines+markers",yaxis="y", name="yaxis3 data") %>%
  add_trace(x = c(1, 2, 3), y = c(40, 50, 60),type="scatter",mode="lines+markers",yaxis="y2", name="yaxis4 data") %>%
  layout(yaxis=list(side="left"),
         yaxis2=list(side="right",overlaying="y3"),
         showlegend=TRUE)
# Bottom left
p3 <- plot_ly() %>% add_trace(x = c(1, 2, 3), y = c(2, 52, 62),type="scatter",mode="lines+markers",yaxis="y", name="yaxis5 data") %>%
  add_trace(x = c(1, 2, 3), y = c(40, 50, 60),type="scatter",mode="lines+markers",yaxis="y2", name="yaxis6 data") %>%
  layout(yaxis=list(side="left"),
         yaxis2=list(side="right",overlaying="y5"),
         showlegend=TRUE)
# Bottom right
p4 <-plot_ly() %>% add_trace(x = c(1, 2, 3), y = c(2, 52, 62),type="scatter",mode="lines+markers",yaxis="y", name="yaxis7 data") %>%
  add_trace(x = c(1, 2, 3), y = c(40, 50, 60),type="scatter",mode="lines+markers",yaxis="y2", name="yaxis8 data") %>%
  layout(yaxis=list(side="left"),
         yaxis2=list(side="right",overlaying="y7"),
         showlegend=TRUE)

p <- subplot(p1,p2,p3,p4,nrows = 2, margin = 0.05)%>% layout(legend = list(x = 1.05, y = 1))%>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )
p
```

## Multiple Axes

Using Plotly for creating a figure with multiple axes

```{r}
library(plotly)   
 
fig <- plot_ly(width = 700)   
fig <- fig %>% add_trace(x = ~1:3, y = ~4:6, name = "yaxis1 data", mode = "lines+markers", type = "scatter")   
 
y2 <- list(   
  tickfont = list(color = "#ff7f0e"),   
  titlefont = list(color = "#ff7f0e"), 
  overlaying = "y",   
  side = "left",   
  anchor="free", 
  position=0.15, 
  title = "yaxis2 title") 
 
 
fig <- fig %>% add_trace(x = ~2:4, y = ~10*(4:6), name = "yaxis2 data", yaxis = "y2", mode = "lines+markers", type = "scatter")  
 
y3 <- list(   
  tickfont = list(color = "#d62728"),   
  titlefont = list(color = "#d62728"),  
  overlaying = "y",   
  side = "right",   
  title = "yaxis3 title") 
 
 
fig <- fig %>% add_trace(x = ~4:6, y = ~1000*(4:6), name = "yaxis3 data", yaxis = "y3", mode = "lines+markers", type = "scatter")  
 
y4 <- list(   
  tickfont = list(color = "#9467bd"),   
  titlefont = list(color = "#9467bd"),   
  overlaying = "y",   
  side = "right",   
  anchor="free", 
  position=0.85, 
  title = "yaxis4 title") 
 
 
fig <- fig %>% add_trace(x = ~5:7, y = ~10000*(4:6), name = "yaxis4 data", yaxis = "y4", mode = "lines+markers", type = "scatter")  
 
fig <- fig %>% layout( 
  title = "multiple y-axes example", yaxis2 = y2, yaxis3 = y3, yaxis4 = y4,   
  xaxis = list(title = '', domain = c(0.3, 0.7)),   
  yaxis = list(title="yaxis title",  
               tickfont = list(color = "#1f77b4"),   
               titlefont = list(color = "#1f77b4") 
               )   
)%>%
  layout(plot_bgcolor='#e5ecf6',  
          xaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff'),  
          yaxis = list(  
            zerolinecolor = '#ffff',  
            zerolinewidth = 2,  
            gridcolor = 'ffff')  
          )   
 
fig
```

### Reference
All of the y-axis properties are found here: https://plotly.com/r/reference/layout/yaxis/.  For more information on creating subplots see the [Subplots in R](https://plotly.com/r/subplot-charts/) section.
