---
title : Distplots in R
description: How to make interactive Distplots in R with Plotly.
display_as: statistical
language: r
layout: base
name: Distplots
order: 4
page_type: example_index
permalink: r/distplot/
thumbnail: thumbnail/distplot.jpg
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Combined statistical representations with histogram

Several representations of statistical distributions are available in plotly, such as [histograms](https://plotly.com/r/histograms/), [violin plots](https://plotly.com/r/violin/), [box plots](https://plotly.com/r/box-plots/#) (see [the complete list here](https://plotly.com/r/statistical-charts/). It is also possible to combine several representations in the same plot.


```{r}

library(plotly)
library(ggplot2)
library(reshape2)
data(tips)

p <- ggplot(tips, aes(x=total_bill, weight = tip, color=sex, fill = sex)) +
  geom_histogram(binwidth=2.5) +
  ylab("sum of tip") + 
  geom_rug(sides="t", length = unit(0.3, "cm"))
fig <- ggplotly(p)
fig
```


## Combined statistical representations with distplot figure factory

The distplot figure factory displays a combination of statistical representations of numerical data, such as histogram, kernel density estimation or normal curve, and rug plot.

#### Basic Distplot

A histogram, a kde plot and a rug plot are displayed.

```{r}
library(ggplot2)  
library(plotly)

set.seed(1)    
hist_data <- data.frame(rnorm(1000, mean = 0, sd = 1))   
colnames(hist_data) = c('x')  
gg <- ggplot(hist_data,aes(x = x, color = 'density')) +  
  geom_histogram(aes(y = ..density..), bins = 7,  fill = '#67B7D1', alpha = 0.5) +  
  geom_density(color = '#67B7D1') +  
  geom_rug(color = '#67B7D1') + 
  ylab("") + 
  xlab("")  + theme(legend.title=element_blank()) +
  scale_color_manual(values = c('density' = '#67B7D1'))


ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff')) 
```

#### Plot Multiple Datasets

```{r}
library(ggplot2)    
library(plotly) 
set.seed(1)    
x1 <- rnorm(200, mean = 0, sd = 1) - 2  
x2 <- rnorm(200, mean = 0, sd = 1)   
x3 <- rnorm(200, mean = 0, sd = 1) + 2  
x4 <- rnorm(200, mean = 0, sd = 1) +4  
group_labels = c('Group 1', 'Group 2', 'Group 3', 'Group 4')  

df1 <- data.frame(x1, group_labels[1])  
colnames(df1) <- c('x', 'Group') 
df2 <- data.frame(x2, group_labels[2]) 
colnames(df2) <- c('x', 'Group') 
df3 <- data.frame(x3, group_labels[3]) 
colnames(df3) <- c('x', 'Group') 
df4 <- data.frame(x4, group_labels[4]) 
colnames(df4) <- c('x', 'Group') 
df <- rbind(df1,df2,df3,df4) 
colnames(df) <- c('x', 'Group') 

gg <- ggplot(data = df ) +  
  geom_histogram(aes(x=x, y = ..density.., fill=Group),bins = 29, alpha = 0.7) + 
 geom_density(aes(x=x, color=Group)) + geom_rug(aes(x=x, color=Group))+ 
  ylab("") + 
  xlab("")


ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff')) 
```

#### Use Multiple Bin Sizes

Different bin sizes are used for the different datasets with the `bin_size` argument.

```{r}
library(ggplot2) 
library(plotly)

set.seed(1)  
x1 <- rnorm(1000, mean = 0, sd = 1) - 2
x2 <- rnorm(1000, mean = 0, sd = 1) 
x3 <- rnorm(1000, mean = 0, sd = 1) + 2
x4 <- rnorm(1000, mean = 0, sd = 1) +4
group_labels = c('Group 1', 'Group 2', 'Group 3', 'Group 4')

df = data.frame(x1,x2,x3,x4, group_labels)

gg <- ggplot(df,aes() ) + 
  geom_histogram(aes(x = x1, y = ..density.., fill = '#67B7D1'), alpha = 0.7, bins = 29) +
  geom_histogram(aes(x = x2, y = ..density.., fill = '#ff8080'), alpha = 0.7, bins = 20) +
  geom_histogram(aes(x = x3, y = ..density.., fill = '#ff99dd'), alpha = 0.7, bins = 10) +
  geom_histogram(aes(x = x4, y = ..density.., fill = '#ff9900'), alpha = 0.7, bins = 5) +
  geom_density(aes(x = x1),color = '#67B7D1') + 
  geom_density(aes(x = x2),color = '#ff8080') +
  geom_density(aes(x = x3),color = '#ff99dd') +
  geom_density(aes(x = x4),color = '#ff9900') +
  geom_rug(aes(x = x1),color = '#67B7D1') + 
  geom_rug(aes(x = x2),color = '#ff8080') +
  geom_rug(aes(x = x3),color = '#ff99dd') +
  geom_rug(aes(x = x4),color = '#ff9900') + 
  theme(legend.title=element_blank()) +
  scale_fill_identity(labels = c('Group 1', 'Group 2', 'Group 3', 'Group 4'),
                      guide = "legend")  +
  labs(x = '',
       y = '')

ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff')) 
```


#### Plot Normal Curve

```{r}
library(ggplot2)    
library(plotly) 
set.seed(1)    
x1 <- rnorm(200, mean = 0, sd = 1)  
x2 <- rnorm(200, mean = 0, sd = 1) + 2 
 
group_labels = c('Group 1', 'Group 2')  
 
df1 <- data.frame(x1, group_labels[1])  
colnames(df1) <- c('x', 'Group') 
df2 <- data.frame(x2, group_labels[2]) 
colnames(df2) <- c('x', 'Group') 
 
df <- rbind(df1,df2) 
colnames(df) <- c('x', 'Group') 
 
gg <- ggplot(data = df , aes(x=x)) +  
  geom_histogram(aes(y = ..density.., fill=Group),bins = 30, alpha = 0.7)+ 
  geom_density(aes(color=Group))+  
  geom_rug(aes(color=Group))+ 
  labs(x = '',  
       y = '',  
       title = 'Distplot with Normal Distribution')  

ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff')) 
```

#### Plot Only Curve and Rug

```{r}
library(ggplot2)    
library(plotly) 
set.seed(1)    
x1 <- rnorm(200, mean = 0, sd = 1) - 1
x2 <- rnorm(200, mean = 0, sd = 1) 
x3 <- rnorm(200, mean = 0, sd = 1) + 1

group_labels = c('Group 1', 'Group 2', 'Group 3')  

df1 <- data.frame(x1, group_labels[1])  
colnames(df1) <- c('x', 'Group') 
df2 <- data.frame(x2, group_labels[2]) 
colnames(df2) <- c('x', 'Group') 
df3 <- data.frame(x3, group_labels[3]) 
colnames(df3) <- c('x', 'Group') 

df <- rbind(df1,df2,df3) 
colnames(df) <- c('x', 'Group') 

gg <- ggplot(data = df ) +  
 geom_density(aes(x=x, color=Group)) + geom_rug(aes(x=x, color=Group)) + 
  ylab("") + 
  xlab("")

ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),
         title = 'Curve and Rug Plot') 
```

#### Plot Only Hist and Rug

```{r}
library(ggplot2)    
library(plotly) 
set.seed(1)    
x1 <- rnorm(200, mean = 0, sd = 1) - 1
x2 <- rnorm(200, mean = 0, sd = 1) 
x3 <- rnorm(200, mean = 0, sd = 1) + 1

group_labels = c('Group 1', 'Group 2', 'Group 3')  

df1 <- data.frame(x1, group_labels[1])  
colnames(df1) <- c('x', 'Group') 
df2 <- data.frame(x2, group_labels[2]) 
colnames(df2) <- c('x', 'Group') 
df3 <- data.frame(x3, group_labels[3]) 
colnames(df3) <- c('x', 'Group') 

df <- rbind(df1,df2,df3) 
colnames(df) <- c('x', 'Group') 

gg <- ggplot(data = df ) +  
  geom_histogram(aes(x=x, y = ..density.., fill=Group),bins = 29, alpha = 0.7) + 
  geom_rug(aes(x=x, color=Group)) + 
  ylab("") + 
  xlab("")


ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),
         title = 'Hist and Rug Plot') 
```

#### Plot Hist and Rug with Different Bin Sizes

```{r}
library(ggplot2)  
library(plotly)

set.seed(1) 
x1 <- rnorm(1000, mean = 0, sd = 1) - 2
x2 <- rnorm(1000, mean = 0, sd = 1) 
x3 <- rnorm(1000, mean = 0, sd = 1) + 2

df <- data.frame(x1, x2, x3)

gg <- ggplot(df,aes() ) + 
  geom_histogram(aes(x = x1, y = ..density.., fill = '#67B7D1'), alpha = 0.7, bins = 5) +
  geom_histogram(aes(x = x2, y = ..density.., fill = '#ff8080'), alpha = 0.7, bins = 17) +
  geom_histogram(aes(x = x3, y = ..density.., fill = '#ff99dd'), alpha = 0.7, bins = 29) +
  geom_rug(aes(x = x1),color = '#67B7D1') + 
  geom_rug(aes(x = x2),color = '#ff8080') +
  geom_rug(aes(x = x3),color = '#ff99dd') +
  labs(x = '',
       y = '',
       title = 'Hist and Rug Plot') +
  theme(legend.title=element_blank()) +
  scale_fill_identity(labels = c('Group 1', 'Group 2', 'Group 3'),
                      guide = "legend") 


ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),
         title = 'Hist and Rug Plot') 
```

#### Plot Only Hist and Curve

```{r}
library(ggplot2)    
library(plotly) 
set.seed(1)    
x1 <- rnorm(200, mean = 0, sd = 1) - 2  
x2 <- rnorm(200, mean = 0, sd = 1)   
x3 <- rnorm(200, mean = 0, sd = 1) + 2  
x4 <- rnorm(200, mean = 0, sd = 1) +4  
group_labels = c('Group 1', 'Group 2', 'Group 3', 'Group 4')  

df1 <- data.frame(x1, group_labels[1])  
colnames(df1) <- c('x', 'Group') 
df2 <- data.frame(x2, group_labels[2]) 
colnames(df2) <- c('x', 'Group') 
df3 <- data.frame(x3, group_labels[3]) 
colnames(df3) <- c('x', 'Group') 
df4 <- data.frame(x4, group_labels[4]) 
colnames(df4) <- c('x', 'Group') 
df <- rbind(df1,df2,df3,df4) 
colnames(df) <- c('x', 'Group') 

gg <- ggplot(data = df ) +  
  geom_histogram(aes(x=x, y = ..density.., fill=Group),bins = 29, alpha = 0.7) + 
 geom_density(aes(x=x, color=Group)) + 
  ylab("") + 
  xlab("")


ggplotly(gg)%>% 
  layout(plot_bgcolor='#e5ecf6',   
             xaxis = list(   
               title='Time', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),   
             yaxis = list(   
               title='Value A', 
               zerolinecolor = '#ffff',   
               zerolinewidth = 2,   
               gridcolor = 'ffff'),
         title = 'Hist and Curve Plot') 
```


